---
title: "Testing cross-populations PGS methods"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
knitr::opts_knit$set(root.dir = '/scratch/prj/oliverpainfel/Software/MyGit/GenoPred/GenoPredPipe')

library(data.table)

```

***

We will test using the mini test data to speed up testing. The configuration files are in ~/oliverpainfel/Software/MyGit/GenoPred/pipeline/misc/cross_pop_test.

```{bash}
snakemake -j1 --use-conda --configfile=misc/cross_pop_test/config_cross_pop.yaml output_all -n
```

***

Test the new set up with OpenSNP data as the target. Initially restrict to chr22 to see if it works as expected.

We need to edit the config file used in the previous run on OpenSNP data.

<details><summary>Show code</summary>

```{r}
setwd('/users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline')

config <- readLines('misc/opensnp/config.yaml')

config[grepl('^config_file:', config)]<- 'config_file: misc/opensnp/config_cross_pop.yaml'
config<-config[-which(grepl('^testing:', config))]
config <- config[!grepl('^score_list:', config)]
config[grepl('^outdir:', config)]<- 'outdir: /users/k1806347/oliverpainfel/Data/OpenSNP/GenoPred/test_cross_pop_1'
config[grepl('^pgs_methods:', config)]<- "pgs_methods: ['ptclump','tlprs']"
config<-c(config, 'gwas_groups: misc/cross_pop_test/gwas_groups.txt')
config<-c(config, "tlprs_methods: ['ptclump']")

write.table(config, 'misc/opensnp/config_cross_pop.yaml', col.names = F, row.names = F, quote = F)

```

```{bash}
snakemake --profile slurm --use-conda --configfile=misc/opensnp/config_cross_pop.yaml output_all -n
```

## Evaluate PGS

<details><summary>Show code</summary>

```{r}
# Test correlation between PGS and observed height

setwd('/users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/')
library(data.table)
library(ggplot2)
library(cowplot)

source('../functions/misc.R')
source_all('../functions')

# Read in pheno data
pheno <- fread('/users/k1806347/oliverpainfel/Data/OpenSNP/processed/pheno/height.txt')

# Read in PGS
pgs <- read_pgs(config = 'misc/opensnp/config_cross_pop.yaml', name = 'opensnp')$opensnp

# Estimate correlation between pheno and pgs
cor <- NULL
for(pop_i in names(pgs)){
  for(gwas_i in names(pgs[[pop_i]])){
    for(pgs_method_i in names(pgs[[pop_i]][[gwas_i]])){
      pgs_i <- pgs[[pop_i]][[gwas_i]][[pgs_method_i]]
      pheno_pgs<-merge(pheno, pgs_i, by = c('FID','IID'))
      
      for(model_i in names(pgs_i)[-1:-2]){
        y <- scale(pheno_pgs$height)
        x <- scale(pheno_pgs[[model_i]])
        
        if(all(is.na(x))){
          next
        }
        
        coef_i <- coef(summary(mod <- lm(y ~ x)))
        
        tmp <- data.table(
          pop = pop_i,
          gwas = gwas_i,
          pgs_method = pgs_method_i,
          name = model_i,
          r = coef_i[2,1],
          se = coef_i[2,2],
          p = coef_i[2,4],
          n = nobs(mod))
      cor <- rbind(cor, tmp)
      
      }
    }
  }
}

# The European sample is the only one large enough for interpretable results
# Subset EUR results
cor_eur <- cor[cor$pop == 'EUR', ]

# Restrict to best and and pseudoval only
cor_eur_subset <- NULL
for(pop_i in unique(cor_eur$pop)){
  for(gwas_i in unique(cor_eur$gwas[cor_eur$pop == pop_i])){
    for(pgs_method_i in unique(cor_eur$pgs_method[cor_eur$pop == pop_i & cor_eur$gwas == gwas_i])){
      
      # Subset relevant results
      cor_eur_i <- cor_eur[
        cor_eur$pop == pop_i &
        cor_eur$gwas == gwas_i &
        cor_eur$pgs_method == pgs_method_i,]
  
      # Top R
      if(pgs_method_i %in% c('ptclump','ldpred2','megaprs','prscs','prscsx','lassosum','dbslmm') | any(grepl('tlprs', pgs_method_i))){
        top_i <- cor_eur_i[which(cor_eur_i$r == max(cor_eur_i$r, na.rm = T))[1],]
        top_i$model <- 'Top'
        cor_eur_subset <- rbind(cor_eur_subset, top_i)
      }
  
      # PseudoVal
      if(pgs_method_i %in% c('ptclump','sbayesr','ldpred2','megaprs','prscs','prscsx','lassosum','dbslmm')){
        cor_eur_i$name <- gsub(paste0(gwas_i, '_'), '', cor_eur_i$name)
        pseudo_param <- find_pseudo(config = 'misc/opensnp/config_cross_pop.yaml', gwas = gwas_i, pgs_method = pgs_method_i)
        pseudo_i <- cor_eur_i[cor_eur_i$name == pseudo_param,]
        pseudo_i$model <- 'Pseudo'
        cor_eur_subset <- rbind(cor_eur_subset, pseudo_i)
      }
    }
  }
}

# Plot the results
cor_eur_subset$model <- factor(cor_eur_subset$model, levels = c('Top','Pseudo','External'))
dir.create('/users/k1806347/oliverpainfel/Software/MyGit/GenoPred/docs/Images/OpenSNP')

plot_obj_eur <- 
  ggplot(cor_eur_subset, aes(x = pgs_method, y = r, fill = model)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), width = 0.7) +
  geom_errorbar(
    aes(ymin = r - se, ymax = r + se),
    width = .2,
    position = position_dodge(width = 0.7)
  ) +
  labs(
    y = "Correlation (SE)",
    x = 'PGS Method',
    fill = 'Model',
    title = paste0("OpenSNP - EUR\n(N = ", cor_eur_subset$n[1], ")")
  ) +
  theme_half_open() +
  background_grid() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, size=12)) +
  facet_grid(. ~ gwas)


cor_eur_tlprs<-cor_eur[cor_eur$pgs_method == 'tlprs_ptclump',]
cor_eur_tlprs[grepl('EUR_0_0.01', cor_eur_tlprs$name),]


cor_eas <- cor[cor$pop == 'EAS', ]
# Restrict to best and and pseudoval only
cor_eas_subset <- NULL
for(pop_i in unique(cor_eas$pop)){
  for(gwas_i in unique(cor_eas$gwas[cor_eas$pop == pop_i])){
    for(pgs_method_i in unique(cor_eas$pgs_method[cor_eas$pop == pop_i & cor_eas$gwas == gwas_i])){
      
      # Subset relevant results
      cor_eas_i <- cor_eas[
        cor_eas$pop == pop_i &
        cor_eas$gwas == gwas_i &
        cor_eas$pgs_method == pgs_method_i,]
  
      # Top R
      if(pgs_method_i %in% c('ptclump','ldpred2','megaprs','prscs','prscsx','lassosum','dbslmm') | any(grepl('tlprs', pgs_method_i))){
        top_i <- cor_eas_i[which(cor_eas_i$r == max(cor_eas_i$r, na.rm = T))[1],]
        top_i$model <- 'Top'
        cor_eas_subset <- rbind(cor_eas_subset, top_i)
      }
  
      # PseudoVal
      if(pgs_method_i %in% c('ptclump','sbayesr','ldpred2','megaprs','prscs','prscsx','lassosum','dbslmm')){
        cor_eas_i$name <- gsub(paste0(gwas_i, '_'), '', cor_eas_i$name)
        pseudo_param <- find_pseudo(config = 'misc/opensnp/config_cross_pop.yaml', gwas = gwas_i, pgs_method = pgs_method_i)
        pseudo_i <- cor_eas_i[cor_eas_i$name == pseudo_param,]
        pseudo_i$model <- 'Pseudo'
        cor_eas_subset <- rbind(cor_eas_subset, pseudo_i)
      }
    }
  }
}

# Plot the results
cor_eas_subset$model <- factor(cor_eur_subset$model, levels = c('Top','Pseudo','External'))
dir.create('/users/k1806347/oliverpainfel/Software/MyGit/GenoPred/docs/Images/OpenSNP')
cor_eas_tlprs<-cor_eas[cor_eas$pgs_method == 'tlprs_ptclump',]

```
