---
title: "Integration tests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
knitr::opts_knit$set(root.dir = '/scratch/prj/oliverpainfel/Software/MyGit/GenoPred/GenoPredPipe')

library(data.table)

```

This document describes the setup and execution of integration tests.

***

# Setup

I have run GenoPredPipe using the test data. Store an .rds file containing some core outputs that can be used for comparison in the future as I make further developments.

```{r}

# Create a function that reads in key outputs
read_genopredpipe_out <- function(x) {
  # Create an empty list to populate with reference outputs
  ref <- list()
  
  ###########################
  # 1KG reference
  ###########################
  
  ref$onekg <- list()
  
  ##########
  # bim
  ##########
  
  ref_bim <- NULL
  for (i in 1:22) {
    ref_bim <-
      rbind(ref_bim, fread(
        paste0(x, '/resources/data/1kg/1KGPhase3.w_hm3.chr', i, '.bim')
      ))
  }
  
  ref$onekg$bim <- ref_bim
  
  ##########
  # populations
  ##########
  
  keep_list <-
    fread(paste0(x, '/resources/data/1kg/super_pop_keep.list'), header = F)
  
  ref_keep <- list()
  for (i in 1:nrow(keep_list)) {
    ref_keep[[keep_list$V1[i]]] <-
      fread(paste0(x, '/', keep_list$V2[i]), header = F)
  }
  
  ref$onekg$pops <- ref_keep
  
  ##########
  # ref pcs
  ##########
  
  ref_pcs <- list()
  for (i in 1:nrow(keep_list)) {
    ref_pcs[[keep_list$V1[i]]] <-
      fread(
          paste0(
            x,
            '/resources/data/1kg/pc_score_files/',
            keep_list$V1[i],
            '/1KGPhase3.w_hm3.',
            keep_list$V1[i],
            '.eigenvec.var'
          ), nrows=1000
      )
  }
  
  ref$onekg$pcs <- ref_pcs
  
  ###########################
  # GWAS sumstats
  ###########################
  
  gwas_list <- c('COAD01', 'BODY04')
  
  ref$sumstats <- list()
  for (i in gwas_list) {
    ref$sumstats[[i]] <-
      fread(
        paste0(
          x,
          '/resources/data/gwas_sumstat/',
          i,
          '/',
          i,
          '.cleaned.gz'
      ), nrows=1000)
  }
  
  ###########################
  # PGS score files
  ###########################
  
  ref$score <- list()
  
  ########
  # ptclump
  ########
  
  ref$score$ptclump <- list()
  for (i in gwas_list) {
    ref$score$ptclump[[i]] <-
      fread(
        paste0(
          x,
          '/resources/data/1kg/prs_score_files/pt_clump/',
          i,
          '/1KGPhase3.w_hm3.',
          i,
          '.score.gz'
        ), nrows=1000
      )
  }
  
  ########
  # dbslmm
  ########
  
  ref$score$dbslmm <- list()
  for (i in gwas_list) {
    ref$score$dbslmm[[i]] <-
      fread(
        paste0(
          x,
          '/resources/data/1kg/prs_score_files/dbslmm/',
          i,
          '/1KGPhase3.w_hm3.',
          i,
          '.score'
        ), nrows=1000
      )
  }
  
  ###########################
  # Target data
  ###########################
  
  targ_names <- c('example_bgen', 'example_plink1', 'example_vcf')
  ref$target <- list()
  
  ##########
  # bim
  ##########
  
  for (j in targ_names) {
    ref$target[[j]] <- list()
    ref$target[[j]]$bim <- NULL
    for (i in 1:22) {
      ref$target[[j]]$bim <-
        rbind(ref$target[[j]]$bim, fread(
          paste0(
            x,
            '/resources/data/target_output/',
            j,
            '/',
            j,
            '.1KGphase3.hm3.chr',
            i,
            '.bim'
          )
        ))
    }
  }
  
  ############
  # ancestry
  ############
  
  ref$target$example_plink1$ancestry <- list()
  ref$target$example_plink1$ancestry$eigenvec <- fread(
    paste0(
      x,
      '/resources/data/target_output/example_plink1/ancestry/ancestry_all/example_plink1.Ancestry.eigenvec'
    )
  )
  ref$target$example_plink1$ancestry$probs <- fread(
    paste0(
      x,
      '/resources/data/target_output/example_plink1/ancestry/ancestry_all/example_plink1.Ancestry.model_pred'
    )
  )
  
  ############
  # PCs
  ############
  
  ref$target$example_plink1$pcs <- list()
  for (i in 1:nrow(keep_list)) {
    ref$target$example_plink1$pcs[[i]] <-
      fread(
        paste0(
          x,
          '/resources/data/target_output/example_plink1/projected_pc/',
          keep_list$V1[i],
          '/example_plink1.',
          keep_list$V1[i],
          '.eigenvec'
        )
      )
  }
  
  ############
  # PGS
  ############
  
  ref$target$example_plink1$pgs <- list()
  
  ####
  # ptclump
  ####
  
  ref$target$example_plink1$pgs$ptclump <- list()
  
  for (i in keep_list$V1) {
    ref$target$example_plink1$pgs$ptclump[[i]] <- list()
    ref$target$example_plink1$pgs$ptclump[[i]]$COAD01 <-
      fread(
        paste0(
          x,
          '/resources/data/target_output/example_plink1/prs/',
          i,
          '/pt_clump/COAD01/example_plink1.COAD01.',
          i,
          '.profiles'
        )
      )
  }
  
  ####
  # dbslmm
  ####
  
  for (i in keep_list$V1) {
    ref$target$example_plink1$pgs$dbslmm[[i]] <- list()
    ref$target$example_plink1$pgs$dbslmm[[i]]$COAD01 <-
      fread(
        paste0(
          x,
          '/resources/data/target_output/example_plink1/prs/',
          i,
          '/dbslmm/COAD01/example_plink1.COAD01.',
          i,
          '.profiles'
        )
      )
  }
  
  return(ref)
}

# Generate reference for comparison
ref<-read_genopredpipe_out('/scratch/prj/oliverpainfel/Software/MyGit/GenoPred/GenoPredPipe')
  
```

**Note.** This does not include all outputs from the pipeline, only those that are generated when running the following command with the test data:

```{bash}
snakemake --profile slurm --use-conda run_create_reports
```

***

# Compare new runs to reference

```{r}

# Generate the rds with key output for test data
targ<-read_genopredpipe_out('/users/k1806347/oliverpainfel/test/GenoPred/GenoPredPipe')

# all.equal tests whether objects are identical and returns details of differences
all.equal(ref, targ)

```

# Save the ref data for future comparisons

```{r}

saveRDS(ref, '/scratch/prj/oliverpainfel/Software/MyGit/GenoPred/GenoPredPipe/private/test_ref.rds')

```















