---
title: "GenoPredPipe"
params:
  name: ""
  output: ""
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 3
    toc_float: true
    fig_caption: yes

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')

library(data.table)
library(knitr)
library(ggplot2)
library(cowplot)
library(kableExtra)
library(stringr)

knit_hooks$set(optipng = hook_optipng)
knit_hooks$set(pngquant = hook_pngquant)

```

***

<font size="4"> 
ID: `r gsub('_',' ',params$name)`
</font>

***

# Descriptives

```{R, eval=T, echo=F}

harmoniser_log<-readLines(paste0(params$output,'/',params$name,'/',params$name,'.1KGphase3.hm3.chr.hm3_harmoniser.log'))

fam<-fread(paste0(params$output,'/',params$name,'/',params$name,'.1KGphase3.hm3.chr22.fam'))

ref_bim<-NULL
for(i in 1:22){
  ref_bim<-rbind(ref_bim, fread(paste0('resources/data/1kg/1KGPhase3.w_hm3.chr',i,'.bim')))
}
nvar_in_ref<-nrow(ref_bim)

ref_overlap<-gsub('%.*','%',harmoniser_log[grepl('of variants in reference are in the target.',harmoniser_log)])

ref_overlap_after_imp<-gsub('%.*','%',harmoniser_log[grepl('of variants in reference are in the target.',harmoniser_log)])

```

This sample contains `r nrow(fam)` individuals. The sample had undergone genotype imputation in advance of GenoPredPipe. Of the `r nvar_in_ref` variants in the reference, `r ref_overlap_after_imp` were available as hard calls in the target.

***

# Ancestry

## Super Population Classification

```{R, eval=T, echo=F, message=F, warning=F}
# Read in 1KG population codes
pop_codes<-fread('resources/data/1kg/1kg_pop_codes.tsv')
pop_codes<-pop_codes[,1:3]
names(pop_codes)<-c('pop','code','super_code')
pop_codes<-pop_codes[pop_codes$code != '',]

super_pop_codes<-fread('resources/data/1kg/1kg_super_pop_codes.tsv')
names(super_pop_codes)<-c('pop','code')

model_pred<-fread(paste0(params$output,'/',params$name,'/ancestry/ancestry_all/',params$name,'.Ancestry.model_pred'))
probs_melt<-melt(model_pred[,-1:-2])

model_pred$max_prob<-apply(model_pred[,-1:-2], 1, max)
n_hard_call<-sum(model_pred$max_prob > 0.5)
model_pred<-model_pred[model_pred$max_prob > 0.5,]
model_pred$max_prob<-NULL

model_pred_hard<-NULL
for(i in names(model_pred[,-1:-2])){
	tmp_keep<-model_pred[apply(model_pred[,-1:-2], 1, function(x) x[i] == max(x)),1:2]
	model_pred_hard<-rbind(model_pred_hard, data.frame(tmp_keep, Population=i)) 
}

hard_call_table<-data.frame(melt(table(model_pred_hard$Population)))
hard_call_table<-merge(hard_call_table, super_pop_codes, by.x='Var1',by.y='code')
hard_call_table<-hard_call_table[,c('pop','value')]
names(hard_call_table)<-c('Super Population', 'N')

```

Of the `r nrow(fam)` individuals in the sample, the ancestry model classified `r n_hard_call` individuals to a super population with a probability >50%. The number of individuals assigned to each super population were:

```{r, eval=T, echo=F}
kable(hard_call_table, align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

***

```{R, eval=T, echo=F, message=F, warning=F, fig.cap="Distribution of Super Population Probabilities", fig.dim = c(8, 4)}

ggplot(probs_melt, aes(x=value, fill=variable)) + 
  geom_histogram(alpha=0.5) +
  labs(x='Probability', y='N', fill='Super Population') +
  xlim(c(-0.05,1.05)) +
  theme_minimal_hgrid()

```

***

<details><summary>Show principal component plots</summary>

```{r, results='asis', echo=F}
cat(paste0("![Principal Component Scores Compared to Reference Super Populations](",normalizePath(params$output, mustWork = FALSE),'/',params$name,'/ancestry/ancestry_all/', params$name,".Ancestry.PCs_plot_super_pop.png)"))
```

```{r, eval=T, echo=F}
kable(super_pop_codes, align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

</details>

***
