---
title: Cross-population evaluation of polygenic scores
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 2
    css: styles/styles.css
    includes:
      in_header: header.html
      after_body: footer.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{css, echo=F}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

***

In this study, we will investigate the utility if polygenic scores for stratifying individual into aetiological subtypes, and the predictive utility of modelling interactions between polygenic score and subtype membership. Our initial investigation will look at pathways specific PGS.

We have previously conducted a very simple proof of concept study [here](https://opain.github.io/GenoClust/proof_of_principle.html). Here will conduct a more realistic simulation to determine feasibility and refine the methodology, and then apply the methodology to complex disease of interest.

Things to think about:
- Derive clusters using cases only, and then project in to full sample, or derive clusters within full sample?
- Clustering method: K-means, Gaussian mixture model, hierarchical clustering

***

# Proof-of-concept

## Simulated data

We could simulate genotype and phenotype data using HAPNEST, two create two phenotypes with causal variants within different pathways.

Simulate two binary phenotypes, with a genetic correlation of 0. Specify causal variants present within a gene set for each disease:
- Disease 1: GO:0006281 DNA repair
- Disease 2: GO:0048167 — Regulation of synaptic plasticity

***

### Simulate genetic data

Simulate chr22 genotype data for 100k EUR individuals, using HAPNEST.

<details><summary>Show code</summary>

```{bash}
cd ~/oliverpainfel/Data/HAPNEST

# Generate genotype and phenotype data
singularity exec \
  --bind data/:/data/ \
  --bind /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/misc/hapnest/config.synth_2.yaml:/data/config.synth_2.yaml \
  containers/intervene-synthetic-data_latest.sif \
  generate_geno \
  8 \
  data/config.synth_2.yaml
  
# Keep getting an error due to memory allocation. In the end I ran it with 1 core
  
```

</details>

***

### Simulate phenotype data

We want to simulate pairs of phenotypes that have a range of heritability, polygenicity, and genetic correlation. Specifically, we want to ensure these traits are enriched for gene-sets, with differing degrees of overlap between the pairs of traits. 

```{r}
#conda activate .snakemake/conda/329e0288cb99508f5e6c50a0996b234c_
library(bigsnpr)

# Attach the bigSNP object from a previously saved `.rds` file
rds <- snp_readBed('~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/synth_2_chr-22.bed', backingfile = tempfile())
bigSNP <- snp_attach(rds)
G <- bigSNP$genotypes
map <- bigSNP$map

# Read gene ontology sets (.gmt)
gmt <- readLines('/scratch/prj/oliverpainfel/recovered/Data/MSigDB/c5.go.bp.v7.5.1.entrez.gmt')
gene_sets <- strsplit(gmt, "\t")
gene_list <- setNames(lapply(gene_sets, `[`, -c(1,2)), sapply(gene_sets, `[`, 1))

# Read in gene annotations
gene_loc <- fread('/scratch/prj/oliverpainfel/recovered/Data/Gene_Locations/MAGMA_NCBI37.3/NCBI37.3.gene.loc')
names(gene_loc)<-c('ID','CHR','START','STOP','STRAND','NAME')
gene_loc <- gene_loc[gene_loc$CHR == 22,] #out simulation will only use chr22 data

# Annotate SNPs with genes
map$gene <- sapply(1:nrow(map), function(i) {
  chr <- map$chromosome[i]
  pos <- map$physical.pos[i]
  gene_match <- gene_annot %>% 
    filter(chr == chromosome & pos >= start & pos <= end) %>%
    pull(gene_symbol)
  if(length(gene_match) > 0) paste(gene_match, collapse = ",") else NA
})

# SNPs in enriched gene sets
snp_gene_sets <- lapply(gene_sets, function(genes) {
  snps_in_set <- map %>% filter(grepl(paste(genes, collapse="|"), gene))
  snps_in_set$snp_id
})

# Set simulation parameters
set.seed(1)

h2_trait1 <- 0.3
h2_trait2 <- 0.4
rg <- 0.5  # genetic correlation
polygenicity_fraction <- 0.01  # fraction of causal SNPs
overlap_fraction <- 0.5  # fraction of overlapping causal SNPs

# Select causal variants for each trait
n_snps <- ncol(G)
n_causal <- ceiling(polygenicity_fraction * n_snps)
causal_snps_trait1 <- sample(unlist(snp_gene_sets), n_causal)

# Overlap between traits
n_overlap <- ceiling(overlap_fraction * n_causal)
causal_snps_overlap <- sample(causal_snps_trait1, n_overlap)
remaining_snps <- setdiff(unlist(snp_gene_sets), causal_snps_overlap)
causal_snps_trait2 <- c(causal_snps_overlap, sample(remaining_snps, n_causal - n_overlap))

# Generate effect sizes for each trait
library(MASS)
Sigma <- matrix(c(1, rg, rg, 1), ncol=2)
betas_raw <- mvrnorm(n_causal, mu=c(0,0), Sigma=Sigma)

# Adjust betas to meet desired heritability
scale_betas <- function(G, snps, h2, betas_raw_col) {
  G_sub <- G[, snps]
  var_g <- var(big_prodVec(G_sub, betas_raw_col))
  scale_factor <- sqrt(h2 / var_g)
  betas_raw_col * scale_factor
}

beta_trait1 <- scale_betas(G, causal_snps_trait1, h2_trait1, betas_raw[,1])
beta_trait2 <- scale_betas(G, causal_snps_trait2, h2_trait2, betas_raw[,2])

# Simulate phenotyp data
n_indiv <- nrow(G)

simulate_pheno <- function(G, causal_snps, betas, h2) {
  genetic_component <- big_prodVec(G[, causal_snps], betas)
  env_sd <- sqrt(var(genetic_component)*(1-h2)/h2)
  genetic_component + rnorm(n_indiv, mean=0, sd=env_sd)
}

pheno_trait1 <- simulate_pheno(G, causal_snps_trait1, beta_trait1, h2_trait1)
pheno_trait2 <- simulate_pheno(G, causal_snps_trait2, beta_trait2, h2_trait2)

phenotypes <- data.frame(trait1 = pheno_trait1, trait2 = pheno_trait2)

# Read in causal variants
library(data.table)
effect_sizes<-fread('~/oliverpainfel/Data/HAPNEST/synth_2/causal.snplist')
names(effect_sizes)<-c('rsid','trait_1','trait_2')

rsids<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/rsid_map_list_chr22.txt')
effect_sizes<-merge(effect_sizes, rsids, by='rsid')

h2 <- 0.05  # Desired heritability
set.seed(1)
traits<-list()
for(i in 1:2){
  effect_sizes_i<-effect_sizes[effect_sizes[[paste0('trait_',i)]] == 1,]
  causal_indices <- match(effect_sizes_i$id_hg38, snp_names)
  G_causal <- big_prodVec(G, effect_sizes_i[[paste0('trait_',i)]], ind.col = causal_indices)
  var_G <- var(G_causal)
  var_E <- (var_G / h2) - var_G
  E <- rnorm(length(G_causal), mean = 0, sd = sqrt(var_E))
  Y <- G_causal + E
  traits[[paste0('trait_', i)]]<-Y
  print(var(G_causal) / (var(G_causal) + var(E)))
}

cor(do.call(cbind, traits))

library(bigsnpr)

# Prepare the phenotype
Y1 <- traits[["trait_1"]]
Y2 <- traits[["trait_2"]]

# Perform GWAS for trait 1
gwas_1 <- big_univLinReg(G, Y1)

# Perform GWAS for trait 2
gwas_2 <- big_univLinReg(G, Y2)

# Combine results into a data frame
gwas_results <- data.frame(
  SNP = snp_names,
  beta_trait1 = gwas_1$estim,
  beta_trait2 = gwas_2$estim,
  se_trait1 = gwas_1$std.err,
  se_trait2 = gwas_2$std.err
)

gwas_results$p_trait1 <- 2*pnorm(q=abs(gwas_results$beta_trait1/gwas_results$se_trait1), lower.tail=FALSE)
gwas_results$p_trait2 <- 2*pnorm(q=abs(gwas_results$beta_trait2/gwas_results$se_trait2), lower.tail=FALSE)

# Looks good. Lets proceed with these simulated phenotypes
write.table(data.frame(FID = bigSNP$fam$family.ID, IID = bigSNP$fam$sample.ID, outcome = traits[["trait_1"]]), 
            '~/oliverpainfel/Data/HAPNEST/synth_2/trait_1.txt', row.names = F, quote=F)
write.table(data.frame(FID = bigSNP$fam$family.ID, IID = bigSNP$fam$sample.ID, outcome = traits[["trait_2"]]), 
            '~/oliverpainfel/Data/HAPNEST/synth_2/trait_2.txt', row.names = F, quote=F)

```

***

### Define causal variants

Download GO data to identify genes within these sets. 

```{r}
# Read in .gmt file
gmt <- readLines('/scratch/prj/oliverpainfel/recovered/Data/MSigDB/c5.go.bp.v7.5.1.entrez.gmt')
gene_sets <- strsplit(gmt, "\t")
gene_list <- setNames(lapply(gene_sets, `[`, -c(1,2)), sapply(gene_sets, `[`, 1))

# Subset sets of interest
gene_list <- gene_list[ c('GOBP_DNA_REPAIR', 'GOBP_REGULATION_OF_SYNAPTIC_PLASTICITY')]

# Identify variants within genes within each set
gene_loc <- fread('/scratch/prj/oliverpainfel/recovered/Data/Gene_Locations/MAGMA_NCBI37.3/NCBI37.3.gene.loc')
names(gene_loc)<-c('ID','CHR','START','STOP','STRAND','NAME')
gene_loc <- gene_loc[gene_loc$CHR == 22,] #out simulation will only use chr22 data

# Read in hm3 SNP-list
rsids<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/rsid_map_list_chr22.txt')
snplist<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/hapmap_variant_list_chr22.txt', header=F)$V1
snplist_tab<-data.frame(do.call(rbind, lapply(strsplit(snplist, split = ':'), function(x) c(x))))
names(snplist_tab) <- c('CHR', 'BP', 'A1', 'A2')
snplist_tab$CHR <- gsub('chr', '', snplist_tab$CHR)
rsids<-rsids[match(snplist, rsids$id_hg38),]
snplist_tab$SNP <- rsids$rsid

# Create SNP-list for each gene set
dir.create('~/oliverpainfel/Data/HAPNEST/synth_2', recursive = T)

set.seed(1)
snp_sets <- list()
for(i in names(gene_list)){
  snp_sets[[i]] <- NULL
  genes <- gene_list[[i]]
  genes <- genes[genes %in% gene_loc$ID]
  for(k in genes){
    CHR <- gene_loc$CHR[gene_loc$ID == k] 
    START <- gene_loc$START[gene_loc$ID == k] 
    STOP <- gene_loc$STOP[gene_loc$ID == k]
    
    snp_set <- snplist_tab$SNP[which(snplist_tab$CHR == CHR & snplist_tab$BP >= START & snplist_tab$BP <= STOP)]
    if(!(length(snp_set) > 0)){
      next
    }
    
    # Select 1 variant from each gene to be causal
    snp_sets[[i]] <- c(snp_sets[[i]], sample(snp_set, 1))
  }
  snp_sets[[i]] <- unique(snp_sets[[i]])
}

# 
causal_var<-data.frame(SNP = unlist(snp_sets))
for(i in 1:length(snp_sets)){
  causal_var[[paste0('Trait_', i)]]<-as.numeric(causal_var$SNP %in% snp_sets[[i]])
}

# Save as csv
write.table(causal_var, '~/oliverpainfel/Data/HAPNEST/synth_2/causal.snplist', row.names = F, col.names = F, quote = F, sep=',')

```

***

### Run HAPNEST

<details><summary>Show code</summary>

```{bash}
cd ~/oliverpainfel/Data/HAPNEST

# Generate genotype and phenotype data
singularity exec \
  --bind data/:/data/ \
  --bind /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/misc/hapnest/config.synth_2.yaml:/data/config.synth_2.yaml \
  containers/intervene-synthetic-data_latest.sif \
  generate_geno \
  8 \
  data/config.synth_2.yaml
  
# Keep getting an error due to memory allocation. In the end I ran it with 1 core

singularity exec \
  --bind data/:/data/ \
  --bind /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/misc/hapnest/config.synth_2.yaml:/data/config.synth_2.yaml \
  containers/intervene-synthetic-data_latest.sif \
  generate_pheno \
  data/config.synth_2.yaml
  
# Can't figure out how run with causal variants specified. Lets do this using R
  
```

```{r}
#conda activate .snakemake/conda/329e0288cb99508f5e6c50a0996b234c_
library(bigsnpr)

# Attach the bigSNP object from a previously saved `.rds` file
rds <- snp_readBed('~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/synth_2_chr-22.bed', backingfile = tempfile())
bigSNP <- snp_attach(rds)

# Access the genotype matrix
G <- bigSNP$genotypes

# Get SNP names
snp_names <- bigSNP$map$marker.ID

# Read in causal variants
library(data.table)
effect_sizes<-fread('~/oliverpainfel/Data/HAPNEST/synth_2/causal.snplist')
names(effect_sizes)<-c('rsid','trait_1','trait_2')

rsids<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/rsid_map_list_chr22.txt')
effect_sizes<-merge(effect_sizes, rsids, by='rsid')

h2 <- 0.05  # Desired heritability
set.seed(1)
traits<-list()
for(i in 1:2){
  effect_sizes_i<-effect_sizes[effect_sizes[[paste0('trait_',i)]] == 1,]
  causal_indices <- match(effect_sizes_i$id_hg38, snp_names)
  G_causal <- big_prodVec(G, effect_sizes_i[[paste0('trait_',i)]], ind.col = causal_indices)
  var_G <- var(G_causal)
  var_E <- (var_G / h2) - var_G
  E <- rnorm(length(G_causal), mean = 0, sd = sqrt(var_E))
  Y <- G_causal + E
  traits[[paste0('trait_', i)]]<-Y
  print(var(G_causal) / (var(G_causal) + var(E)))
}

cor(do.call(cbind, traits))

library(bigsnpr)

# Prepare the phenotype
Y1 <- traits[["trait_1"]]
Y2 <- traits[["trait_2"]]

# Perform GWAS for trait 1
gwas_1 <- big_univLinReg(G, Y1)

# Perform GWAS for trait 2
gwas_2 <- big_univLinReg(G, Y2)

# Combine results into a data frame
gwas_results <- data.frame(
  SNP = snp_names,
  beta_trait1 = gwas_1$estim,
  beta_trait2 = gwas_2$estim,
  se_trait1 = gwas_1$std.err,
  se_trait2 = gwas_2$std.err
)

gwas_results$p_trait1 <- 2*pnorm(q=abs(gwas_results$beta_trait1/gwas_results$se_trait1), lower.tail=FALSE)
gwas_results$p_trait2 <- 2*pnorm(q=abs(gwas_results$beta_trait2/gwas_results$se_trait2), lower.tail=FALSE)

# Looks good. Lets proceed with these simulated phenotypes
write.table(data.frame(FID = bigSNP$fam$family.ID, IID = bigSNP$fam$sample.ID, outcome = traits[["trait_1"]]), 
            '~/oliverpainfel/Data/HAPNEST/synth_2/trait_1.txt', row.names = F, quote=F)
write.table(data.frame(FID = bigSNP$fam$family.ID, IID = bigSNP$fam$sample.ID, outcome = traits[["trait_2"]]), 
            '~/oliverpainfel/Data/HAPNEST/synth_2/trait_2.txt', row.names = F, quote=F)

```

</details>

***

### Run GSEA

We need to run a GWAS of each phenotyepe, and then MAGMA to see whether our simulation has lead to an enrichment of selected pathways

***

#### Perform PCA

<details><summary>Show code</summary>

```{r}
# conda activate .snakemake/conda/ea13b6c549c70695534894feeeecf0b3_
setwd('~/oliverpainfel/GenoPred/pipeline/')

start.time <- Sys.time()
library("optparse")

option_list = list(
  make_option("--target_plink_chr", action="store", default=NULL, type='character',
              help="Path to per chromosome target PLINK files [required]"),
  make_option("--maf", action="store", default=0.05, type='numeric',
              help="Minor allele frequency threshold [optional]"),
  make_option("--geno", action="store", default=0.02, type='numeric',
              help="Variant missingness threshold [optional]"),
  make_option("--hwe", action="store", default=1e-6, type='numeric',
              help="Hardy Weinberg p-value threshold. [optional]"),
  make_option("--n_pcs", action="store", default=10, type='numeric',
              help="Number of PCs (min=4) [optional]"),
  make_option("--plink2", action="store", default='plink2', type='character',
              help="Path PLINK2 software binary [required]"),
  make_option("--keep_list", action="store", default=NULL, type='character',
              help="File containing list of keep files and corresponding population code [optional]"),
  make_option("--unrel", action="store", default=NA, type='character',
              help="File containing list of unrelated individuals [optional]"),
  make_option("--n_cores", action="store", default=1, type='numeric',
              help="Number of cores for parallel computing [optional]"),
  make_option("--test", action="store", default=NA, type='character',
              help="Specify test mode [optional]"),
  make_option("--output", action="store", default=NULL, type='character',
              help="Path for output files [required]")
)

opt = parse_args(OptionParser(option_list=option_list))

opt$target_plink_chr<-'~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/synth_2_chr-'
opt$output<-'~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/pca/'
opt$test<-'chr22'

library(GenoUtils)
library(data.table)
source('../functions/misc.R')
source_all('../functions')

# Create temp directory
tmp_dir<-tempdir()

if(!is.na(opt$test)){
  CHROMS <- as.numeric(gsub('chr','',opt$test))
}

############
# Create file listing variants in regions of long range LD
############

targ_pvar <- read_bim(opt$target_plink_chr, chr = CHROMS)
targ_pvar <- remove_regions(dat = targ_pvar, regions = long_ld_coord)

###########
# Perform PCA on QC'd and independent variants
###########

# Create QC'd SNP-list
target_qc_snplist <- plink_qc_snplist(bfile = opt$target_plink_chr, plink2 = opt$plink2, chr = CHROMS, maf = opt$maf, geno = opt$geno, hwe = opt$hwe, threads = opt$n_cores)

# Remove high LD regions
target_qc_snplist <- target_qc_snplist[target_qc_snplist %in% targ_pvar$SNP]

# Perform LD pruning
ld_indep <- plink_prune(bfile = opt$target_plink_chr, chr = CHROMS, plink2 = opt$plink2, extract = target_qc_snplist, threads = opt$n_cores)

# To improve efficiency, derive PCs using random subset of 1000 individuals.
fam<-fread('~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/synth_2_chr-22.fam')
fam_subset <- fam[sample(1000, replace = F),]

# Run PCA
snp_weights <- plink_pca(bfile = opt$target_plink_chr, keep = fam_subset, chr = CHROMS, plink2 = opt$plink2, extract = ld_indep, n_pc = opt$n_pcs, threads = opt$n_cores)
fwrite(snp_weights, paste0(tmp_dir,'/ref.eigenvec.var'), row.names = F, quote=F, sep=' ', na='NA')

# Project into the full population
target_pcs <- plink_score(bfile = opt$target_plink_chr, chr = CHROMS, plink2 = opt$plink2, score = paste0(tmp_dir,'/ref.eigenvec.var'), threads = opt$n_cores)

dir.create('~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/pca')
fwrite(target_pcs, paste0(opt$output,'pcs.txt'), quote=F, sep=' ', na='NA')

```

</details>

***

#### Perform GWAS

<details><summary>Show code</summary>

```{bash}

module add plink2
for pheno in $(seq 1 2); do
    mkdir -p ~/oliverpainfel/Analysis/HAPNEST/synth_2/gwas/pheno${pheno}
    for chr in $(seq 22 22); do
        sbatch -p neurohack_cpu --mem 20G -n 4 --wrap="plink2 \
          --bfile ~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/synth_2_chr-${chr} \
          --pheno ~/oliverpainfel/Data/HAPNEST/synth_2/trait_${pheno}.txt \
          --1 \
          --covar ~/oliverpainfel/Data/HAPNEST/data/outputs/synth_2/pca/pcs.txt \
          --covar-variance-standardize \
          --linear omit-ref cols=+a1freq,+ax hide-covar \
          --maf 0.01 \
          --geno 0.05 \
          --out ~/oliverpainfel/Analysis/HAPNEST/synth_2/gwas/pheno${pheno}/pheno${pheno}.chr${chr}"
    done
done

```

```{r}
library(data.table)
  rsids<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/rsid_map_list_chr22.txt')

for(i in 2:2){
  ss <- fread(paste0('/users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/gwas/pheno', i, '/pheno', i,'.chr22.outcome.glm.linear'))
  ss$REF<-NULL
  ss$ALT<-NULL
  names(ss)[names(ss) == 'AX'] <- 'A2'
  
  ss<-merge(ss, rsids, by.x = 'ID', by.y = 'id_hg38')
  ss$ID <- ss$rsid
  ss$rsid <- NULL
  
  fwrite(ss, paste0('/users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/gwas/pheno', i, '/pheno', i,'.chr22.outcome.glm.linear'), row.names = F, quote=F, sep=' ', na='NA')
}
```


</details>

***

#### Run MAGMA

<details><summary>Show code</summary>

```{bash}

# Gene association
mkdir /users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/magma

for pheno in $(seq 2 2); do
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval /users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/gwas/pheno${pheno}/pheno${pheno}.chr22.outcome.glm.linear use=ID,P ncol=OBS_CT \
  --gene-model snp-wise=top \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out /users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/magma/pheno${pheno}_gene_level
done

# GSEA
for pheno in $(seq 2 2); do
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results /users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/magma/pheno${pheno}_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c5.go.bp.v7.5.1.entrez.gmt \
  --out /users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/magma/pheno${pheno}_gsea
done

```

```{r}
# Investigate gene level associations
# Why are the gene sets not showing as enriched
res<-fread('/users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/magma/pheno2_gene_level.genes.out')

# Read in .gmt file
gmt <- readLines('/scratch/prj/oliverpainfel/recovered/Data/MSigDB/c5.go.bp.v7.5.1.entrez.gmt')
gene_sets <- strsplit(gmt, "\t")
gene_list <- setNames(lapply(gene_sets, `[`, -c(1,2)), sapply(gene_sets, `[`, 1))

# Subset sets of interest
gene_list <- gene_list[ c('GOBP_DNA_REPAIR', 'GOBP_REGULATION_OF_SYNAPTIC_PLASTICITY')]

res[res$GENE %in% gene_list[['GOBP_DNA_REPAIR']],]
res[res$GENE %in% gene_list[['GOBP_REGULATION_OF_SYNAPTIC_PLASTICITY']],]

# Read in causal variants
causal_var<-fread('~/oliverpainfel/Data/HAPNEST/synth_2/causal.snplist')
rsids<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/rsid_map_list_chr22.txt')
causal_var<-merge(causal_var, rsids, by.x='V1', by.y='rsid')
causal_var<-causal_var[order(causal_var$id_hg38),]

# Many of the genes are not coming up as significant, despite containing SNPs with crazy high P values. Try running MAGMA with top,1 gene model. This could pick up the signal better. This helps, but gene p values are capped at 1e-50, and many genes are still non-significant. Manually check top SNP p values within these genes.
# Read in SNPs for gene 9463
gwas <- fread(paste0('/users/k1806347/oliverpainfel/Analysis/HAPNEST/synth_2/gwas/pheno2/pheno2.chr22.outcome.glm.linear'))
mean(abs(gwas[gwas$POS > 38418262 & gwas$POS < 38481708,]$T_STAT))
# The gene is highly enriched. Why is MAGMA not finding this.
# Make sure the SNPs we are setting as causal are in the MAGMA analysis/annot file
```

</details>

***

## Artificial misdiagnosis

Look at the UKB independent sumstats that I have used for PGS comparison. Check PGS R2 and gene set enrichment.

T2D and IBD both have strong PGS R2, and presumably different gene set enrichment.

***

### Run GSEA

```{bash}

mkdir -p ~/oliverpainfel/Analyses/GenoClust/artificial/gwas
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST004773/GCST004773-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T2D.gz
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST004131/GCST004131-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/IBD.gz
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST90013445/GCST90013445-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T1D.gz
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST90013534/GCST90013534-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA.gz
cp /scratch/prj/oliverpainfel/recovered/Data/GWAS_sumstats/prs_comparison/cleaned/COAD01.cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD.gz

/scratch/prj/oliverpainfel/recovered/Data/GWAS_sumstats/prs_comparison/cleaned/

gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T2D.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/IBD.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T1D.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD.gz

mkdir -p ~/oliverpainfel/Analyses/GenoClust/artificial/magma/

# Gene association
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T2D use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/IBD use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T1D use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gene_level

# GSEA
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gsea
# Surprisingly T2D has very few significantly enriched gene sets

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gsea
  
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gsea
  
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gsea

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gsea
  
```

```{r}
# Compare GSEA results across traits
T2D<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gsea.gsa.out', skip = 3)
IBD<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gsea.gsa.out', skip = 3)
T1D<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gsea.gsa.out', skip = 3)
RA<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gsea.gsa.out', skip = 3)
CAD<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gsea.gsa.out', skip = 3)

T2D <- T2D[, c('FULL_NAME','BETA','P'), with=F]
T2D$TRAIT <- 'T2D'
IBD <- IBD[, c('FULL_NAME','BETA','P'), with=F]
IBD$TRAIT <- 'IBD'
T1D <- T1D[, c('FULL_NAME','BETA','P'), with=F]
T1D$TRAIT <- 'T1D'
RA <- RA[, c('FULL_NAME','BETA','P'), with=F]
RA$TRAIT <- 'RA'
CAD <- CAD[, c('FULL_NAME','BETA','P'), with=F]
CAD$TRAIT <- 'CAD'

both <- rbind(T2D, IBD, T1D, RA, CAD)

both_wide <- reshape(both, 
                     idvar = "FULL_NAME", 
                     timevar = "TRAIT", 
                     direction = "wide")

cor(both_wide[, grepl('BETA', names(both_wide)), with = F], use='p')

both_wide$FDR.T2D <- p.adjust(both_wide$P.T2D, method = 'fdr')
both_wide$FDR.IBD <- p.adjust(both_wide$P.IBD, method = 'fdr')
both_wide$FDR.T1D <- p.adjust(both_wide$P.T1D, method = 'fdr')
both_wide$FDR.RA <- p.adjust(both_wide$P.RA, method = 'fdr')
both_wide$FDR.CAD <- p.adjust(both_wide$P.CAD, method = 'fdr')

library(VennDiagram)

venn.plot <- venn.diagram(
  x = list(T2D = both_wide$FULL_NAME[both_wide$FDR.T2D < 0.05], IBD = both_wide$FULL_NAME[both_wide$FDR.IBD < 0.05]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(T1D = both_wide$FULL_NAME[which(both_wide$FDR.T1D < 0.05)], IBD = both_wide$FULL_NAME[which(both_wide$FDR.IBD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[which(both_wide$FDR.RA < 0.05)], IBD = both_wide$FULL_NAME[which(both_wide$FDR.IBD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[which(both_wide$FDR.RA < 0.05)], T1D = both_wide$FULL_NAME[which(both_wide$FDR.T1D < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[which(both_wide$FDR.RA < 0.05)], CAD = both_wide$FULL_NAME[which(both_wide$FDR.CAD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(IBD = both_wide$FULL_NAME[which(both_wide$FDR.IBD < 0.05)], CAD = both_wide$FULL_NAME[which(both_wide$FDR.CAD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

```
***

### Meta-analyse GWAS

Lets use RA and CAD as they have a low correlation, but decent number of significant sets.

```{bash, eval=F, echo=T}

# Run METAL
/scratch/prj/oliverpainfel/recovered/Software/generic-metal/metal

SCHEME SAMPLESIZE
MARKER SNP
WEIGHT N
ALLELE A1 A2
EFFECT BETA
STDERR SE
PVALUE P
PROCESS /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD
PROCESS /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA
OUTFILE /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA .tbl

ANALYZE

QUIT

```

```{r, eval=F, echo=T}
# Format the METAL output
library(data.table)

# Read in the sumstats
meta<-fread('/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA1.tbl')

# Remove SNPs that were not present in both studies
meta <- meta[!grepl('\\?', meta$Direction),]

# Format for GenoPred
names(meta)<-c('SNP','A1','A2','N','BETA','P','Direction')
meta$A1<-toupper(meta$A1)
meta$A2<-toupper(meta$A2)
meta$SE<-1
meta<-meta[,c('SNP','A1','A2','BETA','SE','N','P'),with=F]

fwrite(meta, '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.tbl.reformat', quote=F, sep=' ', na='NA')

```

```{bash}
# Rerun MAGMA to see how the results compare to the original GWAS
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.tbl.reformat use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gsea
  
```

```{r}
# Compare GSEA results across traits
RA<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gsea.gsa.out', skip = 3)
CAD<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gsea.gsa.out', skip = 3)
MIX<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gsea.gsa.out', skip = 3)

RA <- RA[, c('FULL_NAME','BETA','P'), with=F]
RA$TRAIT <- 'RA'
CAD <- CAD[, c('FULL_NAME','BETA','P'), with=F]
CAD$TRAIT <- 'CAD'
MIX <- MIX[, c('FULL_NAME','BETA','P'), with=F]
MIX$TRAIT <- 'MIX'

both <- rbind(RA, CAD, MIX)

both_wide <- reshape(both, 
                     idvar = "FULL_NAME", 
                     timevar = "TRAIT", 
                     direction = "wide")

cor(both_wide[, grepl('BETA', names(both_wide)), with = F], use='p')

both_wide$FDR.RA <- p.adjust(both_wide$P.RA, method = 'fdr')
both_wide$FDR.CAD <- p.adjust(both_wide$P.CAD, method = 'fdr')
both_wide$FDR.MIX <- p.adjust(both_wide$P.MIX, method = 'fdr')

library(VennDiagram)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[both_wide$FDR.RA < 0.05], 
           CAD = both_wide$FULL_NAME[both_wide$FDR.CAD < 0.05],
           MIX = both_wide$FULL_NAME[both_wide$FDR.MIX < 0.05]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

# It is still picking up some pathways for both disorders, but also many new pathways.
both_wide[both_wide$FDR.MIX < 0.05 & both_wide$FDR.RA > 0.05 & both_wide$FDR.CAD > 0.05,]
```

***

### Create phenotype data in UKB

```{r}
#####
# CAD - Use definition from https://doi.org/10.1001/jama.2019.22241
#####

library(ukbkings)
library(dplyr)
library(stringr)

# Extract ICD code and sex from UKB
project_dir <- "/datasets/ukbiobank/ukb82087"
f <- bio_field(project_dir)
f %>%
    select(field, name) %>%
    filter(str_detect(field, "^31-0.0|41270|41271|41272|20002|20004|6150")) %>%
    bio_field_add("~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset.txt")

bio_phen(
    project_dir,
    field = "~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset.txt",
    out = "~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset"
)

df <- readRDS("~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset.rds")

# Load required libraries
library(dplyr)

# Assume 'ukb_data' is your loaded data frame
# ukb_data <- read.csv("your_data_file.csv") # Example for loading data

# CAD code definitions
cad_icd10 <- c("I21", "I22", "I23", "I241", "I252")
cad_icd9 <- c("410", "411", "412", "42789")
cad_opcs4 <- paste0("K", c("401", "402", "403", "404",
                           "411", "412", "413", "414",
                           "451", "452", "453", "454", "455"))
cad_self_reported <- c("1075")
cad_operation <- c("1070", "1095")
cad_vascular <- c("1")

# Columns to search in
icd10_cols <- grep("^41270", names(df), value = TRUE) # ICD-10 columns
icd9_cols <- grep("^41271", names(df), value = TRUE) # ICD-9 columns
opcs4_cols <- grep("^41272", names(df), value = TRUE) # OPCS-4 columns
self_report_cols <- grep("^20002", names(df), value = TRUE) # Self-reported illness
operation_cols <- grep("^20004", names(df), value = TRUE) # Operation codes
vascular_cols <- grep("^6150", names(df), value = TRUE) # Vascular/heart problems

# Create a binary CAD phenotype
df <- df %>%
  mutate(
    CAD = if_else(
      rowSums(across(all_of(icd10_cols), ~ . %in% cad_icd10), na.rm = TRUE) > 0 |
      rowSums(across(all_of(icd9_cols), ~ . %in% cad_icd9), na.rm = TRUE) > 0 |
      rowSums(across(all_of(opcs4_cols), ~ . %in% cad_opcs4), na.rm = TRUE) > 0 |
      rowSums(across(all_of(self_report_cols), ~ . %in% cad_self_reported), na.rm = TRUE) > 0 |
      rowSums(across(all_of(operation_cols), ~ . %in% cad_operation), na.rm = TRUE) > 0 |
      rowSums(across(all_of(vascular_cols), ~ . %in% cad_vascular), na.rm = TRUE) > 0,
      1, 0
    )
  )

# Summary of CAD cases
table(df$CAD)

cad <- data.frame(eid = df$eid, CAD = df$CAD)

#####
# RA
#####

keep_files<-list.files(path = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/output/ukb/pcs/within_sample/', pattern = '.keep')

pop_dat<-NULL
for(i in keep_files){
  tmp<-fread(paste0('/users/k1806347/oliverpainfel/Data/ukb/GenoPred/output/ukb/pcs/within_sample/', i))
  names(tmp)<-c('FID','IID')
  tmp$POP<-gsub('.keep','', gsub('ukb.outlier_detection.','',i))
  pop_dat<-rbind(pop_dat, tmp)
}

# Update row number IDs to project specific IDs
psam<-fread('/scratch/prj/ukbiobank/recovered/ukb82087/imputed/ukb82087_imp_chr1_MAF1_INFO4_v1.psam')
psam$rn<-1:nrow(psam)
psam<-psam[,c('IID','rn'), with = F]

pop_dat$FID<-NULL
pop_dat<-merge(pop_dat, psam, by.x='IID', by.y='rn')
pop_dat<-data.frame(
  eid=pop_dat$IID.y,
  POP=pop_dat$POP
)

# This has been defined already during our UKB benchmark
ra <- fread('/scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/phenotypes/ra.unrel.txt')
names(ra)<-c('eid','RA')

both <- merge(ra, cad, by='eid')
both$MIX <- ifelse(both$RA == 1 | both$CAD == 1, 1, 0)
nrow(both[both$RA == 1 & both$CAD == 1,])

# Restrict analysis to EUR individuals
both<-both[both$eid %in% pop_dat$eid[pop_dat$POP == 'EUR']]

# There are only 954  RA cases but >22k CAD cases. This could make it hard for clustering to pick out this cluster - Lets see.
# These are unrelated individuals in UKB already.
# To speed things up, lets use a subset of 25k controls
set.seed(1)
controls<-both$eid[both$MIX == 0]
keep<-sample(controls, 25000)
both<-both[!(both$eid %in% controls[!(controls %in% keep)])]

write.table(both, '~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.txt', row.names = F, col.names = T, quote = F)

# Save with row.number IDs
both<-merge(both, psam, by.x='eid', by.y='IID')

for(i in c('RA','CAD','MIX')){
  pheno<-data.frame(
    FID=both$rn,
    IID=both$rn,
    outcome=both[[i]]
  )
  write.table(pheno, paste0('~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.', i, '_only.txt'), row.names = F, col.names = T, quote = F)
}

keep<-data.frame(FID = both$rn, IID = both$rn)

write.table(keep, '~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.keep', row.names = F, col.names = F, quote = F)

```

***

### Run GenoPred

```{bash}
mkdir ~/oliverpainfel/Data/ukb/genoclust_subset

for chr in $(seq 1 22); do
  ~/oliverpainfel/Software/plink2 \
    --pfile ~/oliverpainfel/Data/ukb/GenoPred/output/ukb/geno/ukb.ref.chr${chr} \
    --keep ~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.keep \
    --make-pgen \
    --out ~/oliverpainfel/Data/ukb/genoclust_subset/ukb.chr${chr}
done

```

```{r}
library(data.table)

dir.create('/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset')

######
# target_list
######

target_list <- data.frame(
  name='ukb',
  path='/users/k1806347/oliverpainfel/Data/ukb/genoclust_subset/ukb',
  type='plink2',
  indiv_report=F,
  unrel='/users/k1806347/oliverpainfel/Data/ukb/phenotypes/unrelated.row_number.txt'
)

write.table(target_list, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/target_list.txt', col.names=T, row.names=F, quote=F)

######
# gwas_list
######

gwas_list<-data.frame(
  name=c('CAD','RA','MIX','T2D','T1D', 'IBD','BMI', 'HbA1c'),
  path=c(
    '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD',
    '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA',
    '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.tbl.reformat',
    '/scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST004773/GCST004773-cleaned.gz',
    '/scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST90013445/GCST90013445-cleaned.gz',
    '/scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST004131/GCST004131-cleaned.gz',
    '/scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST002783/GCST002783-cleaned.gz',
    '/scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST007954/GCST007954-cleaned.gz'
  ),
  population='EUR',
  n=NA,
  sampling=NA,
  prevalence=NA,
  mean=NA,
  sd=NA,
  label=paste0('"', c('CAD','RA','MIX','T2D','T1D', 'IBD','BMI', 'HbA1c'),'"')
)

write.table(gwas_list, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/gwas_list.txt', col.names=T, row.names=F, quote=F)

######
# config
######

config<-c(
  "outdir: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/output_genoclust",
  "config_file: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml",
  "gwas_list: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/gwas_list.txt",
  "target_list: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/target_list.txt",
  "pgs_methods: ['sbayesrc']",
  "cores_prep_pgs: 10",
  "cores_target_pgs: 10",
  "sbayesrc_ldref: /users/k1806347/oliverpainfel/Data/hgdp_1kg/sbayesrc/hm3",
  "gene_sets: /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt"
)

write.table(config, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', col.names = F, row.names = F, quote = F)

```

```{bash}
cd /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline
conda activate genopred

snakemake \
  --profile slurm \
  --use-conda \
  --configfile=/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml \
  output_all target_pgs_partitioned -n
```

***

### Genome-wide PGS

#### Check correlation

Check correlation between the PGS for CAD, RA and MIX with CAD, RA and MIX outcomes.

```{r}
setwd('/users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline')
library(data.table)
library(ggplot2)
library(cowplot)

source('../functions/misc.R')
source_all('../functions')

# Define pgs_methods used
pgs_methods <- read_param(config = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', param = 'pgs_methods', return_obj = F)

# Read in PGS
pgs <- read_pgs(config = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', name = 'ukb')$ukb
pgs <- Reduce(function(x, y) merge(x, y, by = c("FID", "IID"), all = TRUE), lapply(pgs$TRANS, function(x) x[[1]]))

# Read in outcome data
outcome_names<-c('CAD','RA','MIX')
outcomes <- list()
for(i in outcome_names){
  outcomes[[i]]<-fread(paste0('~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.', i, '_only.txt'))
  names(outcomes[[i]])[names(outcomes[[i]]) == 'outcome']<-paste0(i,'_outcome')
}
outcomes <- Reduce(function(x, y) merge(x, y, by = c("FID", "IID"), all = TRUE), outcomes)

# Test correlation
both <- merge(pgs, outcomes, by = c('FID', 'IID'))

round(cor(both[,-1:-2]), 2)

#              CAD_SBayesRC RA_SBayesRC MIX_SBayesRC CAD_outcome RA_outcome MIX_outcome
# CAD_SBayesRC         1.00       -0.01         0.24        0.20      -0.02        0.20
# RA_SBayesRC         -0.01        1.00         0.87        0.00       0.12        0.03
# MIX_SBayesRC         0.24        0.87         1.00        0.06       0.11        0.09
# CAD_outcome          0.20        0.00         0.06        1.00      -0.10        0.97
# RA_outcome          -0.02        0.12         0.11       -0.10       1.00        0.15
# MIX_outcome          0.20        0.03         0.09        0.97       0.15        1.00

# This looks good so far:
# - The MIX PGS is more correlated with the RA PGS than CAD PGS, due to larger genetic effects on RA
# - The MIX outcome is more correlated with the CAD outcome, due to large number of CAD cases
# - The MIX PGS predicts CAD and RA outcome worse that CAD and RA PGS respectively

# The reason for PGS stratification is becoming clearer - If we were to extract genetic effects relevant to CAD, then the CAD PGS would predict CAD cases better. Likewise for RA.

# It would be interesting to see whether we can cluster individuals based on these scores alone. This would be similar to the T2D paper, which stratifies PGS by relationship with other traits. Our pathway specific strategy will take a different approach.

library(pROC)

auc(
  both$MIX_outcome,
  predict(glm(MIX_outcome ~ RA_SBayesRC + CAD_SBayesRC + MIX_SBayesRC, data = both), newdata = both))

auc(
  both$MIX_outcome,
  predict(glm(MIX_outcome ~ MIX_SBayesRC, data = both), newdata = both))

auc(
  both$MIX_outcome,
  predict(glm(MIX_outcome ~ RA_SBayesRC + CAD_SBayesRC, data = both), newdata = both))

```

***

#### Cluster

```{r}

library(NbClust)

both$group<-NULL
both$group[both$CAD_outcome == 1]<-'CAD'
both$group[both$RA_outcome == 1]<-'RA'
both$group[both$CAD_outcome == 1 & both$RA_outcome == 1]<-'BOTH'
both$group[both$CAD_outcome != 1 & both$RA_outcome != 1]<-'CONTROL'

# Extract PGS data for MIX cases only
both_cases <- both[which(both$MIX_outcome == 1),]
pgs_cases <- both_cases[, c('CAD_SBayesRC', 'RA_SBayesRC'), with=F]
pgs_cases <- scale(pgs_cases) 
# Not sure whether I should be scaling since they are reference standardised

# Determine optimal number of clusters
n_clust_sol <-
  NbClust(
    data = pgs_cases[1:2000,],
    distance = "euclidean",
    min.nc = 2,
    max.nc = 10,
    method = 'ward.D2',
    index = 'all'
  )

n_clust_opt<-length(unique(n_clust_sol$Best.partition))

#####
# K-means
#####

# Now try k-means clustering with 3 clusters
k_res<-kmeans(pgs_cases, n_clust_opt)

# Plot the mean of each group
k_res_centers<-data.frame(Group=as.character(1:n_clust_opt),
                          k_res$centers)

library(reshape2)
k_res_centers_melt<-melt(k_res_centers, id='Group')

ggplot(k_res_centers_melt, aes(x=variable, y=value, group=Group, color=Group)) +
  geom_point(size=5) +
  geom_line() +
  labs(x='Polygenic Score', y='Cluster Mean', title='Mean Polygenic Score Across Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Hierarchical clustering
#####

# Compute distance matrix
dist_matrix <- dist(pgs_cases, method = "euclidean")

# Perform hierarchical clustering
hclust_res <- hclust(dist_matrix, method = "ward.D2")

# Cut tree into n_clust_opt clusters
hclust_clusters <- cutree(hclust_res, k = n_clust_opt)

# Convert clusters to a factor for plotting
hclust_clusters <- as.factor(hclust_clusters)

# Create a data frame with cluster assignments
pgs_cases_clustered <- data.frame(pgs_cases, Cluster = hclust_clusters)

# Compute the mean of each cluster for each PGS
hclust_centers <- aggregate(. ~ Cluster, data = pgs_cases_clustered, FUN = mean)

# Reshape the data for plotting
hclust_centers_melt <- melt(hclust_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
hclust_centers_melt$Cluster <- as.character(hclust_centers_melt$Cluster)

ggplot(hclust_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across Hierarchical Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Gaussian mixture model
#####
library(mclust)

gmm_model <- Mclust(pgs_cases, G = n_clust_opt)
summary(gmm_model)

gmm_clusters <- gmm_model$classification  # Hard cluster labels
gmm_probabilities <- gmm_model$z  # Soft probabilities for each cluster

# Create a data frame with cluster assignments
pgs_cases_clustered <- data.frame(pgs_cases, Cluster = gmm_clusters)

# Compute the mean of each cluster for each PGS
gmm_centers <- aggregate(. ~ Cluster, data = pgs_cases_clustered, FUN = mean)

# Reshape the data for plotting
gmm_centers_melt <- melt(gmm_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
gmm_centers_melt$Cluster <- as.character(gmm_centers_melt$Cluster)

ggplot(gmm_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across GMM Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#########
# Compare the clusters to the case groups
k_means_conf_matrix <- table(k_res$cluster, both_cases$group)
hclust_D2_conf_matrix <- table(hclust_clusters, both_cases$group)
gmm_conf_matrix <- table(gmm_clusters, both_cases$group)

library(mclust)
adjustedRandIndex(k_res$cluster, both_cases$group)
adjustedRandIndex(hclust_clusters, both_cases$group)
adjustedRandIndex(gmm_clusters, both_cases$group)

# The accuracy of the hclust solution is higher
pca_res <- prcomp(pgs_cases)
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(k_res$cluster))) + 
  geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(hclust_clusters))) + 
  geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(gmm_clusters))) + 
  geom_point()

ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(k_res$cluster))) + 
  geom_point()
ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(hclust_clusters))) + 
  geom_point()
ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(gmm_clusters))) + 
  geom_point()

ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(group))) + 
  geom_point()

ggplot(both_cases[both_cases$group != 'BOTH',], aes(x = RA_SBayesRC, y = CAD_SBayesRC, color = as.factor(group))) +
  geom_density_2d() +  # Density contours
  theme_minimal()

library(Rtsne)

tsne_res <- Rtsne(pgs_cases[!duplicated(pgs_cases),])

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(k_res$cluster[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(hclust_clusters[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(gmm_clusters[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')


########
# Predict MIX outcome 

# Project cluster solutions into full dataset
both_cluster<-both

library(clue)
# kmeans
kmeans_pred<-cl_predict(k_res, newdata = both, type = "memberships")
both_cluster$kmean<-apply(kmeans_pred, 1, function(x) which(x == 1))

# hclust
centroids <- aggregate(pgs_cases, list(hclust_clusters), mean)[, -1]  # Remove the cluster ID column
assign_to_nearest_cluster <- function(new_data, centroids) {
  apply(as.matrix(new_data), 1, function(row) {
    which.min(colSums((t(centroids) - row)^2))  # Compute Euclidean distance to centroids
  })
}
both_cluster$hclust <-
  assign_to_nearest_cluster(
    new_data = both[, c('CAD_SBayesRC', 'RA_SBayesRC'), with = F], centroids = centroids)

# gmm
both_cluster$gmm <-
  predict(gmm_model, 
          newdata = both[, c('CAD_SBayesRC', 'RA_SBayesRC'), with = F])$classification
both_cluster$gmm_prob <-
  predict(gmm_model, 
          newdata = both[, c('CAD_SBayesRC', 'RA_SBayesRC'), with = F])$z[,1]

# MIX_SBayesRC only
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC")),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# MIX_SBayesRC + disorder specific PGS
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC + CAD_SBayesRC + RA_SBayesRC")),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# kmeans interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(kmean)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# hclust interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(hclust)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# gmm interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(gmm)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# Even when defining clusters in the cases, the improvement in prediction is minimal < 1% relative improvement. However, there is a significant interaction between the RA PGS and the cluster that represents RA cases, so the concept is holding to some extent.

# Would pathway specific scores help? The best PGS predicting RA or CAD subgroups would be RA or CAD PGS, so pathway specific PGS are unlikely to improve clustering/prediction. This isn't working well because the PGS are weak predictors...

# Including the disease specific PGS improves prediction a lot over the MIX PGS alone. I think this is hoovering up a lot of the interaction effects. 

# Rerun using full sample to define clusters

#####################

library(NbClust)

both$group<-NULL
both$group[both$CAD_outcome == 1]<-'CAD'
both$group[both$RA_outcome == 1]<-'RA'
both$group[both$CAD_outcome == 1 & both$RA_outcome == 1]<-'BOTH'
both$group[both$CAD_outcome != 1 & both$RA_outcome != 1]<-'CONTROL'

# Extract PGS data for MIX cases only
pgs_all <- both[, c('CAD_SBayesRC', 'RA_SBayesRC'), with=F]
pgs_all <- scale(pgs_all) 
# Not sure whether I should be scaling since they are reference standardised

# Determine optimal number of clusters
n_clust_sol <-
  NbClust(
    data = pgs_all[1:2000,],
    distance = "euclidean",
    min.nc = 2,
    max.nc = 10,
    method = 'ward.D2',
    index = 'all'
  )

n_clust_opt<-length(unique(n_clust_sol$Best.partition))

#####
# K-means
#####

# Now try k-means clustering with 3 clusters
k_res<-kmeans(pgs_all, n_clust_opt)

# Plot the mean of each group
k_res_centers<-data.frame(Group=as.character(1:n_clust_opt),
                          k_res$centers)

library(reshape2)
k_res_centers_melt<-melt(k_res_centers, id='Group')

ggplot(k_res_centers_melt, aes(x=variable, y=value, group=Group, color=Group)) +
  geom_point(size=5) +
  geom_line() +
  labs(x='Polygenic Score', y='Cluster Mean', title='Mean Polygenic Score Across Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Hierarchical clustering
#####

# Compute distance matrix
dist_matrix <- dist(pgs_all, method = "euclidean")

# Perform hierarchical clustering
hclust_res <- hclust(dist_matrix, method = "ward.D2")

# Cut tree into n_clust_opt clusters
hclust_clusters <- cutree(hclust_res, k = n_clust_opt)

# Convert clusters to a factor for plotting
hclust_clusters <- as.factor(hclust_clusters)

# Create a data frame with cluster assignments
pgs_all_clustered <- data.frame(pgs_all, Cluster = hclust_clusters)

# Compute the mean of each cluster for each PGS
hclust_centers <- aggregate(. ~ Cluster, data = pgs_all_clustered, FUN = mean)

# Reshape the data for plotting
hclust_centers_melt <- melt(hclust_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
hclust_centers_melt$Cluster <- as.character(hclust_centers_melt$Cluster)

ggplot(hclust_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across Hierarchical Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Gaussian mixture model
#####
library(mclust)

gmm_model <- Mclust(pgs_all, G = n_clust_opt)
summary(gmm_model)

gmm_clusters <- gmm_model$classification  # Hard cluster labels
gmm_probabilities <- gmm_model$z  # Soft probabilities for each cluster

# Create a data frame with cluster assignments
pgs_all_clustered <- data.frame(pgs_all, Cluster = gmm_clusters)

# Compute the mean of each cluster for each PGS
gmm_centers <- aggregate(. ~ Cluster, data = pgs_all_clustered, FUN = mean)

# Reshape the data for plotting
gmm_centers_melt <- melt(gmm_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
gmm_centers_melt$Cluster <- as.character(gmm_centers_melt$Cluster)

ggplot(gmm_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across GMM Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#########
# Compare the clusters to the case groups
k_means_conf_matrix <- table(k_res$cluster, both$group)
hclust_D2_conf_matrix <- table(hclust_clusters, both$group)
gmm_conf_matrix <- table(assigned_clusters, both$group)

library(mclust)
adjustedRandIndex(k_res$cluster, both$group)
adjustedRandIndex(hclust_clusters, both$group)
adjustedRandIndex(gmm_clusters, both$group)

# The accuracy of the hclust solution is higher
pca_res <- prcomp(pgs_all)
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(k_res$cluster))) + geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(hclust_clusters))) + geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(gmm_clusters))) + geom_point()

########
# Predict MIX outcome 

# Project cluster solutions into full dataset
both_cluster<-both

both_cluster$kmean<-k_res$cluster
both_cluster$hclust<-hclust_clusters
both_cluster$gmm<-gmm_clusters

# MIX_SBayesRC only
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC")),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# MIX_SBayesRC + disorder specific PGS
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC + CAD_SBayesRC + RA_SBayesRC")),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# No interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], collapse = ' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# kmeans interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(kmean)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# hclust interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(hclust)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# gmm interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(gmm)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# Full sample show slightly better prediction, but there is poorer seperate of case subtypes.

```

***

Try adding in PGS for secondary traits.

```{r}
library(NbClust)

both$group<-NULL
both$group[both$CAD_outcome == 1]<-'CAD'
both$group[both$RA_outcome == 1]<-'RA'
both$group[both$CAD_outcome == 1 & both$RA_outcome == 1]<-'BOTH'
both$group[both$CAD_outcome != 1 & both$RA_outcome != 1]<-'CONTROL'

# Extract PGS data for MIX cases only
both_cases <- both[which(both$MIX_outcome == 1),]
pgs_cases <- both_cases[, grepl('SBayesRC', names(both_cases)), with=F]
pgs_cases <- scale(pgs_cases) 
# Not sure whether I should be scaling since they are reference standardised

# Determine optimal number of clusters
n_clust_sol <-
  NbClust(
    data = pgs_cases[1:2000,],
    distance = "euclidean",
    min.nc = 2,
    max.nc = 10,
    method = 'ward.D2',
    index = 'all'
  )

n_clust_opt<-length(unique(n_clust_sol$Best.partition))

#####
# K-means
#####

# Now try k-means clustering with 3 clusters
k_res<-kmeans(pgs_cases, n_clust_opt)

# Plot the mean of each group
k_res_centers<-data.frame(Group=as.character(1:n_clust_opt),
                          k_res$centers)

library(reshape2)
k_res_centers_melt<-melt(k_res_centers, id='Group')

ggplot(k_res_centers_melt, aes(x=variable, y=value, group=Group, color=Group)) +
  geom_point(size=5) +
  geom_line() +
  labs(x='Polygenic Score', y='Cluster Mean', title='Mean Polygenic Score Across Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Hierarchical clustering
#####

# Compute distance matrix
dist_matrix <- dist(pgs_cases, method = "euclidean")

# Perform hierarchical clustering
hclust_res <- hclust(dist_matrix, method = "ward.D2")

# Cut tree into n_clust_opt clusters
hclust_clusters <- cutree(hclust_res, k = n_clust_opt)

# Convert clusters to a factor for plotting
hclust_clusters <- as.factor(hclust_clusters)

# Create a data frame with cluster assignments
pgs_cases_clustered <- data.frame(pgs_cases, Cluster = hclust_clusters)

# Compute the mean of each cluster for each PGS
hclust_centers <- aggregate(. ~ Cluster, data = pgs_cases_clustered, FUN = mean)

# Reshape the data for plotting
hclust_centers_melt <- melt(hclust_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
hclust_centers_melt$Cluster <- as.character(hclust_centers_melt$Cluster)

ggplot(hclust_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across Hierarchical Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Gaussian mixture model
#####
library(mclust)

gmm_model <- Mclust(pgs_cases, G = n_clust_opt)
summary(gmm_model)

gmm_clusters <- gmm_model$classification  # Hard cluster labels
gmm_probabilities <- gmm_model$z  # Soft probabilities for each cluster

# Create a data frame with cluster assignments
pgs_cases_clustered <- data.frame(pgs_cases, Cluster = gmm_clusters)

# Compute the mean of each cluster for each PGS
gmm_centers <- aggregate(. ~ Cluster, data = pgs_cases_clustered, FUN = mean)

# Reshape the data for plotting
gmm_centers_melt <- melt(gmm_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
gmm_centers_melt$Cluster <- as.character(gmm_centers_melt$Cluster)

ggplot(gmm_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across GMM Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#########
# Compare the clusters to the case groups
k_means_conf_matrix <- table(k_res$cluster, both_cases$group)
hclust_D2_conf_matrix <- table(hclust_clusters, both_cases$group)
gmm_conf_matrix <- table(gmm_clusters, both_cases$group)

library(mclust)
adjustedRandIndex(k_res$cluster, both_cases$group)
adjustedRandIndex(hclust_clusters, both_cases$group)
adjustedRandIndex(gmm_clusters, both_cases$group)

# The accuracy of the hclust solution is higher
pca_res <- prcomp(pgs_cases)
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(k_res$cluster))) + 
  geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(hclust_clusters))) + 
  geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(gmm_clusters))) + 
  geom_point()

ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(k_res$cluster))) + 
  geom_point()
ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(hclust_clusters))) + 
  geom_point()
ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(gmm_clusters))) + 
  geom_point()

ggplot(both_cases, aes(RA_SBayesRC, CAD_SBayesRC, color = as.factor(group))) + 
  geom_point()

ggplot(both_cases[both_cases$group != 'BOTH',], aes(x = RA_SBayesRC, y = CAD_SBayesRC, color = as.factor(group))) +
  geom_density_2d() +  # Density contours
  theme_minimal()

library(Rtsne)

tsne_res <- Rtsne(pgs_cases[!duplicated(pgs_cases),])

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(k_res$cluster[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(hclust_clusters[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(gmm_clusters[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')


########
# Predict MIX outcome 

# Project cluster solutions into full dataset
both_cluster<-both

library(clue)
# kmeans
kmeans_pred<-cl_predict(k_res, newdata = both, type = "memberships")
both_cluster$kmean<-apply(kmeans_pred, 1, function(x) which(x == 1))

# hclust
centroids <- aggregate(pgs_cases, list(hclust_clusters), mean)[, -1]  # Remove the cluster ID column
assign_to_nearest_cluster <- function(new_data, centroids) {
  apply(as.matrix(new_data), 1, function(row) {
    which.min(colSums((t(centroids) - row)^2))  # Compute Euclidean distance to centroids
  })
}
both_cluster$hclust <-
  assign_to_nearest_cluster(
    new_data = both[, grepl('SBayesRC', names(both_cases)), with = F], centroids = centroids)

# gmm
both_cluster$gmm <-
  predict(gmm_model, 
          newdata = both[, grepl('SBayesRC', names(both_cases)), with = F])$classification
both_cluster$gmm_prob <-
  predict(gmm_model, 
          newdata = both[, grepl('SBayesRC', names(both_cases)), with = F])$z[,1]


# MIX_SBayesRC only
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC")),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# MIX_SBayesRC + disorder specific PGS
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC + CAD_SBayesRC + RA_SBayesRC")),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# No interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], collapse = ' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# kmeans interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(kmean)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# hclust interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(hclust)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# gmm interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  ", 
                      paste0(names(both_cluster)[grepl('SBayesRC', names(both_cluster))], '*as.factor(gmm)', collapse=' + '))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# 

```

***

### Partitioned PGS

#### Correlation

```{r}
# Read in the partitioned PGS
setwd('/users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline')
library(data.table)
library(ggplot2)
library(cowplot)

source('../functions/misc.R')
source_all('../functions')

# Read in PGS
pgs_gw <- read_pgs(config = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', name = 'ukb', pseudo_only = T, gwas=c('RA','CAD','MIX'))$ukb

pgs_gw <- Reduce(function(x, y) merge(x, y, by = c("FID", "IID"), all = TRUE), lapply(pgs_gw$TRANS, function(x) x[[1]]))

# Read in PGS
pgs_partitioned <- read_pgs_2(config = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', name = 'ukb', partitioned = T, gwas=c('RA','CAD','MIX'))$ukb

pgs_partitioned <- Reduce(function(x, y) merge(x, y, by = c("FID", "IID"), all = TRUE), lapply(pgs_partitioned$TRANS, function(x) x[[1]]))

pgs <- merge(pgs_gw, pgs_partitioned, by = c('FID', 'IID'))
cor_mat<-cor(pgs[,-1:-2])
round(cor_mat[, !grepl('set', colnames(cor_mat))], 2)

# Read in outcome data
outcome_names<-c('CAD','RA','MIX')
outcomes <- list()
for(i in outcome_names){
  outcomes[[i]]<-fread(paste0('~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.', i, '_only.txt'))
  names(outcomes[[i]])[names(outcomes[[i]]) == 'outcome']<-paste0(i,'_outcome')
}
outcomes <- Reduce(function(x, y) merge(x, y, by = c("FID", "IID"), all = TRUE), outcomes)

both<-merge(outcomes, pgs, by = c('FID','IID'))

# Predict each outcome using GW and stratified PGS
library(glmnet)

res<-NULL
for(i in c('MIX','RA','CAD')){
  out1 <-
    cv.glmnet(x = as.matrix(both[, grepl(paste0(i, '_SBayesRC'), names(both)), with = F]), y = both$MIX_outcome)
  
  out2 <-
    cv.glmnet(x = as.matrix(cbind(0, both[, paste0(i, '_SBayesRC'), with = F])), y = both$MIX_outcome)

  # Get predicted values using the optimal lambda
  pred1 <- predict(out1, newx = as.matrix(both[, grepl(paste0(i, '_SBayesRC'), names(both)), with = F]), s = "lambda.min")
  pred2 <- predict(out2, newx = as.matrix(cbind(0, both[, paste0(i, '_SBayesRC'), with = F])), s = "lambda.min")
  
  # Compute correlation (Observed vs. Predicted)
  cor1 <- cor(pred1, both$MIX_outcome)
  cor2 <- cor(pred2, both$MIX_outcome)
  
  res<-rbind(res,
             data.frame(
               outcome = i,
               model = c('gw-pgs + p-pgs', 'gw-pgs'),
               cor = c(cor1, cor2), 
               n_pred = c(1, sum(grepl(paste0(i, '_SBayesRC'), names(both))))
             ))
}

ggplot(res, aes(x=outcome, y = cor, fill = model)) +
  geom_bar(stat="identity", position = position_dodge())

res_wide<-reshape(res, direction = 'wide', timevar = 'model', idvar = 'outcome')
res_wide$diff<-res_wide$`cor.gw-pgs + p-pgs` - res_wide$`cor.gw-pgs`
res_wide$perc_diff<-res_wide$diff / res_wide$`cor.gw-pgs`

#   outcome cor.gw-pgs + p-pgs n_pred.gw-pgs + p-pgs cor.gw-pgs n_pred.gw-pgs         diff   perc_diff
# 1     MIX         0.09020498                     1 0.08504137            33 0.0051636170 0.060718886
# 3      RA         0.03816091                     1 0.03159778            44 0.0065631312 0.207708618
# 5     CAD         0.19752735                     1 0.19699081             6 0.0005365409 0.002723685

# The absolute and relative improvement is highest for RA, then MIX, with very little improvement for CAD.
# The number of pathway specific PGS is also highest for RA and then MIX.
# The improvement could be due to over fitting.
# Confusing that the CAD PGS is so strong, but the number sig gene sets is higher for RA. I suppose this is due to the CAD GWAS being larger, but also more complex biologically.

# As example of disease stratification. Predict CAD vs RA, using MIX GW PGS and paritinoed PGS.
both$CAD_vs_RA_outcome <- NA
both$CAD_vs_RA_outcome[both$CAD_outcome == 1 | both$RA_outcome == 1] <- 0
both$CAD_vs_RA_outcome[both$CAD_outcome == 1] <- 1
both$CAD_vs_RA_outcome[both$CAD_outcome == 1 & both$RA_outcome == 1] <- NA
both_tmp<-both[complete.cases(both),]

i<-'MIX'
out1 <-
  cv.glmnet(x = as.matrix(both_tmp[, grepl(paste0(i, '_SBayesRC'), names(both_tmp)), with = F]), y = both_tmp$CAD_vs_RA_outcome)

out2 <-
  cv.glmnet(x = as.matrix(cbind(0, both_tmp[, paste0(i, '_SBayesRC'), with = F])), y = both_tmp$CAD_vs_RA_outcome)

# Get predicted values using the optimal lambda
pred1 <- predict(out1, newx = as.matrix(both_tmp[, grepl(paste0(i, '_SBayesRC'), names(both_tmp)), with = F]), s = "lambda.min")
pred2 <- predict(out2, newx = as.matrix(cbind(0, both_tmp[, paste0(i, '_SBayesRC'), with = F])), s = "lambda.min")

# Compute correlation (Observed vs. Predicted)
cor1 <- cor(pred1, both_tmp$CAD_vs_RA_outcome)
cor2 <- cor(pred2, both_tmp$CAD_vs_RA_outcome)

tmp <- data.frame(
  outcome = 'CAD vs RA',
  model = c('gw-pgs + p-pgs', 'gw-pgs'),
  cor = c(cor1, cor2)
)

ggplot(tmp, aes(x=outcome, y = cor, fill = model)) +
  geom_bar(stat="identity", position = position_dodge())

```

***

#### Clustering

```{r}

library(NbClust)

both$group<-NULL
both$group[both$CAD_outcome == 1]<-'CAD'
both$group[both$RA_outcome == 1]<-'RA'
both$group[both$CAD_outcome == 1 & both$RA_outcome == 1]<-'BOTH'
both$group[both$CAD_outcome != 1 & both$RA_outcome != 1]<-'CONTROL'

# Extract PGS data for MIX cases only
both_cases <- both[which(both$MIX_outcome == 1),]
pgs_cases <- both_cases[, grepl('MIX_SBayesRC', names(both_cases)), with=F]
pgs_cases <- scale(pgs_cases) 
# Not sure whether I should be scaling since they are reference standardised

# Determine optimal number of clusters
n_clust_sol <-
  NbClust(
    data = pgs_cases[1:2000,],
    distance = "euclidean",
    min.nc = 2,
    max.nc = 10,
    method = 'ward.D2',
    index = 'all'
  )

n_clust_opt<-length(unique(n_clust_sol$Best.partition))

#####
# K-means
#####

# Now try k-means clustering with 3 clusters
k_res<-kmeans(pgs_cases, n_clust_opt)

# Plot the mean of each group
k_res_centers<-data.frame(Group=as.character(1:n_clust_opt),
                          k_res$centers)

library(reshape2)
k_res_centers_melt<-melt(k_res_centers, id='Group')

ggplot(k_res_centers_melt, aes(x=variable, y=value, group=Group, color=Group)) +
  geom_point(size=5) +
  geom_line() +
  labs(x='Polygenic Score', y='Cluster Mean', title='Mean Polygenic Score Across Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Hierarchical clustering
#####

# Compute distance matrix
dist_matrix <- dist(pgs_cases, method = "euclidean")

# Perform hierarchical clustering
hclust_res <- hclust(dist_matrix, method = "ward.D2")

# Cut tree into n_clust_opt clusters
hclust_clusters <- cutree(hclust_res, k = n_clust_opt)

# Convert clusters to a factor for plotting
hclust_clusters <- as.factor(hclust_clusters)

# Create a data frame with cluster assignments
pgs_cases_clustered <- data.frame(pgs_cases, Cluster = hclust_clusters)

# Compute the mean of each cluster for each PGS
hclust_centers <- aggregate(. ~ Cluster, data = pgs_cases_clustered, FUN = mean)

# Reshape the data for plotting
hclust_centers_melt <- melt(hclust_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
hclust_centers_melt$Cluster <- as.character(hclust_centers_melt$Cluster)

ggplot(hclust_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across Hierarchical Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#####
# Gaussian mixture model
#####
library(mclust)

gmm_model <- Mclust(pgs_cases, G = n_clust_opt)
summary(gmm_model)

gmm_clusters <- gmm_model$classification  # Hard cluster labels
gmm_probabilities <- gmm_model$z  # Soft probabilities for each cluster

# Create a data frame with cluster assignments
pgs_cases_clustered <- data.frame(pgs_cases, Cluster = gmm_clusters)

# Compute the mean of each cluster for each PGS
gmm_centers <- aggregate(. ~ Cluster, data = pgs_cases_clustered, FUN = mean)

# Reshape the data for plotting
gmm_centers_melt <- melt(gmm_centers, id = "Cluster")

# Convert Cluster to character for consistent plotting
gmm_centers_melt$Cluster <- as.character(gmm_centers_melt$Cluster)

ggplot(gmm_centers_melt, aes(x = variable, y = value, group = Cluster, color = Cluster)) +
  geom_point(size = 5) +
  geom_line() +
  labs(x = 'Polygenic Score', y = 'Cluster Mean', title = 'Mean Polygenic Score Across GMM Clusters') +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  background_grid(major = 'y', minor = 'y')

#########
# Compare the clusters to the case groups
k_means_conf_matrix <- table(k_res$cluster, both_cases$group)
hclust_D2_conf_matrix <- table(hclust_clusters, both_cases$group)
gmm_conf_matrix <- table(gmm_clusters, both_cases$group)

library(mclust)
adjustedRandIndex(k_res$cluster, both_cases$group)
adjustedRandIndex(hclust_clusters, both_cases$group)
adjustedRandIndex(gmm_clusters, both_cases$group)

# The accuracy of the hclust solution is higher
pca_res <- prcomp(pgs_cases)
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(k_res$cluster))) + 
  geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(hclust_clusters))) + 
  geom_point()
ggplot(data.frame(pca_res$x), aes(PC1, PC2, color = as.factor(gmm_clusters))) + 
  geom_point()

library(Rtsne)

tsne_res <- Rtsne(pgs_cases[!duplicated(pgs_cases),])

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(k_res$cluster[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(hclust_clusters[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')

ggplot(data.frame(tsne_res$Y), aes(
  x = X1,
  y = X2,
  color = as.factor(gmm_clusters[!duplicated(pgs_cases)])
)) + 
  geom_point() +
  labs(color = 'Cluster')


########
# Predict MIX outcome 

# Project cluster solutions into full dataset
both_cluster<-both

library(clue)
# kmeans
kmeans_pred<-cl_predict(k_res, newdata = both, type = "memberships")
both_cluster$kmean<-apply(kmeans_pred, 1, function(x) which(x == 1))

# hclust
centroids <- aggregate(pgs_cases, list(hclust_clusters), mean)[, -1]  # Remove the cluster ID column
assign_to_nearest_cluster <- function(new_data, centroids) {
  apply(as.matrix(new_data), 1, function(row) {
    which.min(colSums((t(centroids) - row)^2))  # Compute Euclidean distance to centroids
  })
}
both_cluster$hclust <-
  assign_to_nearest_cluster(
    new_data = both[, grepl('MIX_SBayesRC', names(both_cases)), with = F], centroids = centroids)

# gmm
both_cluster$gmm <-
  predict(gmm_model, 
          newdata = both[, grepl('MIX_SBayesRC', names(both_cases)), with = F])$classification
both_cluster$gmm_prob <-
  predict(gmm_model, 
          newdata = both[, grepl('MIX_SBayesRC', names(both_cases)), with = F])$z[,1]

# MIX_SBayesRC only
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC")),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# MIX_SBayesRC + pathway specific PGS
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~  MIX_SBayesRC + ", paste(names(both_cluster)[grepl('MIX_SBayesRC', names(both_cluster))], collapse='+'))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# kmeans interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~ ", paste(names(both_cluster)[grepl('MIX_SBayesRC', names(both_cluster))], '*as.factor(kmean)', collapse='+'))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# hclust interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~ ", paste(names(both_cluster)[grepl('MIX_SBayesRC', names(both_cluster))], '*as.factor(hclust)', collapse='+'))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# gmm interactions
sum_mod<-summary(
  mod<-glm(
    as.formula(paste0("MIX_outcome ~ ", paste(names(both_cluster)[grepl('MIX_SBayesRC', names(both_cluster))], '*as.factor(gmm)', collapse='+'))),
    data = both_cluster))
cor(both_cluster$MIX_outcome, predict(mod, data=both_cluster))

# Prediction does improve, but likely due to over fitting. Given extreme heterogeneity in this example, I am not that.
```

***

### Non-negative factorisation

See whether non-negative matrix factorisation could recapitulate disease-specific PGS.

Steps:
  - Perform LD clumping on MIX GWAS
  - Create matrix of GWAS Z-scores for top MIX hits, and disease specific GWAS
  - Perform non-negative matrix factorisation
  - Generate new PGS based on factor loadings for each variant.
  - Evaluate performance in UKB

```{r}
# conda activate .snakemake/conda/329e0288cb99508f5e6c50a0996b234c_
library("optparse")

option_list = list(
make_option("--ref_plink_chr", action="store", default=NULL, type='character',
		help="Path to per chromosome reference PLINK files [required]"),
make_option("--ref_keep", action="store", default=NULL, type='character',
    help="Keep file to subset individuals in reference for clumping [optional]"),
make_option("--plink2", action="store", default='plink2', type='character',
    help="Path PLINKv2 software binary [required]"),
make_option("--sumstats", action="store", default=NULL, type='character',
		help="GWAS summary statistics in LDSC format [required]"),
make_option("--pTs", action="store", default='5e-8,1e-6,1e-4,1e-2,0.1,0.2,0.3,0.4,0.5,1', type='character',
		help="List of p-value thresholds for scoring [optional]"),
make_option("--nested", action="store", default=T, type='logical',
    help="Specify as F to use non-overlapping p-value intervals [optional]"),
make_option("--top_hla", action="store", default=T, type='logical',
		help="Retain only top assocaited variant in HLA/MHC region [optional]")
)

opt = parse_args(OptionParser(option_list = option_list))

opt$ref_plink_chr<-'resources/data/ref/ref.chr'
opt$ref_keep<-'resources/data/ref/keep_files/EUR.keep'
opt$sumstats<-'~/oliverpainfel/Data/ukb/GenoPred/output_genoclust/reference/gwas_sumstat/MIX/MIX-cleaned.gz'

# Load dependencies
library(GenoUtils)
library(data.table)
source('../functions/misc.R')
source_all('../functions')

# Create temp directory
tmp_dir<-tempdir()

#####
# Format pT option
#####

opt$pTs <- as.numeric(unlist(strsplit(opt$pTs, ',')))

#####
# Read in sumstats
#####

# Read in, check and format GWAS summary statistics
gwas <- read_sumstats(sumstats = opt$sumstats, chr = CHROMS, extract = opt$extract, req_cols = c('CHR','BP','SNP','A1','A2','BETA','P'))

#####
# Prepare for ptclump
#####

if(opt$top_hla){
  # Assumes BP column is GRCh37
  hla <- gwas[(gwas$CHR == 6 & gwas$BP > 28e6 & gwas$BP < 34e6),]
  top_hla <- hla$SNP[hla$P == min(hla$P)][1]
  gwas <- gwas[!(gwas$CHR == 6 & gwas$BP > 28e6 & gwas$BP < 34e6 & gwas$SNP != top_hla),]
}

#####
# Clump SNPs in GWAS based on LD in the reference
#####

clumped <- plink_clump(pfile = opt$ref_plink_chr, plink2 = opt$plink2, chr = CHROMS, sumstats = gwas, keep = opt$ref_keep)

gwas_clumped<-gwas[gwas$SNP %in% clumped,]

write.table(gwas_clumped, '~/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.clumped', row.names=F, quote=F)

```

```{r}
# Load dependencies
library(GenoUtils)
library(data.table)
source('../functions/misc.R')
source_all('../functions')

mix_clumped <- fread('~/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.clumped')
mix_clumped <- mix_clumped[mix_clumped$P < 5e-8,]

mix_gwas<-fread('~/oliverpainfel/Data/ukb/GenoPred/output_genoclust/reference/gwas_sumstat/MIX/MIX-cleaned.gz')
cad_gwas<-fread('~/oliverpainfel/Data/ukb/GenoPred/output_genoclust/reference/gwas_sumstat/CAD/CAD-cleaned.gz')
ra_gwas<-fread('~/oliverpainfel/Data/ukb/GenoPred/output_genoclust/reference/gwas_sumstat/RA/RA-cleaned.gz')


# Would normally have to find proxies or impute.
mix_gwas <- mix_gwas[mix_gwas$SNP %in% mix_clumped$SNP,]
cad_gwas <- cad_gwas[cad_gwas$SNP %in% mix_clumped$SNP,]
ra_gwas <- ra_gwas[ra_gwas$SNP %in% mix_clumped$SNP,]

# Flip mix_gwas to be risk increasing
mix_gwas$A1_new[mix_gwas$BETA < 0]<-mix_gwas$A2[mix_gwas$BETA < 0]
mix_gwas$A2_new[mix_gwas$BETA < 0]<-mix_gwas$A1[mix_gwas$BETA < 0]
mix_gwas$A1[!is.na(mix_gwas$A1_new)]<-mix_gwas$A1_new[!is.na(mix_gwas$A1_new)]
mix_gwas$A2[!is.na(mix_gwas$A2_new)]<-mix_gwas$A2_new[!is.na(mix_gwas$A2_new)]
mix_gwas$BETA[mix_gwas$BETA < 0]<- -mix_gwas$BETA[mix_gwas$BETA < 0]

# Compute Z score
cad_gwas$Z_CAD<-cad_gwas$BETA/cad_gwas$SE
ra_gwas$Z_RA<-ra_gwas$BETA/ra_gwas$SE

# Scale Z scores by sample size
cad_gwas$Z_MIX <- cad_gwas$Z_MIX * sqrt(cad_gwas$N)
ra_gwas$Z_MIX <- ra_gwas$Z_MIX * sqrt(ra_gwas$N)

# Flip effects to match reference alleles
ref <- mix_gwas[, c('SNP','A1','A2'), with=F]
cad_gwas<-cad_gwas[,c('SNP','A1','A2','Z_CAD'), with=F]
ra_gwas<-ra_gwas[,c('SNP','A1','A2','Z_RA'), with=F]
cad_gwas <- map_score(ref = ref, score = cad_gwas)
ra_gwas <- map_score(ref = ref, score = ra_gwas)

z_dat<-Reduce(function(dtf1, dtf2) merge(dtf1, dtf2, by = c("SNP", "A1", "A2"), all = TRUE), 
       list(
         cad_gwas,
         ra_gwas))

library(NMF)

# Separate positive and negative Z-scores
z_dat <- z_dat[,-1:-3]

# Create positive matrix (keep only positive values, set negatives to 0)
Z_pos <- z_dat
Z_pos[Z_pos < 0] <- 0

# Create negative matrix (keep only negative values, set positives to 0, then multiply by -1)
Z_neg <- z_dat
Z_neg[Z_neg > 0] <- 0
Z_neg<-abs(Z_neg)

# Combine them side by side (columns are doubled)
X_matrix <- cbind(Z_pos, Z_neg)

# Convert to matrix format
X_matrix <- as.matrix(X_matrix)

# Determine optimal K (Number of Clusters)
nmf_rank <- 2:5
nmf_estimates <- nmf(X_matrix, nmf_rank, method = "brunet", nrun = 10)

# Extract best rank (K)
best_K <- nmf_rank[which.min(nmf_estimates$measures$dispersion)]
print(paste("Optimal K:", best_K))

# Run final NMF with optimal K
nmf_result <- nmf(X_matrix, rank = best_K, method = "brunet")

# Extract factorized matrices W and H
W <- basis(nmf_result)  # SNP-cluster associations
H <- coef(nmf_result)   # Trait-cluster associations

H_normalized <- apply(H, 2, function(x) x / sum(x))

# Visualize contribution of MIX GWAS to CAD vs. RA
barplot(H_normalized, beside = TRUE)

# Apply cut off of 0.75 to weights
W[W < 0.75] <- 0

# Create score file based on cluster weights
score <- data.frame(ref, W)

dir.create('~/oliverpainfel/Analyses/GenoClust/artificial/nmf/')
for(i in 1:ncol(W)){
  score_i <- data.frame(ref, W[,i])
  names(score_i)<-c('rsID','effect_allele','other_allele', 'effect_weight')
  write.table(
    score_i,
    paste0(
      '~/oliverpainfel/Analyses/GenoClust/artificial/nmf/cluster_',
      i,
      '.score'
    ),
    row.names = F,
    quote = F
  )
}

```

```{r}
library(data.table)

dir.create('/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset')

######
# score_list
######

score_list<-data.frame(
  name=c('nmf_1','nmf_2'),
  path=c(
    '/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/nmf/cluster_1.score',
    '/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/nmf/cluster_2.score'
  ),
  label=paste0('"', c('nmf_1','nmf_2'),'"')
)

write.table(score_list, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/score_list.txt', col.names=T, row.names=F, quote=F)

######
# config
######

config<-c(
  "outdir: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/output_genoclust",
  "config_file: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml",
  "gwas_list: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/gwas_list.txt",
  "score_list: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/score_list.txt",
  "target_list: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/target_list.txt",
  "pgs_methods: ['sbayesrc']",
  "cores_prep_pgs: 10",
  "cores_target_pgs: 10",
  "sbayesrc_ldref: /users/k1806347/oliverpainfel/Data/hgdp_1kg/sbayesrc/hm3"
)

write.table(config, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', col.names = F, row.names = F, quote = F)

```

```{bash}
cd /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline
conda activate genopred

snakemake \
  --profile slurm \
  --use-conda \
  --configfile=/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml \
  output_all -n
```

I haven't fully understood this NMF process yet. 

