---
title: Cross-population evaluation of polygenic scores
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 2
    css: styles/styles.css
    includes:
      in_header: header.html
      after_body: footer.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{css, echo=F}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

***

In this study, we will investigate the utility if polygenic scores for stratifying individual into aetiological subtypes, and the predictive utility of modelling interactions between polygenic score and subtype membership. Our initial investigation will look at pathways specific PGS.

We have previously conducted a very simple proof of concept study [here](https://opain.github.io/GenoClust/proof_of_principle.html). Here will conduct a more realistic simulation to determine feasibility and refine the methodology, and then apply the methodology to complex disease of interest.

Things to think about:
- Derive clusters using cases only, and then project in to full sample, or derive clusters within full sample?
- Clustering method: K-means, Gaussian mixture model, hierarchical clustering

***

# Proof-of-concept

## Simulated data

We could simulate genotype and phenotype data using HAPNEST, two create two phenotypes with causal variants within different pathways.

Simulate two binary phenotypes, with a genetic correlation of 0. Specify causal variants present within a gene set for each disease:
- Disease 1: GO:0006281 DNA repair
- Disease 2: GO:0048167 — Regulation of synaptic plasticity

***

### Define causal variants

Download GO data to identify genes within these sets. 

```{r}
# Read in .gmt file
gmt <- readLines('/scratch/prj/oliverpainfel/recovered/Data/MSigDB/c5.go.bp.v7.5.1.entrez.gmt')
gene_sets <- strsplit(gmt, "\t")
gene_list <- setNames(lapply(gene_sets, `[`, -c(1,2)), sapply(gene_sets, `[`, 1))

# Subset sets of interest
gene_list <- gene_list[ c('GOBP_DNA_REPAIR', 'GOBP_REGULATION_OF_SYNAPTIC_PLASTICITY')]

# Identify variants within genes within each set
gene_loc <- fread('/scratch/prj/oliverpainfel/recovered/Data/Gene_Locations/MAGMA_NCBI37.3/NCBI37.3.gene.loc')
names(gene_loc)<-c('ID','CHR','START','STOP','STRAND','NAME')
gene_loc <- gene_loc[gene_loc$CHR == 22,] #out simulation will only use chr22 data

# Read in hm3 SNP-list
rsids<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/rsid_map_list_chr22.txt')
snplist<-fread('~/oliverpainfel/Data/HAPNEST/data/inputs/processed/1KG+HGDP/hapmap_variant_list_chr22.txt', header=F)$V1
snplist_tab<-data.frame(do.call(rbind, lapply(strsplit(snplist, split = ':'), function(x) c(x))))
names(snplist_tab) <- c('CHR', 'BP', 'A1', 'A2')
snplist_tab$CHR <- gsub('chr', '', snplist_tab$CHR)
rsids<-rsids[match(snplist, rsids$id_hg38),]
snplist_tab$SNP <- rsids$rsid

# Create SNP-list for each gene set
dir.create('~/oliverpainfel/Data/HAPNEST/synth_2', recursive = T)

set.seed(1)
snp_sets <- list()
for(i in names(gene_list)){
  snp_sets[[i]] <- NULL
  genes <- gene_list[[i]]
  genes <- genes[genes %in% gene_loc$ID]
  for(k in genes){
    CHR <- gene_loc$CHR[gene_loc$ID == k] 
    START <- gene_loc$START[gene_loc$ID == k] 
    STOP <- gene_loc$STOP[gene_loc$ID == k]
    
    snp_set <- snplist_tab$SNP[which(snplist_tab$CHR == CHR & snplist_tab$BP >= START & snplist_tab$BP <= STOP)]
    if(!(length(snp_set) > 0)){
      next
    }
    
    # Select 1 variant from each gene to be causal
    snp_sets[[i]] <- c(snp_sets[[i]], sample(snp_set, 1))
  }
  snp_sets[[i]] <- unique(snp_sets[[i]])
}

# 
causal_var<-data.frame(SNP = unlist(snp_sets))
for(i in 1:length(snp_sets)){
  causal_var[[paste0('Trait_', i)]]<-as.numeric(causal_var$SNP %in% snp_sets[[i]])
}

# Save as csv
write.table(causal_var, '~/oliverpainfel/Data/HAPNEST/synth_2/causal.snplist', row.names = F, col.names = F, quote = F, sep=',')

```

***

### Run HAPNEST

<details><summary>Show code</summary>

```{bash}
cd ~/oliverpainfel/Data/HAPNEST

# Generate genotype and phenotype data
singularity exec \
  --bind data/:/data/ \
  --bind /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/misc/hapnest/config.synth_2.yaml:/data/config.synth_2.yaml \
  containers/intervene-synthetic-data_latest.sif \
  generate_geno \
  8 \
  data/config.synth_2.yaml
  
# Keep getting an error due to memory allocation. In the end I ran it with 1 core

singularity exec \
  --bind data/:/data/ \
  --bind /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/misc/hapnest/config.synth_2.yaml:/data/config.synth_2.yaml \
  containers/intervene-synthetic-data_latest.sif \
  generate_pheno \
  data/config.synth_2.yaml
  
```

</details>

***

### Run GSEA

We need to run a GWAS of each phenotyepe, and then MAGMA to see whether our simulation has lead to an enrichment of selected pathways

***

#### Perform PCA

<details><summary>Show code</summary>

```{r}
# conda activate .snakemake/conda/ea13b6c549c70695534894feeeecf0b3_
setwd('~/oliverpainfel/GenoPred/pipeline/')

start.time <- Sys.time()
library("optparse")

option_list = list(
  make_option("--target_plink_chr", action="store", default=NULL, type='character',
              help="Path to per chromosome target PLINK files [required]"),
  make_option("--maf", action="store", default=0.05, type='numeric',
              help="Minor allele frequency threshold [optional]"),
  make_option("--geno", action="store", default=0.02, type='numeric',
              help="Variant missingness threshold [optional]"),
  make_option("--hwe", action="store", default=1e-6, type='numeric',
              help="Hardy Weinberg p-value threshold. [optional]"),
  make_option("--n_pcs", action="store", default=10, type='numeric',
              help="Number of PCs (min=4) [optional]"),
  make_option("--plink2", action="store", default='plink2', type='character',
              help="Path PLINK2 software binary [required]"),
  make_option("--keep_list", action="store", default=NULL, type='character',
              help="File containing list of keep files and corresponding population code [optional]"),
  make_option("--unrel", action="store", default=NA, type='character',
              help="File containing list of unrelated individuals [optional]"),
  make_option("--n_cores", action="store", default=1, type='numeric',
              help="Number of cores for parallel computing [optional]"),
  make_option("--test", action="store", default=NA, type='character',
              help="Specify test mode [optional]"),
  make_option("--output", action="store", default=NULL, type='character',
              help="Path for output files [required]")
)

opt = parse_args(OptionParser(option_list=option_list))

opt$target_plink_chr<-'/users/k1806347/oliverpainfel/HAPNEST/data/outputs/synth_1/synth_1_chr-'
opt$output<-'~/oliverpainfel/Analysis/HAPNEST/synth_1/pca/'
opt$test<-'chr22'

keep_list<-data.frame(
  file = c(
    '~/oliverpainfel/Analysis/HAPNEST/synth_1/training.EUR.txt',
    '~/oliverpainfel/Analysis/HAPNEST/synth_1/training.EAS.txt',
    '~/oliverpainfel/Analysis/HAPNEST/synth_1/training.AFR.txt'),
  POP = c('EUR','EAS','AFR')
)

library(GenoUtils)
library(data.table)
source('../functions/misc.R')
source_all('../functions')

# Create temp directory
tmp_dir<-tempdir()

if(!is.na(opt$test)){
  CHROMS <- as.numeric(gsub('chr','',opt$test))
}

############
# Create file listing variants in regions of long range LD
############

targ_pvar <- read_bim(opt$target_plink_chr, chr = CHROMS)
targ_pvar <- remove_regions(dat = targ_pvar, regions = long_ld_coord)

for(pop in keep_list$POP){
  # Read in keep file for population
  keep_file <- fread(keep_list$file[keep_list$POP == pop], header=F)
  if(ncol(keep_file) == 1){
    keep_file <- data.table(
      FID = keep_file$V1,
      IID = keep_file$V1)
  } else {
    keep_file <- data.table(
      FID = keep_file$V1,
      IID = keep_file$V2)
  }

  ###########
  # Perform PCA on QC'd and independent variants
  ###########

  # Create QC'd SNP-list
  target_qc_snplist <- plink_qc_snplist(bfile = opt$target_plink_chr, plink2 = opt$plink2, chr = CHROMS, keep = keep_file, maf = opt$maf, geno = opt$geno, hwe = opt$hwe, threads = opt$n_cores)

  # Remove high LD regions
  target_qc_snplist <- target_qc_snplist[target_qc_snplist %in% targ_pvar$SNP]

  # Perform LD pruning
  ld_indep <- plink_prune(bfile = opt$target_plink_chr, chr = CHROMS, keep = keep_file, plink2 = opt$plink2, extract = target_qc_snplist, threads = opt$n_cores)

  # To improve efficiency, derive PCs using random subset of 1000 individuals.
  keep_file_subset <- keep_file[sample(1000, replace = T),]
  keep_file_subset <- keep_file_subset[!duplicated(keep_file_subset),]

  # Run PCA
  snp_weights <- plink_pca(bfile = opt$target_plink_chr, keep = keep_file_subset, chr = CHROMS, plink2 = opt$plink2, extract = ld_indep, n_pc = opt$n_pcs, threads = opt$n_cores)
  fwrite(snp_weights, paste0(tmp_dir,'/ref.eigenvec.var'), row.names = F, quote=F, sep=' ', na='NA')

  # Project into the full population
  target_pcs <- plink_score(bfile = opt$target_plink_chr, keep = keep_file, chr = CHROMS, plink2 = opt$plink2, score = paste0(tmp_dir,'/ref.eigenvec.var'), threads = opt$n_cores)

  fwrite(target_pcs, paste0(opt$output, pop,'.pcs.txt'), quote=F, sep=' ', na='NA')
}

```

</details>

***

#### Perform GWAS

<details><summary>Show code</summary>

```{bash}

module add plink2
for pheno in $(seq 1 1); do
    awk 'BEGIN {OFS="\t"} NR==1 {print "FID", "IID", "pheno"} NR>1 {print $1, $1, $NF}' /users/k1806347/oliverpainfel/HAPNEST/data/outputs/synth_1/synth_1_chr.pheno${pheno} > /users/k1806347/oliverpainfel/HAPNEST/data/outputs/synth_1/synth_1_chr.pheno${pheno}.plink
    
  for pop in $(echo EUR EAS AFR); do
  
    mkdir -p ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}
    for chr in $(seq 22 22); do
        sbatch -p neurohack_cpu --mem 20G -n 4 --wrap="plink2 \
          --bfile /users/k1806347/oliverpainfel/HAPNEST/data/outputs/synth_1/synth_1_chr-22 \
          --pheno /users/k1806347/oliverpainfel/HAPNEST/data/outputs/synth_1/synth_1_chr.pheno${pheno}.plink \
          --1 \
          --covar ~/oliverpainfel/Analysis/HAPNEST/synth_1/pca/${pop}.pcs.txt \
          --covar-variance-standardize \
          --logistic omit-ref cols=+a1freq,+ax hide-covar \
          --maf 0.01 \
          --geno 0.05 \
          --out ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.chr${chr}"
    done
  done
done

# Once complete, merge results across chromosomes
for pheno in $(seq 1 1); do
  for pop in $(echo EUR EAS AFR); do
    head -n 1 ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.chr22.pheno.glm.logistic.hybrid > ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.GW.pheno.glm.logistic.hybrid
      for chr in $(seq 22 22); do
        tail -n +2 ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.chr22.pheno.glm.logistic.hybrid >> ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.GW.pheno.glm.logistic.hybrid
      done
      
      # Remove REF and ALT columns and rename AX column to A2
      cut -f 4,5 --complement ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.GW.pheno.glm.logistic.hybrid | awk 'BEGIN{FS=OFS="\t"} NR==1 {$7="A2"} 1' > temp.txt && mv temp.txt ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.GW.pheno.glm.logistic.hybrid
  
      gzip ~/oliverpainfel/Analysis/HAPNEST/synth_1/gwas/pheno${pheno}/pheno${pheno}.${pop}.GW.pheno.glm.logistic.hybrid
  done
done

```

</details>

***

#### Run MAGMA

<details><summary>Show code</summary>

```{bash}



```

</details>

***

## Artificial misdiagnosis

Look at the UKB independent sumstats that I have used for PGS comparison. Check PGS R2 and gene set enrichment.

T2D and IBD both have strong PGS R2, and presumably different gene set enrichment.

***

### Run GSEA

```{bash}

mkdir -p ~/oliverpainfel/Analyses/GenoClust/artificial/gwas
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST004773/GCST004773-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T2D.gz
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST004131/GCST004131-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/IBD.gz
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST90013445/GCST90013445-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T1D.gz
cp /scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/GenoPredPipe/output/reference/gwas_sumstat/GCST90013534/GCST90013534-cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA.gz
cp /scratch/prj/oliverpainfel/recovered/Data/GWAS_sumstats/prs_comparison/cleaned/COAD01.cleaned.gz ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD.gz

/scratch/prj/oliverpainfel/recovered/Data/GWAS_sumstats/prs_comparison/cleaned/

gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T2D.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/IBD.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T1D.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA.gz
gunzip ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD.gz

mkdir -p ~/oliverpainfel/Analyses/GenoClust/artificial/magma/

# Gene association
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T2D use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/IBD use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/T1D use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval ~/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gene_level

# GSEA
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gsea
# Surprisingly T2D has very few significantly enriched gene sets

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gsea
  
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gsea
  
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gsea

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gsea
  
```

```{r}
# Compare GSEA results across traits
T2D<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/T2D_gsea.gsa.out', skip = 3)
IBD<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/IBD_gsea.gsa.out', skip = 3)
T1D<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/T1D_gsea.gsa.out', skip = 3)
RA<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gsea.gsa.out', skip = 3)
CAD<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gsea.gsa.out', skip = 3)

T2D <- T2D[, c('FULL_NAME','BETA','P'), with=F]
T2D$TRAIT <- 'T2D'
IBD <- IBD[, c('FULL_NAME','BETA','P'), with=F]
IBD$TRAIT <- 'IBD'
T1D <- T1D[, c('FULL_NAME','BETA','P'), with=F]
T1D$TRAIT <- 'T1D'
RA <- RA[, c('FULL_NAME','BETA','P'), with=F]
RA$TRAIT <- 'RA'
CAD <- CAD[, c('FULL_NAME','BETA','P'), with=F]
CAD$TRAIT <- 'CAD'

both <- rbind(T2D, IBD, T1D, RA, CAD)

both_wide <- reshape(both, 
                     idvar = "FULL_NAME", 
                     timevar = "TRAIT", 
                     direction = "wide")

cor(both_wide[, grepl('BETA', names(both_wide)), with = F], use='p')

both_wide$FDR.T2D <- p.adjust(both_wide$P.T2D, method = 'fdr')
both_wide$FDR.IBD <- p.adjust(both_wide$P.IBD, method = 'fdr')
both_wide$FDR.T1D <- p.adjust(both_wide$P.T1D, method = 'fdr')
both_wide$FDR.RA <- p.adjust(both_wide$P.RA, method = 'fdr')
both_wide$FDR.CAD <- p.adjust(both_wide$P.CAD, method = 'fdr')

library(VennDiagram)

venn.plot <- venn.diagram(
  x = list(T2D = both_wide$FULL_NAME[both_wide$FDR.T2D < 0.05], IBD = both_wide$FULL_NAME[both_wide$FDR.IBD < 0.05]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(T1D = both_wide$FULL_NAME[which(both_wide$FDR.T1D < 0.05)], IBD = both_wide$FULL_NAME[which(both_wide$FDR.IBD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[which(both_wide$FDR.RA < 0.05)], IBD = both_wide$FULL_NAME[which(both_wide$FDR.IBD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[which(both_wide$FDR.RA < 0.05)], T1D = both_wide$FULL_NAME[which(both_wide$FDR.T1D < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[which(both_wide$FDR.RA < 0.05)], CAD = both_wide$FULL_NAME[which(both_wide$FDR.CAD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

venn.plot <- venn.diagram(
  x = list(IBD = both_wide$FULL_NAME[which(both_wide$FDR.IBD < 0.05)], CAD = both_wide$FULL_NAME[which(both_wide$FDR.CAD < 0.05)]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

```
***

### Meta-analyse GWAS

Lets use RA and CAD as they have a low correlation, but decent number of significant sets.

```{bash, eval=F, echo=T}

# Run METAL
/scratch/prj/oliverpainfel/recovered/Software/generic-metal/metal

SCHEME SAMPLESIZE
MARKER SNP
WEIGHT N
ALLELE A1 A2
EFFECT BETA
STDERR SE
PVALUE P
PROCESS /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD
PROCESS /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA
OUTFILE /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA .tbl

ANALYZE

QUIT

```

```{r, eval=F, echo=T}
# Format the METAL output
library(data.table)

# Read in the sumstats
meta<-fread('/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA1.tbl')

# Remove SNPs that were not present in both studies
meta <- meta[!grepl('\\?', meta$Direction),]

# Format for GenoPred
names(meta)<-c('SNP','A1','A2','N','BETA','P','Direction')
meta$A1<-toupper(meta$A1)
meta$A2<-toupper(meta$A2)
meta$SE<-1
meta<-meta[,c('SNP','A1','A2','BETA','SE','N','P'),with=F]

fwrite(meta, '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.tbl.reformat', quote=F, sep=' ', na='NA')

```

```{bash}
# Rerun MAGMA to see how the results compare to the original GWAS
/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --bfile /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma_ref/g1000_eur \
  --pval /scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.tbl.reformat use=SNP,P ncol=N \
  --gene-annot /scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/data/magma/NCBI37.3.genes.annot \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gene_level

/scratch/prj/oliverpainfel/recovered/Software/MyGit/GenoDisc/pipeline/resources/software/magma/magma \
  --gene-results ~/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gene_level.genes.raw \
  --set-annot /scratch/prj/oliverpainfel/recovered/Data/MSigDB/c2.all.v7.5.1.entrez.gmt \
  --out ~/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gsea
  
```

```{r}
# Compare GSEA results across traits
RA<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/RA_gsea.gsa.out', skip = 3)
CAD<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/CAD_gsea.gsa.out', skip = 3)
MIX<-fread('/users/k1806347/oliverpainfel/Analyses/GenoClust/artificial/magma/mix_CAD_RA_gsea.gsa.out', skip = 3)

RA <- RA[, c('FULL_NAME','BETA','P'), with=F]
RA$TRAIT <- 'RA'
CAD <- CAD[, c('FULL_NAME','BETA','P'), with=F]
CAD$TRAIT <- 'CAD'
MIX <- MIX[, c('FULL_NAME','BETA','P'), with=F]
MIX$TRAIT <- 'MIX'

both <- rbind(RA, CAD, MIX)

both_wide <- reshape(both, 
                     idvar = "FULL_NAME", 
                     timevar = "TRAIT", 
                     direction = "wide")

cor(both_wide[, grepl('BETA', names(both_wide)), with = F], use='p')

both_wide$FDR.RA <- p.adjust(both_wide$P.RA, method = 'fdr')
both_wide$FDR.CAD <- p.adjust(both_wide$P.CAD, method = 'fdr')
both_wide$FDR.MIX <- p.adjust(both_wide$P.MIX, method = 'fdr')

library(VennDiagram)

venn.plot <- venn.diagram(
  x = list(RA = both_wide$FULL_NAME[both_wide$FDR.RA < 0.05], 
           CAD = both_wide$FULL_NAME[both_wide$FDR.CAD < 0.05],
           MIX = both_wide$FULL_NAME[both_wide$FDR.MIX < 0.05]),
  filename = NULL,
  output = TRUE
)

dev.off()
grid.draw(venn.plot)

# It is still picking up some pathways for both disorders, but also many new pathways.
both_wide[both_wide$FDR.MIX < 0.05 & both_wide$FDR.RA > 0.05 & both_wide$FDR.CAD > 0.05,]
```

***

### Create phenotype data in UKB

```{r}
#####
# CAD - Use definition from https://doi.org/10.1001/jama.2019.22241
#####

library(ukbkings)
library(dplyr)
library(stringr)

# Extract ICD code and sex from UKB
project_dir <- "/datasets/ukbiobank/ukb82087"
f <- bio_field(project_dir)
f %>%
    select(field, name) %>%
    filter(str_detect(field, "^31-0.0|41270|41271|41272|20002|20004|6150")) %>%
    bio_field_add("~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset.txt")

bio_phen(
    project_dir,
    field = "~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset.txt",
    out = "~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset"
)

df <- readRDS("~/oliverpainfel/Data/ukb/phenotypes/cad_field_subset.rds")

# Load required libraries
library(dplyr)

# Assume 'ukb_data' is your loaded data frame
# ukb_data <- read.csv("your_data_file.csv") # Example for loading data

# CAD code definitions
cad_icd10 <- c("I21", "I22", "I23", "I241", "I252")
cad_icd9 <- c("410", "411", "412", "42789")
cad_opcs4 <- paste0("K", c("401", "402", "403", "404",
                           "411", "412", "413", "414",
                           "451", "452", "453", "454", "455"))
cad_self_reported <- c("1075")
cad_operation <- c("1070", "1095")
cad_vascular <- c("1")

# Columns to search in
icd10_cols <- grep("^41270", names(df), value = TRUE) # ICD-10 columns
icd9_cols <- grep("^41271", names(df), value = TRUE) # ICD-9 columns
opcs4_cols <- grep("^41272", names(df), value = TRUE) # OPCS-4 columns
self_report_cols <- grep("^20002", names(df), value = TRUE) # Self-reported illness
operation_cols <- grep("^20004", names(df), value = TRUE) # Operation codes
vascular_cols <- grep("^6150", names(df), value = TRUE) # Vascular/heart problems

# Create a binary CAD phenotype
df <- df %>%
  mutate(
    CAD = if_else(
      rowSums(across(all_of(icd10_cols), ~ . %in% cad_icd10), na.rm = TRUE) > 0 |
      rowSums(across(all_of(icd9_cols), ~ . %in% cad_icd9), na.rm = TRUE) > 0 |
      rowSums(across(all_of(opcs4_cols), ~ . %in% cad_opcs4), na.rm = TRUE) > 0 |
      rowSums(across(all_of(self_report_cols), ~ . %in% cad_self_reported), na.rm = TRUE) > 0 |
      rowSums(across(all_of(operation_cols), ~ . %in% cad_operation), na.rm = TRUE) > 0 |
      rowSums(across(all_of(vascular_cols), ~ . %in% cad_vascular), na.rm = TRUE) > 0,
      1, 0
    )
  )

# Summary of CAD cases
table(df$CAD)

cad <- data.frame(eid = df$eid, CAD = df$CAD)

#####
# RA
#####

keep_files<-list.files(path = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/output/ukb/pcs/within_sample/', pattern = '.keep')

pop_dat<-NULL
for(i in keep_files){
  tmp<-fread(paste0('/users/k1806347/oliverpainfel/Data/ukb/GenoPred/output/ukb/pcs/within_sample/', i))
  names(tmp)<-c('FID','IID')
  tmp$POP<-gsub('.keep','', gsub('ukb.outlier_detection.','',i))
  pop_dat<-rbind(pop_dat, tmp)
}

# Update row number IDs to project specific IDs
psam<-fread('/scratch/prj/ukbiobank/recovered/ukb82087/imputed/ukb82087_imp_chr1_MAF1_INFO4_v1.psam')
psam$rn<-1:nrow(psam)
psam<-psam[,c('IID','rn'), with = F]

pop_dat$FID<-NULL
pop_dat<-merge(pop_dat, psam, by.x='IID', by.y='rn')
pop_dat<-data.frame(
  eid=pop_dat$IID.y,
  POP=pop_dat$POP
)

# This has been defined already during our UKB benchmark
ra <- fread('/scratch/prj/ukbiobank/recovered/Edinburgh_Data/usr/ollie_pain/phenotypes/ra.unrel.txt')
names(ra)<-c('eid','RA')

both <- merge(ra, cad, by='eid')
both$MIX <- ifelse(both$RA == 1 | both$CAD == 1, 1, 0)
nrow(both[both$RA == 1 & both$CAD == 1,])

# Restrict analysis to EUR individuals
both<-both[both$eid %in% pop_dat$eid[pop_dat$POP == 'EUR']]

# There are only 954  RA cases but >22k CAD cases. This could make it hard for clustering to pick out this cluster - Lets see.
# These are unrelated individuals in UKB already.
# To speed things up, lets use a subset of 25k controls
set.seed(1)
controls<-both$eid[both$MIX == 0]
keep<-sample(controls, 25000)
both<-both[!(both$eid %in% controls[!(controls %in% keep)])]

write.table(both, '~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.txt', row.names = F, col.names = T, quote = F)

# Save with row.number IDs
both<-merge(both, psam, by.x='eid', by.y='IID')

for(i in c('RA','CAD','MIX')){
  pheno<-data.frame(
    FID=both$rn,
    IID=both$rn,
    outcome=both[[i]]
  )
  write.table(pheno, paste0('~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.', i, '_only.txt'), row.names = F, col.names = T, quote = F)
}

keep<-data.frame(FID = both$rn, IID = both$rn)

write.table(keep, '~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.keep', row.names = F, col.names = F, quote = F)

```

***

### Run GenoPred

```{bash}
mkdir ~/oliverpainfel/Data/ukb/genoclust_subset

for chr in $(seq 1 22); do
  ~/oliverpainfel/Software/plink2 \
    --pfile ~/oliverpainfel/Data/ukb/GenoPred/output/ukb/geno/ukb.ref.chr${chr} \
    --keep ~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.keep \
    --make-pgen \
    --out ~/oliverpainfel/Data/ukb/genoclust_subset/ukb.chr${chr}
done

```

```{r}
library(data.table)

dir.create('/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset')

######
# target_list
######

target_list <- data.frame(
  name='ukb',
  path='/users/k1806347/oliverpainfel/Data/ukb/genoclust_subset/ukb',
  type='plink2',
  indiv_report=F,
  unrel='/users/k1806347/oliverpainfel/Data/ukb/phenotypes/unrelated.row_number.txt'
)

write.table(target_list, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/target_list.txt', col.names=T, row.names=F, quote=F)

######
# gwas_list
######

gwas_list<-data.frame(
  name=c('CAD','RA','MIX'),
  path=c(
    '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/CAD',
    '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/RA',
    '/scratch_tmp/prj/oliverpainfel/Analyses/GenoClust/artificial/gwas/mix_CAD_RA.tbl.reformat'
  ),
  population='EUR',
  n=NA,
  sampling=NA,
  prevalence=NA,
  mean=NA,
  sd=NA,
  label=paste0('"', c('CAD','RA','MIX'),'"')
)

write.table(gwas_list, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/gwas_list.txt', col.names=T, row.names=F, quote=F)

######
# config
######

config<-c(
  "outdir: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/output_genoclust",
  "config_file: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml",
  "gwas_list: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/gwas_list.txt",
  "target_list: /users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/target_list.txt",
  "pgs_methods: ['sbayesrc']",
  "cores_prep_pgs: 10",
  "cores_target_pgs: 10",
  "sbayesrc_ldref: /users/k1806347/oliverpainfel/Data/hgdp_1kg/sbayesrc/hm3"
)

write.table(config, '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', col.names = F, row.names = F, quote = F)

```

```{bash}
cd /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline
conda activate genopred

snakemake \
  --profile slurm \
  --use-conda \
  --configfile=/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml \
  output_all -n
```

***

### Check correlation

Check correlation between the PGS for CAD, RA and MIX with CAD, RA and MIX outcomes.

```{r}
setwd('/users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline')
library(data.table)
library(ggplot2)
library(cowplot)

source('../functions/misc.R')
source_all('../functions')

# Define pgs_methods used
pgs_methods <- read_param(config = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', param = 'pgs_methods', return_obj = F)

# Read in PGS
pgs <- read_pgs(config = '/users/k1806347/oliverpainfel/Data/ukb/GenoPred/configs/genoclust_subset/config.yaml', name = 'ukb')$ukb
pgs <- Reduce(function(x, y) merge(x, y, by = c("FID", "IID"), all = TRUE), lapply(pgs$TRANS, function(x) x[[1]]))

# Read in outcome data
outcome_names<-c('CAD','RA','MIX')
outcomes <- list()
for(i in outcome_names){
  outcomes[[i]]<-fread(paste0('~/oliverpainfel/Data/ukb/phenotypes/ra_cad_mix.subset.', i, '_only.txt'))
  names(outcomes[[i]])[names(outcomes[[i]]) == 'outcome']<-paste0(i,'_outcome')
}
outcomes <- Reduce(function(x, y) merge(x, y, by = c("FID", "IID"), all = TRUE), outcomes)

# Test correlation
both <- merge(pgs, outcomes, by = c('FID', 'IID'))

round(cor(both[,-1:-2]), 2)

#              CAD_SBayesRC RA_SBayesRC MIX_SBayesRC CAD_outcome RA_outcome MIX_outcome
# CAD_SBayesRC         1.00       -0.01         0.24        0.20      -0.02        0.20
# RA_SBayesRC         -0.01        1.00         0.87        0.00       0.12        0.03
# MIX_SBayesRC         0.24        0.87         1.00        0.06       0.11        0.09
# CAD_outcome          0.20        0.00         0.06        1.00      -0.10        0.97
# RA_outcome          -0.02        0.12         0.11       -0.10       1.00        0.15
# MIX_outcome          0.20        0.03         0.09        0.97       0.15        1.00

# This looks good so far:
# - The MIX PGS is more correlated with the RA PGS than CAD PGS, due to larger genetic effects on RA
# - The MIX outcome is more correlated with the CAD outcome, due to large number of CAD cases
# - The MIX PGS predicts CAD and RA outcome worse that CAD and RA PGS respectively

# The reason for PGS stratification is becoming clearer - If we were to extract genetic effects relevant to CAD, then the CAD PGS would predict CAD cases better. Likewise for RA.

# It would be interesting to see whether we can cluster individuals based on these scores alone. This would be similar to the T2D paper, which stratifies PGS by relationship with other traits. Our pathway specific strategy will take a different approach.
```


