---
title: "Mind the Gap UKB"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

***

# Setup install required software

```{bash}
# Download script to instance
dx download mind_the_gap/mind_the_gap_ukb.Rmd

# Install some command line programs
sudo apt update
sudo apt install -y tmux less htop 

# Download plink2
wget https://s3.amazonaws.com/plink2-assets/plink2_linux_avx2_20251205.zip
unzip plink2_linux_avx2_20251205.zip
dx download software/plink2
chmod a+x plink2

# Download liftover and chain file
wget -q https://hgdownload.soe.ucsc.edu/goldenPath/hg19/liftOver/hg19ToHg38.over.chain.gz
wget -q https://hgdownload.soe.ucsc.edu/admin/exe/linux.x86_64/liftOver
chmod +x liftOver

# Download and give permission to GreedyRelated binary for linux
wget https://gitlab.com/-/project/14754196/uploads/6fdc44072a8a866cb77c2cf91f68d662/GreedyRelated_linux
chmod a+x GreedyRelated_linux

# Download my script for removing related individuals
wget https://raw.githubusercontent.com/opain/GenoPred/refs/heads/master/Scripts/ukb_relative_remover/ukb_relative_remover.R -O ukb_relative_remover.R

# Download QCTOOL2
wget https://www.well.ox.ac.uk/~gav/resources/qctool_v2.2.0-CentOS_Linux7.8.2003-x86_64.tgz
tar -zxvf qctool_v2.2.0-CentOS_Linux7.8.2003-x86_64.tgz
mv 'qctool_v2.2.0-CentOS Linux7.8.2003-x86_64'/qctool ~/qctool

# Import previous genopred run output for ancestry inference results
tar -xvf /mnt/project/ukb_genopred.tar genopred/output/ukb/ancestry -C ~/

```

```{r}
install.packages(c('data.table','R.utils','ggplot2','cowplot','doMC','foreach','scales','optparse','RNOmni'))
devtools::install_github("opain/GenoUtils")
```

***

# Identify unrelated individuals

```{bash}
mkdir ukb

# Identify list of related individuals (using threshold of 0.044)
Rscript ukb_relative_remover.R \
  --rel_file /mnt/project/Bulk/Genotype\ Results/Genotype\ calls/ukb_rel.dat \
  --rel_thresh 0.044 \
  --seed 1 \
  --GreedyRelated ./GreedyRelated_linux \
  --output ukb/ukb82087
```

***

# Prepare phenotype data

```{r}
# Read in phenotype data
library(data.table)
library(RNOmni)
pheno <- fread('/mnt/project/mind_the_gap/phenotype/data.csv')

# Extract age and sex
age_sex <- data.table(
  FID = pheno$eid,
  IID = pheno$eid,
  age = pheno$p21022,
  sex = as.numeric(factor(pheno$p31, levels = c('Female','Male'))) - 1
)

# Extract phenotypes
fields<-c('30620','30600','30610','30650','30160','21001','21002','30710','30680','4079','30150','30740','30750','30760','50','30030','30020','30780','30120','30050','30060','30040','30130','30140','30080','30010','30700','4080','30690','30860','30870','30000','30730')
trait<-c('Alanine aminotransferase','Albumin','Alkaline phosphatase','Aspartate transaminase','Basophil','Body mass index','Body weight','C-reactive protein','Calcium','Diastolic blood pressure','Eosinophil','Glucose','HbA1c','HDL-cholesterol','Height','Hematocrit','Hemoglobin','LDL-cholesterol','Lymphocyte','Mean corpuscular hemoglobin','Mean corpuscular hemoglobin concentration','Mean corpuscular volume','Monocyte','Neutrophil','Platelet','Red blood cell','Serum creatinine','Sytolic blood pressure','Total cholesterol','Total protein','Triglycerides','White blood cell','γ-glutamyl transpeptidase')
labels<-c('ALT','ALB','ALP','AST','BAS','BMI','BWT','CRP','Ca','DBP','EOS','GLC','HbA1c','HDL','HT','HCT','HB','LDL','LYM','MCH','MCHC','MCV','MON','NEU','PLT','RBC','CR','SBP','TC','TP','TG','WBC','GGT')

# Rename columns and take first instance
for(i in 1:length(labels)){
  if(any(grepl(paste0('p', fields[i],'_|^', fields[i],'$'), names(pheno)))){
    setnames(pheno, names(pheno)[grepl(paste0('p', fields[i],'_|^', fields[i],'$'), names(pheno))][1], labels[i])
  }
}

pheno$FID <- pheno$eid
pheno$IID <- pheno$eid

# Subset phenotype columns
pheno <- pheno[, names(pheno) %in% c('FID','IID', labels), with = F]

# Read in ancestry data
keep_files <- list()
for(i in list.files(path = '~/genopred/output/ukb/ancestry/keep_files/model_based', pattern = '.keep', full.names = T)){
  pop <- gsub('.*/', '', gsub('.keep','', i))
  keep_files [[pop]] <- fread(i)
}

# Read in UKB PCs 1:16
fam<-fread('/mnt/project/Bulk/Genotype Results/Genotype calls/ukb22418_c1_b0_v2.fam')$V1
pcs <- fread('/mnt/project/Bulk/Genotype\ Results/Genotype\ calls/ukb_sqc_v2.txt')
pcs <- pcs[, 26:(26 + 15), with = F]
names(pcs)<-paste0('PC', 1:16)

pcs<- data.table(
  FID = fam,
  IID = fam,
  pcs
)

# Combine PCs, age and sex
covs <- merge(age_sex, pcs, by = c('FID','IID'))
covar_names<-c(paste0('PC',1:16), 'sex', 'age')

# Read in list of unrelated individuals
related <- fread('ukb/ukb82087.related')$V1
unrelated<-fam[!(fam %in% related)]

# Create phenotype files containing a max of 40k unrelated individuals for each population
# For EUR, create a phenotype for the remaining individuals which we can use for GWAS.
pheno_names <- names(pheno)[!(names(pheno) %in% c('FID','IID'))]

geno_keep <- NULL
dir.create('~/ukb/phenotype')
set.seed(1)
# Subset unrelated
pheno_unrel <- pheno[pheno$FID %in% unrelated,]

for(j in names(keep_files)){
  # Subset population
  pheno_j <- pheno_unrel[pheno_unrel$FID %in% keep_files[[j]]$V1,]
  
  # Subset max 40k individuals for test subset
  pheno_j_test <- pheno_j[sample(1:nrow(pheno_j), min(c(nrow(pheno_j), 40000))),]

  fwrite(
    pheno_j_test[, c('FID', 'IID'), with = F],
    paste0('~/ukb/phenotype/ukb.test.', j, '.keep'),
    quote = F,
    col.names = F,
    na = 'NA',
    sep = ' '
  )  
  
  # Subset remainder of EUR individuals for GWAS
  if(j == 'EUR'){
    pheno_j_train <- pheno_j[!(pheno_j$FID %in% pheno_j_test$FID), ]
    fwrite(
      pheno_j_train[, c('FID', 'IID'), with = F],
      paste0('~/ukb/phenotype/ukb.train.', j, '.keep'),
      quote = F,
      col.names = F,
      na = 'NA',
      sep = ' '
    )  
  }
  
  for(i in pheno_names){
    # Subset phenotype
    pheno_j_test_i <- pheno_j_test[, c('FID','IID', i), with = F]
    
    # Merge with covariate data
    pheno_j_test_i_covs <- merge(pheno_j_test_i, covs, by = c('FID','IID'))
    
    # Remove individuals with incomplete data
    pheno_j_test_i_covs <- pheno_j_test_i_covs[complete.cases(pheno_j_test_i_covs),]
    
    # Normalise phenotype
    pheno_j_test_i_covs$pheno_norm <- RNOmni::RankNorm(pheno_j_test_i_covs[[i]])
    
    # Regress covariates
    mod<-lm(as.formula(paste0('pheno_norm ~ ', paste(covar_names, collapse=' + '))), data=pheno_j_test_i_covs)
    pheno_j_test_i_covs$pheno_norm_resid_scale<-as.numeric(scale(resid(mod)))
  
    pheno_j_test_i_covs<-data.frame(
      FID=pheno_j_test_i_covs$FID,
      IID=pheno_j_test_i_covs$IID,
      outcome=pheno_j_test_i_covs$pheno_norm_resid_scale
    )
    
    fwrite(pheno_j_test_i_covs, paste0('~/ukb/phenotype/', i,'.',j,'.txt'), quote = F, na = 'NA', sep = ' ')
    
    if(j == 'EUR'){
      # Subset phenotype
      pheno_j_train_i <- pheno_j_train[, c('FID','IID', i), with = F]
      
      # Merge with covariate data
      pheno_j_train_i_covs <- merge(pheno_j_train_i, covs, by = c('FID','IID'))
      
      # Remove individuals with incomplete data
      pheno_j_train_i_covs <- pheno_j_train_i_covs[complete.cases(pheno_j_train_i_covs),]
      
      # Normalise phenotype
      pheno_j_train_i_covs$pheno_norm <- RNOmni::RankNorm(pheno_j_train_i_covs[[i]])
      
      # Regress covariates
      mod<-lm(as.formula(paste0('pheno_norm ~ ', paste(covar_names, collapse=' + '))), data=pheno_j_train_i_covs)
      pheno_j_train_i_covs$pheno_norm_resid_scale<-as.numeric(scale(resid(mod)))
    
      pheno_j_train_i_covs<-data.frame(
        FID=pheno_j_train_i_covs$FID,
        IID=pheno_j_train_i_covs$IID,
        outcome=pheno_j_train_i_covs$pheno_norm_resid_scale
      )
      
      fwrite(pheno_j_train_i_covs, paste0('~/ukb/phenotype/', i,'.',j,'.gwas.txt'), quote = F, na = 'NA', sep = ' ')
    }
  }
}

```

```{bash}
dx upload -r ~/ukb/phenotype --path mind_the_gap/
```

***

# Prepare the genetic data

To individuals with phenotype data and variants in SBayesRC PGS.

***

## Imputed data

This section is outdated, as updated training and testing subset definitions. Leaving code here as outputs are used later to subset variants by imputation quality in UKB, and subset SBayesRC 7.1M and HapMap3 SNPs in WGS data.

```{bash}
# Create symlinks to the dispensed imputed genetic data
mkdir -p /home/rstudio-server/ukb/ukb_symlinks

# Link bgen and bgen.bgi files for all chromosomes
for chr in $(seq 1 22); do
  for file in $(echo bgen bgen.bgi); do
    ln -s /mnt/project/Bulk/Imputation/UKB\ imputation\ from\ genotype/ukb22828_c${chr}_b0_v3.${file} /home/rstudio-server/ukb/ukb_symlinks/ukb_imp.chr${chr}.${file}
  done
done

# Link the sample file (same for all chromosomes)
ln -s /mnt/project/Bulk/Imputation/UKB\ imputation\ from\ genotype/ukb22828_c1_b0_v3.sample /home/rstudio-server/ukb/ukb_symlinks/ukb_imp.sample

```

```{bash}
# Subset the bgen files to SBayesRC PGS variants and individuals in phenotype files
# Subset variants first to avoid repeatedly reading the large bgen for each population
# /mnt/project/mind_the_gap/genotype/snplist.txt contains variants in the hapmap3 or 7.1M reference for SBayesRC
mkdir -p $HOME/ukb/phenotype

for pop in AFR AMR CSA EAS EUR MID; do
  sed "s/ /_/g" "/mnt/project/mind_the_gap/phenotype/ukb.test.${pop}.keep" > "$HOME/ukb/phenotype/ukb.test.${pop}.idonly.txt"
done

for chr in $(seq 1 22); do
~/plink2 \
  --bgen ukb/ukb_symlinks/ukb_imp.chr${chr}.bgen ref-last \
  --sample ukb/ukb_symlinks/ukb_imp.sample \
  --extract /mnt/project/mind_the_gap/genotype/snplist.txt \
  --export bgen-1.2 \
  --make-pgen \
  --out ukb/ukb_imp_sbayesrc.chr${chr}
  
  dx upload ukb/ukb_imp_sbayesrc.chr${chr}.pgen --path mind_the_gap/genotype/
  dx upload ukb/ukb_imp_sbayesrc.chr${chr}.pvar --path mind_the_gap/genotype/
  dx upload ukb/ukb_imp_sbayesrc.chr${chr}.psam --path mind_the_gap/genotype/
  
  awk 'NR>2 {$1 = $1"_"$2} 1' ukb/ukb_imp_sbayesrc.chr${chr}.sample > ukb/ukb_imp_sbayesrc.chr${chr}.corrected.sample

  # Run 8 jobs at a time using xargs
  for pop in AFR AMR CSA EAS EUR MID; do
    echo "$chr $pop"
  done | xargs -n 2 -P 8 bash -c '
  chr="$0"
  pop="$1"

  ~/qctool \
    -g "ukb/ukb_imp_sbayesrc.chr${chr}.bgen" \
    -s "ukb/ukb_imp_sbayesrc.chr${chr}.corrected.sample" \
    -incl-samples "$HOME/ukb/phenotype/ukb.test.${pop}.idonly.txt" \
    -snp-stats \
    -osnp "ukb/ukb_imp_sbayesrc_${pop}.chr${chr}.snp-stats.txt"
  dx upload "ukb/ukb_imp_sbayesrc_${pop}.chr${chr}.snp-stats.txt" --path mind_the_gap/genotype/
  '
  
  rm ukb/ukb_imp_sbayesrc*.chr${chr}*
done

for chr in $(seq 1 22); do
~/plink2 \
  --bgen ukb/ukb_symlinks/ukb_imp.chr${chr}.bgen ref-last \
  --sample ukb/ukb_symlinks/ukb_imp.sample \
  --extract /mnt/project/mind_the_gap/genotype/snplist.txt \
  --make-pgen \
  --out ukb/ukb_imp_sbayesrc.all.chr${chr}
  
  dx upload ukb/ukb_imp_sbayesrc.all.chr${chr}.pgen --path mind_the_gap/genotype/
  dx upload ukb/ukb_imp_sbayesrc.all.chr${chr}.pvar --path mind_the_gap/genotype/
  dx upload ukb/ukb_imp_sbayesrc.all.chr${chr}.psam --path mind_the_gap/genotype/
  
  rm ukb/ukb_imp_sbayesrc.all*.chr${chr}*
done

```

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json
```

```{r}
# We need to resolve multiallelic variants in plink files
# Change SNP IDs for variants that are not in the dense sbayesrc reference
library(data.table)
score_dense <- fread('/mnt/project/mind_the_gap/score_files/dense/sbayesrc/ref-BMI.score.gz')
score_sparse <- fread('/mnt/project/mind_the_gap/score_files/sparse/sbayesrc/ref-BMI.score.gz')

score <- merge(
  score_dense[, c('SNP', 'A1', 'A2'), with = F], 
  score_sparse[, c('SNP', 'A1', 'A2'), with = F], 
  by = 'SNP',
  all = T)

score$match <- with(score, ((A1.x == A1.y & A2.x == A2.y) | (A1.x == A2.y & A2.x == A1.y)))
score$match[is.na(score$match)] <- F
score$A1<-NA
score$A2<-NA
score$A1[score$match] <- score$A1.x[score$match]
score$A2[score$match] <- score$A2.x[score$match]
score$A1[is.na(score$A1.x)] <- score$A1.y[is.na(score$A1.x)]
score$A2[is.na(score$A1.x)] <- score$A2.y[is.na(score$A1.x)]
score$A1[is.na(score$A1.y)] <- score$A1.x[is.na(score$A1.y)]
score$A2[is.na(score$A1.y)] <- score$A2.x[is.na(score$A1.y)]

score <- score[, c('SNP', 'A1', 'A2'), with = F]

for(i in 1:22){
  print(i)
  pvar <- fread(paste0('/mnt/project/mind_the_gap/genotype/ukb_imp_sbayesrc.chr',i,'.pvar'))
  score_i <- score[score$SNP %in% pvar$ID,]
  
  pvar_score_dense <- merge(
    pvar,
    score_i,
    by.x = 'ID',
    by.y = 'SNP',
    all.x = T,
    sort = F
  )
  
  pvar_score_dense$omit <- with(pvar_score_dense, !((REF == A1 & ALT == A2) | (REF == A2 & ALT == A1)))
  pvar_score_dense$omit[is.na(pvar_score_dense$omit)] <- T
  
  pvar_score_dense$ID[pvar_score_dense$omit] <- paste0(pvar_score_dense$ID[pvar_score_dense$omit], '_omit')
  while(1){
    if(sum(duplicated(pvar_score_dense$ID)) == 0){
      break
    } else {
      pvar_score_dense$ID[duplicated(pvar_score_dense$ID)] <- paste0(pvar_score_dense$ID[duplicated(pvar_score_dense$ID)],'_dup')
    }
  }

  fwrite(
    pvar_score_dense[, c('#CHROM', 'ID', 'POS', 'REF', 'ALT'), with = F], 
    paste0('~/ukb_imp_sbayesrc.chr',i,'.pvar'),
    sep = ' ', quote = F)
}

```

```{bash}
# Make symlinks
for chr in $(seq 1 22); do
  for file in $(echo pgen psam); do
    ln -s /mnt/project/mind_the_gap/genotype/ukb_imp_sbayesrc.chr${chr}.${file} ~/ukb_imp_sbayesrc.chr${chr}.${file}
  done
done

# This means the updated pvar is not saved in the project folder for the next session.
```

```{bash}
# Merge each chromosome and extract PGS SNP (removing multiallelic)
rm merge_list.txt
for i in $(seq 1 22); do
  echo ~/ukb_imp_sbayesrc.chr${i} >> ~/merge_list.txt
done

~/plink2 \
  --pmerge-list ~/merge_list.txt \
  --extract /mnt/project/mind_the_gap/genotype/snplist.txt \
  --make-pgen \
  --out ~/ukb_imp_sbayesrc

# Subset by population
set -euo pipefail
for pop in AFR AMR CSA EAS EUR MID; do
~/plink2 \
  --pfile ~/ukb_imp_sbayesrc \
  --keep /mnt/project/mind_the_gap/phenotype/ukb${pop}.keep \
  --make-pgen \
  --out ~/ukb_imp_sbayesrc.${pop}
  
dx upload ~/ukb_imp_sbayesrc.${pop}.pgen --path mind_the_gap/genotype/
dx upload ~/ukb_imp_sbayesrc.${pop}.pvar --path mind_the_gap/genotype/
dx upload ~/ukb_imp_sbayesrc.${pop}.psam --path mind_the_gap/genotype/

rm ~/ukb_imp_sbayesrc.${pop}.*

done

# Do the EUR training subset
~/plink2 \
  --pfile ~/ukb_imp_sbayesrc \
  --keep /mnt/project/mind_the_gap/phenotype/ukb${pop}.keep \
  --make-pgen \
  --out ~/ukb_imp_sbayesrc.${pop}
  
dx upload ~/ukb_imp_sbayesrc.${pop}.pgen --path mind_the_gap/genotype/
dx upload ~/ukb_imp_sbayesrc.${pop}.pvar --path mind_the_gap/genotype/
dx upload ~/ukb_imp_sbayesrc.${pop}.psam --path mind_the_gap/genotype/

rm ~/ukb_imp_sbayesrc.${pop}.*

```

***

## WGS data

```{r}
library(data.table)
pvar <- NULL
for(i in 1:22){
  pvar <- rbind(
    pvar,
    fread(
      paste0(
        '/mnt/project/mind_the_gap/genotype/ukb_imp_sbayesrc.chr', i, '.pvar')
    )
  )
}

# Make bed file for liftover
bed_file<-data.table(
  chr=paste0('chr', pvar$`#CHROM`),
  bp_minus1=format(pvar$POS-1, scientific = FALSE),
  bp=format(pvar$POS, scientific = FALSE),
  rsid=pvar$ID
)

fwrite(bed_file, 'snplist.bed', col.names = F, quote = F, sep = ' ')
```

```{bash}
# Liftover
./liftOver snplist.bed hg19ToHg38.over.chain.gz sites.hg38.bed sites.unmapped.bed

# Convert back to range format for plink2 (bed1)
awk 'BEGIN{OFS="\t"} {print $1, $3, $3}' sites.hg38.bed > sites.hg38.txt

# Subset individuals and variants
seq 1 22 | xargs -n1 -P 6 -I{} bash -c '
  set -euo pipefail
  chr="{}"
  
  ~/plink2 \
    --pfile /mnt/project/Bulk/DRAGEN\ WGS/DRAGEN\ population\ level\ WGS\ variants\,\ PLINK\ format\ \[500k\ release\]/ukb24308_c${chr}_b0_v1 \
    --extract range sites.hg38.txt \
    --no-pheno \
    --threads 1 \
    --memory 15000 \
    --make-pgen \
    --out ~/ukb_wgs_sbayesrc.chr${chr}
  
  dx upload ukb_wgs_sbayesrc.chr${chr}.pgen --path mind_the_gap/wgs/
  dx upload ukb_wgs_sbayesrc.chr${chr}.pvar --path mind_the_gap/wgs/
  dx upload ukb_wgs_sbayesrc.chr${chr}.psam --path mind_the_gap/wgs/
  
  rm ukb_wgs_sbayesrc.chr${chr}.*
'
done

```

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json
```

```{r}
# We need to resolve multiallelic variants in plink files
# Change SNP IDs for variants that are not in the dense sbayesrc reference
library(data.table)
score_dense <- fread('/mnt/project/mind_the_gap/score_files/dense/sbayesrc/ref-BMI.score.gz')
score_sparse <- fread('/mnt/project/mind_the_gap/score_files/sparse/sbayesrc/ref-BMI.score.gz')

score <- merge(
  score_dense[, c('SNP', 'A1', 'A2'), with = F], 
  score_sparse[, c('SNP', 'A1', 'A2'), with = F], 
  by = 'SNP',
  all = T)

score$match <- with(score, ((A1.x == A1.y & A2.x == A2.y) | (A1.x == A2.y & A2.x == A1.y)))
score$match[is.na(score$match)] <- F
score$A1<-NA
score$A2<-NA
score$A1[score$match] <- score$A1.x[score$match]
score$A2[score$match] <- score$A2.x[score$match]
score$A1[is.na(score$A1.x)] <- score$A1.y[is.na(score$A1.x)]
score$A2[is.na(score$A1.x)] <- score$A2.y[is.na(score$A1.x)]
score$A1[is.na(score$A1.y)] <- score$A1.x[is.na(score$A1.y)]
score$A2[is.na(score$A1.y)] <- score$A2.x[is.na(score$A1.y)]

score <- score[, c('SNP', 'A1', 'A2'), with = F]

# Insert CHR and BP (GRCh38) into score file
bed <- fread('/mnt/project/mind_the_gap/sites.hg38.bed')

score_bed <- merge(score, bed, by.x = 'SNP', by.y = 'V4')
score_bed<-score_bed[!duplicated(score_bed$SNP),]
score_bed$V1 <- as.numeric(gsub('chr','', score_bed$V1))

for(i in 1:22){
  print(i)
  pvar <- fread(paste0('/mnt/project/mind_the_gap/wgs/ukb_wgs_sbayesrc.chr',i,'.pvar'))
  score_i <- score_bed[score_bed$V1 == i,]
  
  pvar_score_dense <- merge(
    pvar,
    score_i,
    by.x = c('#CHROM', 'POS'),
    by.y = c('V1', 'V3'),
    all.x = T,
    sort = F
  )
  
  pvar_score_dense$omit <- with(pvar_score_dense, !((REF == A1 & ALT == A2) | (REF == A2 & ALT == A1)))
  pvar_score_dense$omit[is.na(pvar_score_dense$omit)] <- T
  
  pvar_score_dense$ID[!pvar_score_dense$omit] <- pvar_score_dense$SNP[!pvar_score_dense$omit]

  fwrite(
    pvar_score_dense[, c('#CHROM', 'ID', 'POS', 'REF', 'ALT'), with = F], 
    paste0('~/ukb_wgs_sbayesrc.chr',i,'.pvar'),
    sep = ' ', quote = F)
}

```

```{bash}
# Make symlinks
for chr in $(seq 1 22); do
  for file in $(echo pgen psam); do
    ln -s /mnt/project/mind_the_gap/wgs/ukb_wgs_sbayesrc.chr${chr}.${file} ~/ukb_wgs_sbayesrc.chr${chr}.${file}
  done
done
```

```{bash}
# Split the data into populations, and EUR train and test and merge chromosomes
# Merge each chromosome and extract PGS SNP (removing multiallelic)
rm -f merge_list.txt
for i in $(seq 1 22); do
  echo ~/ukb_wgs_sbayesrc.chr${i} >> ~/merge_list.txt
done

~/plink2 \
  --pmerge-list ~/merge_list.txt \
  --extract /mnt/project/mind_the_gap/genotype/snplist.txt \
  --make-pgen \
  --out ~/ukb_wgs_sbayesrc

# Subset by population
for pop in AFR AMR CSA EAS EUR MID; do
  echo "$pop"
done | xargs -P 6 -I{} bash -c '
  set -euo pipefail
  POP="{}"
  ~/plink2 \
    --pfile ~/ukb_wgs_sbayesrc \
    --keep /mnt/project/mind_the_gap/phenotype/ukb.test.${POP}.keep \
    --make-pgen \
    --threads 2 \
    --memory 10000 \
    --out ~/ukb_wgs_sbayesrc.test.${POP}
    
  dx upload ~/ukb_wgs_sbayesrc.test.${POP}.pgen --path mind_the_gap/wgs/
  dx upload ~/ukb_wgs_sbayesrc.test.${POP}.pvar --path mind_the_gap/wgs/
  dx upload ~/ukb_wgs_sbayesrc.test.${POP}.psam --path mind_the_gap/wgs/
  
  rm ~/ukb_wgs_sbayesrc.test.${POP}.*
'

# EUR training dataset
pop=EUR
~/plink2 \
  --pfile ~/ukb_wgs_sbayesrc \
  --keep /mnt/project/mind_the_gap/phenotype/ukb.train.${pop}.keep \
  --make-pgen \
  --out ~/ukb_wgs_sbayesrc.train.${pop}
  
dx upload ~/ukb_wgs_sbayesrc.train.${pop}.pgen --path mind_the_gap/wgs/
dx upload ~/ukb_wgs_sbayesrc.train.${pop}.pvar --path mind_the_gap/wgs/
dx upload ~/ukb_wgs_sbayesrc.train.${pop}.psam --path mind_the_gap/wgs/

rm ~/ukb_wgs_sbayesrc.train.${pop}.*
```

***

# Run GWAS in UKB

***

## Imputed data

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json
```

```{r}
library(data.table)

# Combine the phenotypes into single file
traits=c('HT','HB','PLT','BMI','BWT','MCHC','SBP','NEU','HDL','TC')  

pheno_all <- list()
for(trait in traits){
  pheno_all[[trait]]<-fread(paste0('/mnt/project/mind_the_gap/phenotype/', trait, '.EUR.gwas.txt'))
  names(pheno_all[[trait]])[3] <- trait
}

pheno_all <- Reduce(function(dtf1, dtf2) merge(dtf1, dtf2, by = c('FID','IID'), all = TRUE), pheno_all)

fwrite(pheno_all, '~/all_trait.EUR.gwas.txt', sep=' ', na='NA', quote=F)

```

```{bash}
dx upload ~/all_trait.EUR.gwas.txt --path mind_the_gap/phenotype/
```

The code for the GWAS was submitted from CREATE server and is documented elsewhere.

***

## WGS data

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json
```

```{bash}

dx download mind_the_gap/wgs/ukb_wgs_sbayesrc.train.EUR.pgen
dx download mind_the_gap/wgs/ukb_wgs_sbayesrc.train.EUR.pvar
dx download mind_the_gap/wgs/ukb_wgs_sbayesrc.train.EUR.psam

for pheno in $(echo HT HB PLT BMI BWT MCHC SBP NEU HDL TC); do
  echo "$pheno"
done | xargs -P 5 -I{} bash -c '
pheno="{}"
mkdir -p gwas/${pheno}

./plink2 \
  --pfile ~/ukb_wgs_sbayesrc.train.EUR \
  --pheno /mnt/project/mind_the_gap/phenotype/${pheno}.EUR.gwas.txt \
  --linear omit-ref allow-no-covars cols=+a1freq,+ax \
  --threads 3 \
  --memory 15000 \
  --out gwas/${pheno}/ukb.eur_train.${pheno}

  # Remove REF and ALT columns and rename AX column to A2
  cut -f 4,5 --complement gwas/${pheno}/ukb.eur_train.${pheno}.outcome.glm.linear | awk '\''BEGIN{FS=OFS="\t"} NR==1 {$7="A2"} 1'\'' > gwas/${pheno}/temp.txt && mv gwas/${pheno}/temp.txt gwas/${pheno}/${pheno}.gw.txt

  gzip gwas/${pheno}/${pheno}.gw.txt
  dx upload gwas/${pheno}/${pheno}.gw.txt.gz --path mind_the_gap/gwas/
'
```

Export these sumstats to CREATE, where I can process using SBayesRC and ptclump using GenoPred.

***

# Group variants by imputation quality

```{r}
library(data.table)
library(foreach)
library(doMC)
registerDoMC(8)

# Subset EUR variants to remove certain proportions of variants based on INFO (0.5, 0.25, 0.05)
pop = 'EUR'
rsq <- foreach(i = 1:22, .combine = rbind, .options.multicore = list(preschedule = FALSE)) %dopar% {
  tmp <- fread(
    cmd = paste0(
      "grep -v '^#' /mnt/project/mind_the_gap/genotype/ukb_imp_sbayesrc_",
      pop,
      '.chr',
      i,
      '.snp-stats.txt'
    ),
    nThread = 1
  )
tmp
}

props <- c(0.5,0.25,0.05)
for(prop in props){
  print(prop)
  threshold <- quantile(rsq$info, probs = prop, na.rm = TRUE)
  print(threshold)
  rsq_tmp <- rsq[rsq$info > threshold,]
  fwrite(rsq_tmp[,'rsid', with = F, drop = F], paste0('~/snplist_rsq_subset.prop_', prop,'.', threshold, '.', pop,'.txt'), col.names = F, quote = F)
}
```

```{bash}
tar -cvf snplist_rsq_subset.prop.tar snplist_rsq_subset.prop_*
dx upload snplist_rsq_subset.prop.tar --path mind_the_gap/
```

***

# Select variants for random masking

Try removing 50%, 25%, 5% of variants from the PGS at random. Do 10 iterations to avoid randomness impacting results.

```{r}
library(data.table)

# List variants in WGS data
snps<-fread('/mnt/project/mind_the_gap/wgs/ukb_wgs_sbayesrc.test.EUR.pvar')$ID

# Create 10 iterations of random missingness
set.seed(1)
for(k in 1:10){
  for(i in c(0.05, 0.25, 0.5)){
    print(i)
    snps_keep<-sample(snps, floor((1-i)*length(snps)))
    write.table(snps_keep, paste0('~/snplist_random_subset.prop_',i,'.iteration_',k,'.txt'), col.names = F, row.names = F, quote = F)
  }
}

```

```{bash}
tar -cvf snplist_random_subset.prop.10_it.tar snplist_random_subset.prop_*
dx upload snplist_random_subset.prop.10_it.tar --path mind_the_gap/
```

***

# Quantify PGS decay

We will calculate the relative variance and R2 of masked PGS compared to baseline PGS, and the squared correlation between baseline and masked PGS. We will mask variants in PGS randomly (using 10 iterations), or based on INFO. We will also show the impact of reduced variance and R2 on PGS percentiles range, and slope of PGS relative to the outcome, and how rescaling can remedy the situation. We will also compare the observed and estimated relative variance and squared correlation between masked and baseline PGS.

***

## UKB WGS GWAS

### Random masking

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json

tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_random_masked_ukb_wgs_gwas.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# --- 1. CONFIGURATION ---

# Populations to analyze
# ORDER MATTERS: Run AFR first to get the sample size for downsampling EUR
populations <- c('AFR', 'EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_iterations <- 2          # Number of random masking iterations per trait/prop
n_boot_per_iter <- 2      # Bootstraps of INDIVIDUALS per iteration (10 * 100 = 1000 total)

# Variable to store AFR sample size dynamically
n_downsample_target <- NULL 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_iterations', 'n_boot_per_iter', 'n_downsample_target')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            n_full <- nrow(base_merged) # Capture the full sample size for this trait
            
            # --- ITERATION LOOP (Random Masking) ---
            # We will collect all 1000 bootstrap estimates (10 iters * 100 boots) for this trait/param
            boot_estimates_full <- list()
            boot_estimates_down <- list()
            
            for (iter in 1:n_iterations) {
                masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.random.prop_', prop, '.iteration_', iter, '.sscore')
                if (!file.exists(masked_file)) next
                
                pgs_masked <- fread(masked_file)
                if (!param_val %in% names(pgs_masked)) next
                
                pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
                full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
                
                # Define data subsets to analyze
                analysis_sets <- list(list(type="Full", data=full_data))
                
                if (pop == 'EUR') {
                     target_n <- if(!is.null(n_downsample_target)) n_downsample_target else 2000 
                     if (nrow(full_data) > target_n) {
                        down_idx <- sample(nrow(full_data), target_n)
                        analysis_sets[[2]] <- list(type="Downsampled", data=full_data[down_idx, ])
                     }
                }
                
                # --- INDIVIDUAL BOOTSTRAP LOOP ---
                for (aset in analysis_sets) {
                    df <- aset$data
                    n_curr <- nrow(df)
                    if (n_curr < 10) next 
                    
                    # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
                    # Subsetting a data.table 100,000 times is memory-intensive. 
                    # Indexing numeric vectors directly is vastly faster.
                    y_vec <- df[['outcome']]
                    s_vec <- df[[param_val]]
                    sp_vec <- df[[paste0(param_val, "_masked")]]
                    
                    # Bootstrap individuals n_boot_per_iter times
                    boot_stats <- replicate(n_boot_per_iter, {
                        idx <- sample(n_curr, n_curr, replace=TRUE)
                        
                        # Use vector indexing instead of data.table subsetting
                        y <- y_vec[idx]
                        s <- s_vec[idx]
                        s_prime <- sp_vec[idx]
                        
                        r2_s <- cor(y, s)^2
                        r2_s_prime <- cor(y, s_prime)^2
                        
                        # Variances
                        var_s <- var(s)
                        var_s_prime <- var(s_prime)
                        rel_var_obs <- var_s_prime / var_s
                        
                        # Slopes
                        slope_base <- cov(y, s) / var_s
                        slope_mask <- cov(y, s_prime) / var_s_prime
                        
                        # Handle extremely low baseline R2 safely
                        obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                        est_rel_r2 <- cor(s, s_prime)^2 
                        
                        error_val <- est_rel_r2 - obs_rel_r2
                        abs_error_val <- abs(error_val)
                        
                        # Rescaled Slope (Calibration Factor = sqrt(Relative R2 / Relative Variance))
                        cf <- sqrt(est_rel_r2 / rel_var_obs)
                        slope_rescaled <- if (!is.na(cf) && cf > 1e-8) slope_mask / cf else NA
                        
                        # Slope Ratios
                        slope_ratio_uncalib <- if (!is.na(slope_base) && abs(slope_base) > 1e-8) slope_mask / slope_base else NA
                        slope_ratio_calib <- if (!is.na(slope_base) && abs(slope_base) > 1e-8 && !is.na(slope_rescaled)) slope_rescaled / slope_base else NA
                        
                        c(obs_rel_r2, est_rel_r2, error_val, abs_error_val,
                          var_s, var_s_prime, rel_var_obs,
                          slope_base, slope_mask, slope_rescaled,
                          slope_ratio_uncalib, slope_ratio_calib, r2_s, r2_s_prime)
                    })
                    
                    # Store bootstrap results
                    # boot_stats is a matrix of 12 rows x 100 cols
                    if (aset$type == "Full") {
                        boot_estimates_full[[iter]] <- t(boot_stats)
                    } else {
                        boot_estimates_down[[iter]] <- t(boot_stats)
                    }
                } 
            } # End Iteration Loop
            
            # --- SUMMARIZE ALL 1000 BOOTSTRAPS FOR THIS TRAIT ---
            process_bootstraps <- function(boot_list, type_label, n_samples_val) {
                if (length(boot_list) == 0) return(NULL)
                
                # Combine all 100 matrices into one 1000x12 matrix
                all_boots <- do.call(rbind, boot_list)
                all_boots <- all_boots[complete.cases(all_boots), , drop=FALSE] # Remove NAs from low R2
                
                if (nrow(all_boots) == 0) return(NULL)
                
                means <- colMeans(all_boots)
                cis <- apply(all_boots, 2, quantile, probs=c(0.025, 0.975))
                
                data.table(
                    population = pop,
                    trait = trait_val,
                    param = param_val,
                    pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                    prop_missing = prop,
                    sample_type = type_label,
                    n_samples = n_samples_val,
                    total_bootstraps = nrow(all_boots),
                    
                    obs_rel_r2 = means[1],
                    obs_rel_r2_low = cis[1, 1],
                    obs_rel_r2_high = cis[2, 1],

                    est_rel_r2 = means[2],
                    est_rel_r2_low = cis[1, 2],
                    est_rel_r2_high = cis[2, 2],
                    
                    error = means[3],
                    error_low = cis[1, 3],
                    error_high = cis[2, 3],
                    
                    mad = means[4],
                    mad_low = cis[1, 4],
                    mad_high = cis[2, 4],
                    
                    # Newly added metrics
                    mean_var_base = means[5],
                    var_base_low = cis[1, 5],
                    var_base_high = cis[2, 5],
                    
                    mean_var_mask = means[6],
                    var_mask_low = cis[1, 6],
                    var_mask_high = cis[2, 6],
                    
                    mean_rel_var_obs = means[7],
                    rel_var_obs_low = cis[1, 7],
                    rel_var_obs_high = cis[2, 7],
                    
                    mean_slope_base = means[8],
                    slope_base_low = cis[1, 8],
                    slope_base_high = cis[2, 8],
                    
                    mean_slope_mask = means[9],
                    slope_mask_low = cis[1, 9],
                    slope_mask_high = cis[2, 9],
                    
                    mean_slope_rescaled = means[10],
                    slope_rescaled_low = cis[1, 10],
                    slope_rescaled_high = cis[2, 10],
                    
                    mean_slope_ratio_uncalib = means[11],
                    slope_ratio_uncalib_low = cis[1, 11],
                    slope_ratio_uncalib_high = cis[2, 11],
                    
                    mean_slope_ratio_calib = means[12],
                    slope_ratio_calib_low = cis[1, 12],
                    slope_ratio_calib_high = cis[2, 12],
                    
                    mean_r2_base = means[13],
                    r2_base_low = cis[1, 13],
                    r2_base_high = cis[2, 13],
                    
                    mean_r2_mask = means[14],
                    r2_mask_low = cis[1, 14],
                    r2_mask_high = cis[2, 14]
                )
            }
            
            n_down <- if (pop == 'EUR' && !is.null(n_downsample_target)) n_downsample_target else NA
            
            res_full <- process_bootstraps(boot_estimates_full, "Full", n_full)
            res_down <- process_bootstraps(boot_estimates_down, "Downsampled", n_down)
            
            trait_res_list[[length(trait_res_list) + 1]] <- rbind(res_full, res_down)
        }
        return(rbindlist(trait_res_list))
    }
    
    # --- Capture Sample Size from AFR run ---
    if (pop == 'AFR' && !is.null(trait_results_list)) {
        # Estimate N from AFR to use for downsampling EUR
        n_downsample_target <- round(median(trait_results_list$n_samples, na.rm=TRUE))
        print(paste("Set downsampling target for EUR to:", n_downsample_target))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_sparse_results_10it_random.csv', row.names=F)
system('dx upload ~/wgs_sparse_results_10it_random.csv --path mind_the_gap/results/')

```

***

#### Illustrate change in PGS percentile

```{r}
library(data.table)
library(ggplot2)
library(cowplot) # For theme_half_open()
library(stringr) # For str_detect()
library(dplyr)   # For percent_rank()

# --- 1. Define Example Parameters ---
example_trait <- 'BMI'
example_pop <- 'EUR'
example_method <- 'SBayesRC' # Case-insensitive, 'sbayesrc' also works
example_iteration <- 1

# Define paths and proportions
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
props <- c(0.05, 0.25, 0.5)

# This list will hold the binned data for each 'prop'
plot_data_list <- list()

# Read in the baseline PGS
pgs_baseline <- fread(paste0(base_dir_baseline, 'ukb.', example_pop, '.sparse.sscore'))
pgs_baseline <- pgs_baseline[, 'sbayesrc_BMI_SCORE_SBayesRC_SUM', with = F]

for (prop in props){
  # Read in the masked PGS
  pgs_masked <- fread(paste0(base_dir_masked, 'ukb.', example_pop, '.sparse.random.prop_',prop,'.iteration_', example_iteration, '.sscore'))
  pgs_masked <- pgs_masked[, 'sbayesrc_BMI_SCORE_SBayesRC_SUM', with = F]

  # Estimate relative variance
  obs_v_rel <- var(pgs_masked) / var(pgs_baseline)
  print(obs_v_rel)
  
  names(pgs_baseline) <- "pgs_s"
  names(pgs_masked) <- "pgs_s_prime"

  percentile_df <- cbind(pgs_baseline, pgs_masked)
  
  # 1. Create the "True Percentile Rank" from the baseline score (s)
  percentile_df[, rank_true := percent_rank(pgs_s)]
  
  # 2. Create an Empirical CDF function from the *baseline* (s) distribution.
  ecdf_baseline <- ecdf(percentile_df$pgs_s)
  
  # 3. Create the corrected scores by re-scaling the variance
  percentile_df[, pgs_s_corrected_obs := pgs_s_prime * (1/sqrt(as.numeric(obs_v_rel)))]

  # 4. Apply the baseline CDF to all three scores
  percentile_df[, rank_naive_std := ecdf_baseline(pgs_s_prime)]
  percentile_df[, rank_corrected_obs_std := ecdf_baseline(pgs_s_corrected_obs)]
  
  # --- Bin the data for a smooth plot ---
  percentile_df[, true_percentile_bin := cut(rank_true,
                                           breaks = seq(0, 1, by = 0.02),
                                           labels = FALSE,
                                           include.lowest = TRUE)]
  
  # Summarize by bin
  binned_percentiles <- percentile_df[, .(
    mean_true_rank = mean(rank_true, na.rm = TRUE),
    mean_naive_rank = mean(rank_naive_std, na.rm = TRUE),
    mean_corrected_obs_rank = mean(rank_corrected_obs_std, na.rm = TRUE)
  ), by = true_percentile_bin]
  
  # Add metadata for plotting
  binned_percentiles$prop <- prop
  binned_percentiles$call_rate <- paste0('Call rate = ', 1 - prop)
  
  plot_data_list[[length(plot_data_list) + 1]] <- binned_percentiles
} # End of prop loop

#                                 sbayesrc_BMI_SCORE_SBayesRC_SUM
# sbayesrc_BMI_SCORE_SBayesRC_SUM                       0.9033025
#                                 sbayesrc_BMI_SCORE_SBayesRC_SUM
# sbayesrc_BMI_SCORE_SBayesRC_SUM                        0.603311
#                                 sbayesrc_BMI_SCORE_SBayesRC_SUM
# sbayesrc_BMI_SCORE_SBayesRC_SUM                       0.3234202

# --- 4. Combine and Create Plot ---
if (length(plot_data_list) == 0) {
  stop("No data was successfully processed. Cannot create plot.")
}

# Melt the data to long format for ggplot
plot_data_wide <- rbindlist(plot_data_list)
plot_data_long <- melt(plot_data_wide, 
                       id.vars = c("true_percentile_bin", "mean_true_rank", "prop", "call_rate"),
                       measure.vars = c("mean_naive_rank", "mean_corrected_obs_rank"),
                       variable.name = "model_type",
                       value.name = "mean_observed_rank")

# Clean up the labels to match your slope plot
plot_data_long[model_type == "mean_naive_rank", model_type := "Naive"]
plot_data_long[model_type == "mean_corrected_obs_rank", model_type := "Corrected"]

plot_data_long$model_type <- factor(plot_data_long$model_type, 
                                  levels = c("Naive", "Corrected"))

# Factor call_rate to ensure correct panel order
plot_data_long$call_rate <- factor(plot_data_long$call_rate, 
                             levels = c("Call rate = 0.5", "Call rate = 0.75", "Call rate = 0.95"))

# Create the plot
plot_percentile_compression <- ggplot(plot_data_long, 
                                      aes(x = mean_true_rank, y = mean_observed_rank, color = model_type)) +
  
  # The y=x "Truth" line
  geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed") +
  
  # Plot all three lines
  geom_line(size = 1) +
  
  facet_wrap(~call_rate) +
  
  labs(
    title = paste("Standardized Percentile Compression & Correction for", example_trait),
    subtitle = paste("Showing", example_pop, "population,", example_method, "method"),
    x = "True Percentile\n(from baseline PGS 's' distribution)",
    y = "Observed Percentile\n(Score mapped to baseline distribution)",
    color = "Model Type"
  ) +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  theme_half_open() +
  background_grid(major = 'xy') +
  panel_border() +
  theme(legend.position = "bottom")

png('~/percentile_sparse_wgs_with_estimated.png', res = 300, width = 3300, height = 1750, units = 'px')
plot_percentile_compression
dev.off()
system('dx upload ~/percentile_sparse_wgs_with_estimated.png --path mind_the_gap/results/')

```

***

#### Evaluate PGS-check estimates

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# --- 1. CONFIGURATION ---

# Populations to analyze
# ORDER MATTERS: Run AFR first to get the sample size for downsampling EUR
populations <- c('AFR', 'EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_iterations <- 2          # Number of random masking iterations per trait/prop
n_boot_per_iter <- 2      # Bootstraps of INDIVIDUALS per iteration (10 * 100 = 1000 total)

# Variable to store AFR sample size dynamically
n_downsample_target <- NULL 

set.seed(123)


#######
# 2. READ IN ESTIMATED METRICS FROM PGS-CHECK
#######

# Using SBayesRC reference
est_dat <- NULL
for(iteration in 1:n_iterations){
    for(i in props){
        file_path <- paste0('/mnt/project/mind_the_gap/score_files/est_rsq.wgs.sparse.random.', i, '.iteration_', iteration, '.txt')
        if(file.exists(file_path)) {
            tmp <- fread(file_path)
            tmp$prop_missing <- i
            tmp$iteration <- iteration
            est_dat <- rbind(est_dat, tmp)
        }
    }
}
est_dat <- est_dat[, c('population', 'beta_set','relative_variance','relative_R2','prop_missing', 'iteration'), with = F]
names(est_dat) <- c('population', 'param','est_rel_var', 'est_rel_r2', 'prop_missing', 'iteration')
est_dat$param <- paste0(est_dat$param, '_SUM')


# Using 1KG+HGDP reference
est_dat_external <- NULL
for(iteration in 1:n_iterations){
    for(i in props){
        file_path <- paste0('/mnt/project/mind_the_gap/score_files/est_rsq.wgs.sparse.random.', i, '.iteration_', iteration, '.external_ld.txt')
        if(file.exists(file_path)) {
            tmp <- fread(file_path)
            tmp$prop_missing <- i
            tmp$iteration <- iteration
            est_dat_external <- rbind(est_dat_external, tmp)
        }
    }
}
est_dat_external <- est_dat_external[, c('population', 'beta_set','relative_variance','relative_R2','prop_missing', 'iteration'), with = F]
names(est_dat_external) <- c('population', 'param','est_rel_var', 'est_rel_r2', 'prop_missing', 'iteration')
est_dat_external$param <- paste0(est_dat_external$param, '_SUM')


# Using LD-naive approach (No Variance Estimate)
est_dat_naive <- NULL
for(iteration in 1:n_iterations){
    for(i in props){
        file_path <- paste0('/mnt/project/mind_the_gap/score_files/est_rsq.wgs.sparse.random.', i, '.iteration_', iteration, '.ld_naive.txt')
        if(file.exists(file_path)) {
            tmp <- fread(file_path)
            tmp$prop_missing <- i
            tmp$iteration <- iteration
            est_dat_naive <- rbind(est_dat_naive, tmp)
        }
    }
}
est_dat_naive <- est_dat_naive[, c('population', 'beta_set','relative_R2','prop_missing', 'iteration'), with = F]
names(est_dat_naive) <- c('population', 'param', 'est_rel_var', 'prop_missing', 'iteration')
est_dat_naive$param <- paste0(est_dat_naive$param, '_SUM')

# --- 3. OUTER LOOP: POPULATIONS ---
final_results_list <- list()

for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 4. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_iterations', 'n_boot_per_iter', 'n_downsample_target', 'est_dat', 'est_dat_external', 'est_dat_naive')
    ) %dopar% {
        
        # Load Phenotype (kept to ensure identical sample sub-setting as previous scripts)
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            n_full <- nrow(base_merged)
            
            # --- ITERATION LOOP (Random Masking) ---
            boot_estimates_full <- list()
            boot_estimates_down <- list()
            
            for (iter in 1:n_iterations) {
                masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.random.prop_', prop, '.iteration_', iter, '.sscore')
                if (!file.exists(masked_file)) next
                
                pgs_masked <- fread(masked_file)
                if (!param_val %in% names(pgs_masked)) next
                
                pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
                full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
                
                # Fetch Analytical Estimates for this specific Iteration
                c_ref <- est_dat[population == pop & prop_missing == prop & param == param_val & iteration == iter]
                c_ext <- est_dat_external[population == pop & prop_missing == prop & param == param_val & iteration == iter]
                c_nav <- est_dat_naive[population == pop & prop_missing == prop & param == param_val & iteration == iter]
                
                a_r2_ref <- if(nrow(c_ref) > 0) c_ref$est_rel_r2[1] else NA
                a_var_ref <- if(nrow(c_ref) > 0) c_ref$est_rel_var[1] else NA
                
                a_r2_ext <- if(nrow(c_ext) > 0) c_ext$est_rel_r2[1] else NA
                a_var_ext <- if(nrow(c_ext) > 0) c_ext$est_rel_var[1] else NA
                
                a_var_nav <- if(nrow(c_nav) > 0) c_nav$est_rel_var[1] else NA
                
                # Define data subsets to analyze
                analysis_sets <- list(list(type="Full", data=full_data))
                
                if (pop == 'EUR') {
                     target_n <- if(!is.null(n_downsample_target)) n_downsample_target else 2000 
                     if (nrow(full_data) > target_n) {
                        down_idx <- sample(nrow(full_data), target_n)
                        analysis_sets[[2]] <- list(type="Downsampled", data=full_data[down_idx, ])
                     }
                }
                
                # --- INDIVIDUAL BOOTSTRAP LOOP ---
                for (aset in analysis_sets) {
                    df <- aset$data
                    n_curr <- nrow(df)
                    if (n_curr < 10) next 
                    
                    s_vec <- df[[param_val]]
                    sp_vec <- df[[paste0(param_val, "_masked")]]
                    
                    # Bootstrap individuals n_boot_per_iter times
                    boot_stats <- replicate(n_boot_per_iter, {
                        idx <- sample(n_curr, n_curr, replace=TRUE)
                        
                        s <- s_vec[idx]
                        s_prime <- sp_vec[idx]
                        
                        # The "Observed" metrics are the empirical correlations/variances in the target sample
                        obs_sq_cor <- if(sd(s)>0 && sd(s_prime)>0) cor(s, s_prime)^2 else NA
                        obs_rel_var <- if(var(s)>0) var(s_prime) / var(s) else NA
                        
                        # Calculate Errors (Analytical Estimate - Empirical Observation)
                        err_r2_ref <- a_r2_ref - obs_sq_cor
                        err_var_ref <- a_var_ref - obs_rel_var
                        
                        err_r2_ext <- a_r2_ext - obs_sq_cor
                        err_var_ext <- a_var_ext - obs_rel_var
                        
                        err_var_nav <- a_var_nav - obs_rel_var
                        
                        c(obs_sq_cor, obs_rel_var, 
                          err_r2_ref, err_var_ref, err_r2_ext, err_var_ext, err_var_nav,
                          abs(err_r2_ref), abs(err_var_ref), abs(err_r2_ext), abs(err_var_ext), abs(err_var_nav),
                          a_r2_ref, a_var_ref, a_r2_ext, a_var_ext, a_var_nav)
                    })
                    
                    if (aset$type == "Full") {
                        boot_estimates_full[[iter]] <- t(boot_stats)
                    } else {
                        boot_estimates_down[[iter]] <- t(boot_stats)
                    }
                } 
            } # End Iteration Loop
            
            # --- SUMMARIZE ALL 1000 BOOTSTRAPS FOR THIS TRAIT ---
            process_bootstraps <- function(boot_list, type_label, n_samples_val) {
                if (length(boot_list) == 0) return(NULL)
                
                all_boots <- do.call(rbind, boot_list)
                all_boots <- all_boots[complete.cases(all_boots[, 1:2]), , drop=FALSE]
                
                if (nrow(all_boots) == 0) return(NULL)
                
                means <- colMeans(all_boots, na.rm=TRUE)
                cis <- apply(all_boots, 2, quantile, probs=c(0.025, 0.975), na.rm=TRUE)
                
                data.table(
                    population = pop,
                    trait = trait_val,
                    param = param_val,
                    pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                    prop_missing = prop,
                    sample_type = type_label,
                    n_samples = n_samples_val,
                    total_bootstraps = nrow(all_boots),
                    
                    obs_sq_cor = means[1],
                    obs_sq_cor_low = cis[1, 1],
                    obs_sq_cor_high = cis[2, 1],
                    
                    obs_rel_var = means[2],
                    obs_rel_var_low = cis[1, 2],
                    obs_rel_var_high = cis[2, 2],
                    
                    err_r2_ref = means[3],
                    err_r2_ref_low = cis[1, 3],
                    err_r2_ref_high = cis[2, 3],
                    
                    err_var_ref = means[4],
                    err_var_ref_low = cis[1, 4],
                    err_var_ref_high = cis[2, 4],
                    
                    err_r2_ext = means[5],
                    err_r2_ext_low = cis[1, 5],
                    err_r2_ext_high = cis[2, 5],
                    
                    err_var_ext = means[6],
                    err_var_ext_low = cis[1, 6],
                    err_var_ext_high = cis[2, 6],
                    
                    err_var_nav = means[7],
                    err_var_nav_low = cis[1, 7],
                    err_var_nav_high = cis[2, 7],
                    
                    mad_r2_ref = means[8],
                    mad_r2_ref_low = cis[1, 8],
                    mad_r2_ref_high = cis[2, 8],
                    
                    mad_var_ref = means[9],
                    mad_var_ref_low = cis[1, 9],
                    mad_var_ref_high = cis[2, 9],
                    
                    mad_r2_ext = means[10],
                    mad_r2_ext_low = cis[1, 10],
                    mad_r2_ext_high = cis[2, 10],
                    
                    mad_var_ext = means[11],
                    mad_var_ext_low = cis[1, 11],
                    mad_var_ext_high = cis[2, 11],
                    
                    mad_var_nav = means[12],
                    mad_var_nav_low = cis[1, 12],
                    mad_var_nav_high = cis[2, 12],
                    
                                        est_r2_ref = means[13],
                    est_r2_ref_low = cis[1, 13],
                    est_r2_ref_high = cis[2, 13],
                    
                    est_var_ref = means[14],
                    est_var_ref_low = cis[1, 14],
                    est_var_ref_high = cis[2, 14],
                    
                    est_r2_ext = means[15],
                    est_r2_ext_low = cis[1, 15],
                    est_r2_ext_high = cis[2, 15],
                    
                    est_var_ext = means[16],
                    est_var_ext_low = cis[1, 16],
                    est_var_ext_high = cis[2, 16],
                    
                    est_var_nav = means[17],
                    est_var_nav_low = cis[1, 17],
                    est_var_nav_high = cis[2, 17]
                )
            }
            
            n_down <- if (pop == 'EUR' && !is.null(n_downsample_target)) n_downsample_target else NA
            
            res_full <- process_bootstraps(boot_estimates_full, "Full", n_full)
            res_down <- process_bootstraps(boot_estimates_down, "Downsampled", n_down)
            
            trait_res_list[[length(trait_res_list) + 1]] <- rbind(res_full, res_down)
        }
        return(rbindlist(trait_res_list))
    }
    
    # Capture Sample Size from AFR run
    if (pop == 'AFR' && !is.null(trait_results_list)) {
        n_downsample_target <- round(median(trait_results_list$n_samples, na.rm=TRUE))
        print(paste("Set downsampling target for EUR to:", n_downsample_target))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_sparse_results_10it_random.pgs_check.csv', row.names=F)
system('dx upload ~/wgs_sparse_results_10it_random.pgs_check.csv --path mind_the_gap/results/')

```


#### Compare LD-aware and LD-naive methods

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(dplyr)

# --- 1. Data Processing Loop ---
# Define analysis parameters
props <- c(0.05, 0.25, 0.5)

# Read in the LD naive estimated relative variance
est_dat <- NULL
for(i in props){
  tmp <- fread(paste0('/mnt/project/mind_the_gap/score_files/est_rsq.wgs.sparse.random.',i,'.ld_naive.txt'))
  tmp$prop <- i
  est_dat <- rbind(est_dat, tmp)
}
est_dat<-est_dat[, c('population', 'beta_set','relative_R2','prop'), with = F]
names(est_dat)<-c('population', 'param','est_rel_var_naive', 'prop')
est_dat$param<-paste0(est_dat$param, '_SUM')

# Read in previous results
results_dt <- fread('/mnt/project/mind_the_gap/results/wgs_random_sparse_results.csv')
results_dt <- results_dt[, c('param','trait','prop','obs_rel_var','est_rel_var','pgs_method','population','call_rate'), with = F]

results_dt<-results_dt[results_dt$pgs_method == 'SBayesRC' | (results_dt$pgs_method == 'pT+clump' & grepl('0_1_SUM', results_dt$param)),]

both <- merge(results_dt, est_dat, by = c('param','population','prop'))
dt_melted <- melt(
  both,
  id.vars = c("trait", "population","pgs_method", "obs_rel_var","call_rate"),   # or whatever IDs you need
  measure.vars = c("est_rel_var", "est_rel_var_naive"),
  variable.name = "variance_type",
  value.name   = "relative_variance"
)
dt_melted$variance_type <- factor(
  dt_melted$variance_type,
  levels = c('est_rel_var', 'est_rel_var_naive'),
  labels = c('LD-aware', 'LD-naive')
)

# Compare estimated and observed relative variance
metrics <- dt_melted %>%
  group_by(pgs_method, variance_type) %>%
  summarise(
    pearson_r = coef(lm(scale(obs_rel_var)  ~ scale(relative_variance)))[2],
    pearson_r_ci_low = confint(lm(scale(obs_rel_var)  ~ scale(relative_variance)))[2,1],
    pearson_r_ci_high = confint(lm(scale(obs_rel_var)  ~ scale(relative_variance)))[2,2],
    slope = coef(lm(obs_rel_var  ~ relative_variance))[2],
    slope_ci_low = confint(lm(obs_rel_var  ~ relative_variance))[2,1],
    slope_ci_high = confint(lm(obs_rel_var  ~ relative_variance))[2,2],
    .groups = "drop"
  )

metrics$text <- paste0("r = ", round(metrics$pearson_r, 2),
                       " [", round(metrics$pearson_r_ci_low, 2), ", ", round(metrics$pearson_r_ci_high, 2), "]",
                       "\nβ = ", round(metrics$slope, 2),
                       " [", round(metrics$slope_ci_low, 2), ", ", round(metrics$slope_ci_high, 2), "]")

png('obs_est_rel_var_ld_naive_aware_comparison.png',
    units = 'px',
    width = 2750,
    height = 2500,
    res = 300)

ggplot(dt_melted, aes(x = relative_variance, y = obs_rel_var, fill = call_rate)) +
  geom_abline(colour = 'red') +
  geom_point(shape = 21) +
  facet_grid(pgs_method ~ variance_type) +
  coord_fixed() +
  theme_half_open() +
  panel_border() +
  background_grid() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0,1)) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0,1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(x = "Estimated relative variance", y = 'Observed relative variance', fill = NULL) +
  geom_text(data = metrics, aes(x = 0.3, y = 0.95, label = text), inherit.aes = FALSE, size = 4)

dev.off()
system('dx upload ~/obs_est_rel_var_ld_naive_aware_comparison.png --path mind_the_gap/results/')

```

***

### INFO-based masking

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_ukb_wgs_gwas.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# Populations to analyze
populations <- c('EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_boot <- 10              # Bootstraps of INDIVIDUALS 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_boot')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            
            # --- INFO MASKING (Single File, Deterministic) ---
            masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.prop_', prop, '.sscore')
            if (!file.exists(masked_file)) next
            
            pgs_masked <- fread(masked_file)
            if (!param_val %in% names(pgs_masked)) next
            
            pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
            full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
            
            n_curr <- nrow(full_data)
            if (n_curr < 10) next 
            
            # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
            y_vec <- full_data[['outcome']]
            s_vec <- full_data[[param_val]]
            sp_vec <- full_data[[paste0(param_val, "_masked")]]
            
            # --- INDIVIDUAL BOOTSTRAP LOOP ---
            boot_stats <- replicate(n_boot, {
                idx <- sample(n_curr, n_curr, replace=TRUE)
                
                # Use vector indexing instead of data.table subsetting
                y <- y_vec[idx]
                s <- s_vec[idx]
                s_prime <- sp_vec[idx]
                
                r2_s <- cor(y, s)^2
                r2_s_prime <- cor(y, s_prime)^2
                
                var_s <- var(s)
                var_s_prime <- var(s_prime)
                rel_var_obs <- var_s_prime / var_s
                
                # Handle extremely low baseline R2 safely
                obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                est_rel_r2 <- cor(s, s_prime)^2 
                
                error_val <- est_rel_r2 - obs_rel_r2
                abs_error_val <- abs(error_val)
                
                c(obs_rel_r2, est_rel_r2, error_val, abs_error_val, r2_s, r2_s_prime, var_s, var_s_prime, rel_var_obs)
            })
            
            # Transpose and remove failures (NAs)
            boot_stats <- t(boot_stats)
            boot_stats <- boot_stats[complete.cases(boot_stats), , drop=FALSE]
            
            if (nrow(boot_stats) == 0) next
            
            # Summarize the distribution
            means <- colMeans(boot_stats)
            cis <- apply(boot_stats, 2, quantile, probs=c(0.025, 0.975))
            
            trait_res_list[[length(trait_res_list) + 1]] <- data.table(
                population = pop,
                trait = trait_val,
                param = param_val,
                pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                prop_missing = prop,
                n_samples = n_curr,
                total_bootstraps = nrow(boot_stats),
                
                obs_rel_r2 = means[1],
                obs_rel_r2_low = cis[1, 1],
                obs_rel_r2_high = cis[2, 1],

                est_rel_r2 = means[2],
                est_rel_r2_low = cis[1, 2],
                est_rel_r2_high = cis[2, 2],
                
                error = means[3],
                error_low = cis[1, 3],
                error_high = cis[2, 3],
                
                mad = means[4],
                mad_low = cis[1, 4],
                mad_high = cis[2, 4],
                
                mean_r2_base = means[5],
                r2_base_low = cis[1, 5],
                r2_base_high = cis[2, 5],
                
                mean_r2_mask = means[6],
                r2_mask_low = cis[1, 6],
                r2_mask_high = cis[2, 6],
                
                mean_var_base = means[7],
                var_base_low = cis[1, 7],
                var_base_high = cis[2, 7],
                
                mean_var_mask = means[8],
                var_mask_low = cis[1, 8],
                var_mask_high = cis[2, 8],
                
                mean_rel_var_obs = means[9],
                rel_var_obs_low = cis[1, 9],
                rel_var_obs_high = cis[2, 9]
            )
        }
        return(rbindlist(trait_res_list))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_sparse_results.csv', row.names=F)
system('dx upload ~/wgs_sparse_results.csv --path mind_the_gap/results/')

```

***

#### 1KG+HGDP reference for SBayesRC

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_wgs_gwas_1kghgdp_ref.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_ukb_wgs_gwas_1kghgdp_ref.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# Populations to analyze
populations <- c('EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_boot <- 10              # Bootstraps of INDIVIDUALS 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.1kghgdp_ref.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_boot')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            
            # --- INFO MASKING (Single File, Deterministic) ---
            masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.1kghgdp_ref.prop_', prop, '.sscore')
            if (!file.exists(masked_file)) next
            
            pgs_masked <- fread(masked_file)
            if (!param_val %in% names(pgs_masked)) next
            
            pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
            full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
            
            n_curr <- nrow(full_data)
            if (n_curr < 10) next 
            
            # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
            y_vec <- full_data[['outcome']]
            s_vec <- full_data[[param_val]]
            sp_vec <- full_data[[paste0(param_val, "_masked")]]
            
            # --- INDIVIDUAL BOOTSTRAP LOOP ---
            boot_stats <- replicate(n_boot, {
                idx <- sample(n_curr, n_curr, replace=TRUE)
                
                # Use vector indexing instead of data.table subsetting
                y <- y_vec[idx]
                s <- s_vec[idx]
                s_prime <- sp_vec[idx]
                
                r2_s <- cor(y, s)^2
                r2_s_prime <- cor(y, s_prime)^2
                
                var_s <- var(s)
                var_s_prime <- var(s_prime)
                rel_var_obs <- var_s_prime / var_s
                
                # Handle extremely low baseline R2 safely
                obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                est_rel_r2 <- cor(s, s_prime)^2 
                
                error_val <- est_rel_r2 - obs_rel_r2
                abs_error_val <- abs(error_val)
                
                c(obs_rel_r2, est_rel_r2, error_val, abs_error_val, r2_s, r2_s_prime, var_s, var_s_prime, rel_var_obs)
            })
            
            # Transpose and remove failures (NAs)
            boot_stats <- t(boot_stats)
            boot_stats <- boot_stats[complete.cases(boot_stats), , drop=FALSE]
            
            if (nrow(boot_stats) == 0) next
            
            # Summarize the distribution
            means <- colMeans(boot_stats)
            cis <- apply(boot_stats, 2, quantile, probs=c(0.025, 0.975))
            
            trait_res_list[[length(trait_res_list) + 1]] <- data.table(
                population = pop,
                trait = trait_val,
                param = param_val,
                pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                prop_missing = prop,
                n_samples = n_curr,
                total_bootstraps = nrow(boot_stats),
                
                obs_rel_r2 = means[1],
                obs_rel_r2_low = cis[1, 1],
                obs_rel_r2_high = cis[2, 1],

                est_rel_r2 = means[2],
                est_rel_r2_low = cis[1, 2],
                est_rel_r2_high = cis[2, 2],
                
                error = means[3],
                error_low = cis[1, 3],
                error_high = cis[2, 3],
                
                mad = means[4],
                mad_low = cis[1, 4],
                mad_high = cis[2, 4],
                
                mean_r2_base = means[5],
                r2_base_low = cis[1, 5],
                r2_base_high = cis[2, 5],
                
                mean_r2_mask = means[6],
                r2_mask_low = cis[1, 6],
                r2_mask_high = cis[2, 6],
                
                mean_var_base = means[7],
                var_base_low = cis[1, 7],
                var_base_high = cis[2, 7],
                
                mean_var_mask = means[8],
                var_mask_low = cis[1, 8],
                var_mask_high = cis[2, 8],
                
                mean_rel_var_obs = means[9],
                rel_var_obs_low = cis[1, 9],
                rel_var_obs_high = cis[2, 9]
            )
        }
        return(rbindlist(trait_res_list))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_sparse_1kghgdp_ref_results.csv', row.names=F)
system('dx upload ~/wgs_sparse_1kghgdp_ref_results.csv --path mind_the_gap/results/')

```

***

### Evaluate recovered PGS

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json

tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_recovered_ukb_wgs_gwas.tar

```

#### GCTB

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# --- 1. CONFIGURATION & DATA LOADING ---

# Define analysis parameters
populations <- c('EUR')
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'
props <- c(0.05, 0.25, 0.5)

n_boot <- 5 # Number of bootstrap replicates (Increase this to 100 or 1000 for final run)
set.seed(123)  # For reproducible results

# Read in estimated relative variance and cor(s,s') early so we can pass it into the bootstrap
est_dat <- fread('/mnt/project/mind_the_gap/results/est_rsq.wgs.sparse.recovered.txt')
est_dat <- est_dat[, c('beta_set','var_baseline','var_imputed','relative_variance','cor2_baseline_imputed','prop'), with = F]
names(est_dat) <- c('beta_set','est_var_baseline','est_var_imputed','est_rel_var', 'cor2_baseline_imputed', 'prop_miss')
est_dat$beta_set <- paste0(est_dat$beta_set, '_SUM')

# Create the list to hold all results
final_results_list <- list()

# --- 2. RUN SEQUENTIAL OUTER LOOP (for pop) ---
for (pop in populations) {
  print(paste("Processing Population:", pop))
  
  # Read in baseline PGS
  baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
  pgs_baseline <- fread(baseline_file)
  
  # Read in PGS from recovered score file
  recovered_pgs_all <- fread('/home/rstudio-server/pgs_wgs/wgs_sparse/recovered/ukb.EUR.sparse.sscore')
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Read in masked PGS
    masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.prop_', prop, '.sscore')
    pgs_masked <- fread(masked_file)

    # Subset set recovered PGS to corresponding proportion of masking
    pgs_recovered <- recovered_pgs_all[, c('IID', names(recovered_pgs_all)[grepl(paste0('rsq_', prop, '_SUM'), names(recovered_pgs_all))]), with = F]
    names(pgs_recovered) <- gsub(paste0('_rsq_', prop), '', names(pgs_recovered))
    
    # Identify overlapping individuals
    pgs_baseline <- pgs_baseline[pgs_baseline$IID %in% pgs_masked$IID, ]
    pgs_masked <- pgs_masked[pgs_masked$IID %in% pgs_baseline$IID, ]
    pgs_recovered <- pgs_recovered[pgs_recovered$IID %in% pgs_masked$IID, ]

    # --- 3. RUN PARALLEL INNER LOOP (for trait) ---
    trait_results_list <- foreach(
        trait = traits,
        .combine = 'append', 
        .packages = 'data.table',
        .export = c('pheno_dir', 'n_boot', 'pgs_baseline', 'pgs_masked', 'pgs_recovered', 'pop', 'est_dat') 
    ) %dopar% {
        
        pheno_file <- paste0(pheno_dir, trait, '.', pop, '.txt')
        if (!file.exists(pheno_file)) {
            return(list()) 
        }
        
        pheno <- fread(pheno_file)

        # Subset PGS data
        pgs_baseline_i <- pgs_baseline[, c('IID', names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) &
                                                                      grepl(trait, names(pgs_baseline))]), with = F]
        pgs_masked_i <- pgs_masked[, c('IID', names(pgs_masked)[grepl('SUM', names(pgs_masked)) &
                                                                    grepl(trait, names(pgs_masked))]), with = F]
        pgs_recovered_i <- pgs_recovered[, c('IID', names(pgs_recovered)[grepl('SUM', names(pgs_recovered)) &
                                                                        grepl(trait, names(pgs_recovered))]), with = F]
        
        # Merge PGS data
        pgs_baseline_pheno <- merge(pheno, pgs_baseline_i, by = 'IID', sort = FALSE)
        pgs_masked_pheno <- merge(pheno, pgs_masked_i, by = 'IID', sort = FALSE)
        pgs_recovered_pheno <- merge(pheno, pgs_recovered_i, by = 'IID', sort = FALSE)
        
        # --- CRITICAL ALIGNMENT STEP ---
        # Ensure both data.tables have the same individuals in the same order
        common_iids <- Reduce(intersect, list(pgs_baseline_pheno$IID, 
                                              pgs_masked_pheno$IID, 
                                              pgs_recovered_pheno$IID))

        pgs_baseline_pheno <- pgs_baseline_pheno[IID %in% common_iids][order(IID)]
        pgs_masked_pheno <- pgs_masked_pheno[IID %in% common_iids][order(IID)]
        pgs_recovered_pheno <- pgs_recovered_pheno[IID %in% common_iids][order(IID)]

        # Restrict to common parameters
        common_params <- Reduce(intersect, list(names(pgs_baseline_i)[-1], 
                                                names(pgs_masked_i)[-1], 
                                                names(pgs_recovered_i)[-1]))

        pgs_baseline_pheno <- pgs_baseline_pheno[, c('IID', 'outcome', common_params), with = F]
        pgs_masked_pheno <- pgs_masked_pheno[, c('IID', 'outcome', common_params), with = F]
        pgs_recovered_pheno <- pgs_recovered_pheno[, c('IID', 'outcome', common_params), with = F]

        param_results_list <- list() 

        for (param in common_params) {
            if (!(param %in% names(pgs_masked_pheno))) next
            
            n <- nrow(pgs_baseline_pheno)
            if (n == 0) next 
            
            # Extract analytical estimates for this specific parameter and proportion
            c_est <- est_dat[est_dat$prop_miss == prop & est_dat$beta_set == param, ]
            a_rel_var_rec <- if(nrow(c_est) > 0) c_est$est_rel_var[1] else NA
            a_cor_sq_rec <- if(nrow(c_est) > 0) c_est$cor2_baseline_imputed[1] else NA
            
            # --- BOOTSTRAPPING SECTION ---
            boot_stats <- replicate(n_boot, {
                indices <- sample(1:n, n, replace = TRUE)
                
                boot_base <- pgs_baseline_pheno[indices]
                boot_masked <- pgs_masked_pheno[indices]
                boot_recovered <- pgs_recovered_pheno[indices]
                
                pheno_vals_boot <- boot_base[['outcome']]
                pgs_vals_base <- boot_base[[param]]
                pgs_vals_masked <- boot_masked[[param]]
                pgs_vals_recovered <- boot_recovered[[param]]
                
                # R2 stats
                r2_baseline_boot <- tryCatch(cor(pheno_vals_boot, pgs_vals_base)^2, error = function(e) NA)
                r2_masked_boot <- tryCatch(cor(pheno_vals_boot, pgs_vals_masked)^2, error = function(e) NA)
                r2_recovered_boot <- tryCatch(cor(pheno_vals_boot, pgs_vals_recovered)^2, error = function(e) NA)
                
                # Variances
                var_baseline_boot <- tryCatch(var(pgs_vals_base), error = function(e) NA)
                var_masked_boot <- tryCatch(var(pgs_vals_masked), error = function(e) NA)
                var_recovered_boot <- tryCatch(var(pgs_vals_recovered), error = function(e) NA)
                
                rel_var_masked_boot <- if (is.na(var_baseline_boot) || is.na(var_masked_boot) || var_baseline_boot == 0) NA else var_masked_boot / var_baseline_boot
                rel_var_recovered_boot <- if (is.na(var_baseline_boot) || is.na(var_recovered_boot) || var_baseline_boot == 0) NA else var_recovered_boot / var_baseline_boot
                
                # Correlations
                cor_baseline_masked_boot <- tryCatch(cor(pgs_vals_base, pgs_vals_masked), error = function(e) NA)
                cor_baseline_masked_sq_boot <- cor_baseline_masked_boot^2

                cor_baseline_recovered_boot <- tryCatch(cor(pgs_vals_base, pgs_vals_recovered), error = function(e) NA)
                cor_baseline_recovered_sq_boot <- cor_baseline_recovered_boot^2
                
                # Relative R2
                rel_r2_masked_boot <- if (is.na(r2_baseline_boot) || is.na(r2_masked_boot) || r2_baseline_boot == 0) NA else r2_masked_boot / r2_baseline_boot
                rel_r2_recovered_boot <- if (is.na(r2_baseline_boot) || is.na(r2_recovered_boot) || r2_baseline_boot == 0) NA else r2_recovered_boot / r2_baseline_boot
                
                # Evaluation Metrics for Recovered PGS
                err_rel_var_rec <- a_rel_var_rec - rel_var_recovered_boot
                abs_err_rel_var_rec <- abs(err_rel_var_rec)
                
                err_cor_sq_rec <- a_cor_sq_rec - cor_baseline_recovered_sq_boot
                abs_err_cor_sq_rec <- abs(err_cor_sq_rec)
                
                err_assump_rec <- cor_baseline_recovered_sq_boot - rel_r2_recovered_boot
                abs_err_assump_rec <- abs(err_assump_rec)
                
                err_est_relR2_rec <- a_cor_sq_rec - rel_r2_recovered_boot
                abs_err_est_relR2_rec <- abs(err_est_relR2_rec)

                # Return a named vector
                c(
                    r2_baseline = r2_baseline_boot,
                    r2_masked = r2_masked_boot,
                    r2_recovered = r2_recovered_boot,
                    rel_var_masked = rel_var_masked_boot,
                    rel_var_recovered = rel_var_recovered_boot,
                    cor_baseline_masked = cor_baseline_masked_boot,
                    cor_baseline_masked_sq = cor_baseline_masked_sq_boot,
                    cor_baseline_recovered = cor_baseline_recovered_boot,
                    cor_baseline_recovered_sq = cor_baseline_recovered_sq_boot,
                    rel_r2_masked = rel_r2_masked_boot,
                    rel_r2_recovered = rel_r2_recovered_boot,
                    err_rel_var_rec = err_rel_var_rec,
                    abs_err_rel_var_rec = abs_err_rel_var_rec,
                    err_cor_sq_rec = err_cor_sq_rec,
                    abs_err_cor_sq_rec = abs_err_cor_sq_rec,
                    err_assump_rec = err_assump_rec,
                    abs_err_assump_rec = abs_err_assump_rec,
                    err_est_relR2_rec = err_est_relR2_rec,
                    abs_err_est_relR2_rec = abs_err_est_relR2_rec
                )
            })
            
            # Transpose boot_stats
            boot_stats_t <- t(boot_stats)
            
            # --- Calculate 95% CIs and Means ---
            safe_quantile <- function(col_name) {
                if (!col_name %in% colnames(boot_stats_t) || all(is.na(boot_stats_t[, col_name]))) {
                    return(c(NA, NA))
                }
                quantile(boot_stats_t[, col_name], c(0.025, 0.975), na.rm = TRUE)
            }
            safe_mean <- function(col_name) {
                if (!col_name %in% colnames(boot_stats_t) || all(is.na(boot_stats_t[, col_name]))) {
                    return(NA)
                }
                mean(boot_stats_t[, col_name], na.rm = TRUE)
            }
            
            r2_baseline_ci <- safe_quantile("r2_baseline")
            r2_masked_ci <- safe_quantile("r2_masked")
            r2_recovered_ci <- safe_quantile("r2_recovered")
            rel_var_masked_ci <- safe_quantile("rel_var_masked")
            rel_var_recovered_ci <- safe_quantile("rel_var_recovered")
            cor_baseline_masked_ci <- safe_quantile("cor_baseline_masked")
            cor_baseline_masked_sq_ci <- safe_quantile("cor_baseline_masked_sq")
            cor_baseline_recovered_ci <- safe_quantile("cor_baseline_recovered")
            cor_baseline_recovered_sq_ci <- safe_quantile("cor_baseline_recovered_sq")
            rel_r2_masked_ci <- safe_quantile("rel_r2_masked")
            rel_r2_recovered_ci <- safe_quantile("rel_r2_recovered")
            
            err_rel_var_rec_ci <- safe_quantile("err_rel_var_rec")
            abs_err_rel_var_rec_ci <- safe_quantile("abs_err_rel_var_rec")
            err_cor_sq_rec_ci <- safe_quantile("err_cor_sq_rec")
            abs_err_cor_sq_rec_ci <- safe_quantile("abs_err_cor_sq_rec")
            
            err_assump_rec_ci <- safe_quantile("err_assump_rec")
            abs_err_assump_rec_ci <- safe_quantile("abs_err_assump_rec")
            err_est_relR2_rec_ci <- safe_quantile("err_est_relR2_rec")
            abs_err_est_relR2_rec_ci <- safe_quantile("abs_err_est_relR2_rec")

            # --- Calculate Point Estimates ---
            pheno_point <- pgs_baseline_pheno[['outcome']]
            pgs_s_point <- pgs_baseline_pheno[[param]]
            pgs_s_prime_point <- pgs_masked_pheno[[param]]
            pgs_s_rec_point <- pgs_recovered_pheno[[param]]
            
            r2_baseline_pheno <- cor(pheno_point, pgs_s_point)^2
            r2_masked_pheno <- cor(pheno_point, pgs_s_prime_point)^2
            r2_recovered_pheno <- cor(pheno_point, pgs_s_rec_point)^2
            p_val <- coef(summary(lm(pheno_point ~ pgs_s_point)))[2,4]
            
            cor_baseline_masked <- cor(pgs_s_point, pgs_s_prime_point)
            cor_baseline_masked_sq <- cor_baseline_masked^2
            cor_baseline_recovered <- cor(pgs_s_point, pgs_s_rec_point)
            cor_baseline_recovered_sq <- cor_baseline_recovered^2
            
            var_baseline <- var(pgs_s_point)
            var_masked <- var(pgs_s_prime_point)
            var_recovered <- var(pgs_s_rec_point)
            rel_var_masked_point <- var_masked / var_baseline
            rel_var_recovered_point <- var_recovered / var_baseline

            err_assump_rec <- cor_baseline_recovered_sq - (r2_recovered_pheno / r2_baseline_pheno)
            err_est_relR2_rec <- a_cor_sq_rec - (r2_recovered_pheno / r2_baseline_pheno)

            # --- Store results ---
            param_results_list[[length(param_results_list) + 1]] <- data.table(
                param = param,
                trait = trait,
                population = pop,
                prop = prop,
                n = nrow(pgs_baseline_pheno),
                p_baseline_pheno = p_val,
                
                var_baseline = var_baseline,
                var_masked = var_masked,
                var_recovered = var_recovered,
                
                r2_baseline_pheno = r2_baseline_pheno,
                r2_baseline_low_ci = r2_baseline_ci[1],
                r2_baseline_high_ci = r2_baseline_ci[2],
                
                r2_masked_pheno = r2_masked_pheno,
                r2_masked_low_ci = r2_masked_ci[1],
                r2_masked_high_ci = r2_masked_ci[2],
                
                r2_recovered_pheno = r2_recovered_pheno,
                r2_recovered_low_ci = r2_recovered_ci[1],
                r2_recovered_high_ci = r2_recovered_ci[2],
                
                obs_rel_masked_r2 = r2_masked_pheno / r2_baseline_pheno,
                rel_r2_masked_low_ci = rel_r2_masked_ci[1],
                rel_r2_masked_high_ci = rel_r2_masked_ci[2],
                
                obs_rel_recovered_r2 = r2_recovered_pheno / r2_baseline_pheno,
                rel_r2_recovered_low_ci = rel_r2_recovered_ci[1],
                rel_r2_recovered_high_ci = rel_r2_recovered_ci[2],
                
                obs_rel_masked_var = rel_var_masked_point,
                rel_var_masked_low_ci = rel_var_masked_ci[1],
                rel_var_masked_high_ci = rel_var_masked_ci[2],
                
                obs_rel_recovered_var = rel_var_recovered_point,
                rel_var_recovered_low_ci = rel_var_recovered_ci[1],
                rel_var_recovered_high_ci = rel_var_recovered_ci[2],
                
                cor_baseline_masked = cor_baseline_masked,
                cor_baseline_masked_low_ci = cor_baseline_masked_ci[1],
                cor_baseline_masked_high_ci = cor_baseline_masked_ci[2],
                
                cor_baseline_masked_sq = cor_baseline_masked_sq,
                cor_baseline_masked_sq_low_ci = cor_baseline_masked_sq_ci[1],
                cor_baseline_masked_sq_high_ci = cor_baseline_masked_sq_ci[2],
                
                cor_baseline_recovered = cor_baseline_recovered,
                cor_baseline_recovered_low_ci = cor_baseline_recovered_ci[1],
                cor_baseline_recovered_high_ci = cor_baseline_recovered_ci[2],
                
                cor_baseline_recovered_sq = cor_baseline_recovered_sq,
                cor_baseline_recovered_sq_low_ci = cor_baseline_recovered_sq_ci[1],
                cor_baseline_recovered_sq_high_ci = cor_baseline_recovered_sq_ci[2],
                
                # --- New Validation Columns ---
                est_rel_var_rec = a_rel_var_rec,
                est_cor_sq_rec = a_cor_sq_rec,
                
                err_rel_var_rec = a_rel_var_rec - rel_var_recovered_point,
                err_rel_var_rec_low_ci = err_rel_var_rec_ci[1],
                err_rel_var_rec_high_ci = err_rel_var_rec_ci[2],
                
                mad_rel_var_rec = safe_mean("abs_err_rel_var_rec"),
                mad_rel_var_rec_low_ci = abs_err_rel_var_rec_ci[1],
                mad_rel_var_rec_high_ci = abs_err_rel_var_rec_ci[2],
                
                err_cor_sq_rec = a_cor_sq_rec - cor_baseline_recovered_sq,
                err_cor_sq_rec_low_ci = err_cor_sq_rec_ci[1],
                err_cor_sq_rec_high_ci = err_cor_sq_rec_ci[2],
                
                mad_cor_sq_rec = safe_mean("abs_err_cor_sq_rec"),
                mad_cor_sq_rec_low_ci = abs_err_cor_sq_rec_ci[1],
                mad_cor_sq_rec_high_ci = abs_err_cor_sq_rec_ci[2],
                
                err_assump_rec = err_assump_rec,
                err_assump_rec_low_ci = err_assump_rec_ci[1],
                err_assump_rec_high_ci = err_assump_rec_ci[2],
                
                mad_assump_rec = safe_mean("abs_err_assump_rec"),
                mad_assump_rec_low_ci = abs_err_assump_rec_ci[1],
                mad_assump_rec_high_ci = abs_err_assump_rec_ci[2],

                err_est_relR2_rec = err_est_relR2_rec,
                err_est_relR2_rec_low_ci = err_est_relR2_rec_ci[1],
                err_est_relR2_rec_high_ci = err_est_relR2_rec_ci[2],
                
                mad_est_relR2_rec = safe_mean("abs_err_est_relR2_rec"),
                mad_est_relR2_rec_low_ci = abs_err_est_relR2_rec_ci[1],
                mad_est_relR2_rec_high_ci = abs_err_est_relR2_rec_ci[2]

            )
        } # End of param loop
        
        # The parallel worker returns the list of data.tables for this trait
        return(param_results_list) 
        
    } # End of parallel trait loop

    # --- 4. COMBINE RESULTS (within sequential loop) ---
    if (length(trait_results_list) > 0) {
        final_results_list <- append(final_results_list, trait_results_list)
    }
  } # End of prop loop

} # End of sequential pop loop

results_dt <- rbindlist(final_results_list)
results_dt[, pgs_method := ifelse(str_detect(param, 'sbayesrc'), 'SBayesRC', 'pT+clump')]
results_dt$call_rate <- paste0('Call rate = ', 1 - results_dt$prop)

write.csv(results_dt, '~/wgs_sparse_recovered_results.csv', row.names=F)
system('dx upload ~/wgs_sparse_recovered_results.csv --path mind_the_gap/results/')

```

***

#### R implementation

##### Tuning

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json

tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_recovered_rimp_tuning_ukb_wgs_gwas.tar

```

```{r}

library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# --- 1. CONFIGURATION & DATA LOADING ---

# Define analysis parameters
populations <- c('EUR')
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# New Tuning Parameters
props <- c(0.5)
ridges <- c(0.1, 0.2, 0.4, 0.8, 1.2, 1.5)
refs <- c('ukb', 'external')

# Read in estimated relative variance and cor(s,s') for tuning grid
est_dat <- fread('/mnt/project/mind_the_gap/results/est_rsq.wgs.sparse.recovered.rimp.tuning.txt')
est_dat <- est_dat[, c('beta_set','relative_variance_imputed','relative_R2_imputed','prop', 'ridge', 'ref'), with = F]
names(est_dat) <- c('beta_set','est_rel_var', 'cor2_baseline_imputed', 'prop_miss', 'ridge', 'ref')
est_dat$beta_set <- paste0(est_dat$beta_set, '_SUM')

# Create the list to hold all results
final_results_list <- list()

# --- 2. RUN SEQUENTIAL OUTER LOOP (for pop) ---
for (pop in populations) {
  print(paste("Processing Population:", pop))
  
  # Read in baseline PGS
  baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
  pgs_baseline <- fread(baseline_file)
  
  # Read in PGS from recovered score file (contains all ridges/refs)
  recovered_pgs_all <- fread('/home/rstudio-server/pgs_wgs/wgs_sparse/recovered_rimp_tuning/ukb.EUR.sparse.sscore')
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Read in masked PGS
    masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.prop_', prop, '.sscore')
    pgs_masked <- fread(masked_file)
    
    # Identify overlapping individuals
    pgs_baseline <- pgs_baseline[pgs_baseline$IID %in% pgs_masked$IID, ]
    pgs_masked <- pgs_masked[pgs_masked$IID %in% pgs_baseline$IID, ]
    pgs_recovered <- recovered_pgs_all[recovered_pgs_all$IID %in% pgs_masked$IID, ]

    # --- 3. RUN PARALLEL INNER LOOP (for trait) ---
    trait_results_list <- foreach(
        trait = traits,
        .combine = 'append', 
        .packages = 'data.table',
        .export = c('pheno_dir', 'pgs_baseline', 'pgs_masked', 'pgs_recovered', 'pop', 'est_dat', 'ridges', 'refs') 
    ) %dopar% {
        
        pheno_file <- paste0(pheno_dir, trait, '.', pop, '.txt')
        if (!file.exists(pheno_file)) {
            return(list()) 
        }
        
        pheno <- fread(pheno_file)

        # Subset PGS data for this trait
        pgs_baseline_i <- pgs_baseline[, c('IID', names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) &
                                                                      grepl(trait, names(pgs_baseline))]), with = F]
        pgs_masked_i <- pgs_masked[, c('IID', names(pgs_masked)[grepl('SUM', names(pgs_masked)) &
                                                                    grepl(trait, names(pgs_masked))]), with = F]
        pgs_recovered_i <- pgs_recovered[, c('IID', names(pgs_recovered)[grepl('SUM', names(pgs_recovered)) &
                                                                        grepl(trait, names(pgs_recovered))]), with = F]
        
        # Merge PGS data with phenotype
        pgs_baseline_pheno <- merge(pheno, pgs_baseline_i, by = 'IID', sort = FALSE)
        pgs_masked_pheno <- merge(pheno, pgs_masked_i, by = 'IID', sort = FALSE)
        pgs_recovered_pheno <- merge(pheno, pgs_recovered_i, by = 'IID', sort = FALSE)
        
        # --- CRITICAL ALIGNMENT STEP ---
        common_iids <- Reduce(intersect, list(pgs_baseline_pheno$IID, 
                                              pgs_masked_pheno$IID, 
                                              pgs_recovered_pheno$IID))

        pgs_baseline_pheno <- pgs_baseline_pheno[IID %in% common_iids][order(IID)]
        pgs_masked_pheno <- pgs_masked_pheno[IID %in% common_iids][order(IID)]
        pgs_recovered_pheno <- pgs_recovered_pheno[IID %in% common_iids][order(IID)]

        # Get the common baseline parameters for this trait
        common_params <- names(pgs_baseline_i)[-1]
        
        param_results_list <- list() 

        for (param in common_params) {
            # Skip if missing in masked file
            if (!(param %in% names(pgs_masked_pheno))) next
            
            n <- nrow(pgs_baseline_pheno)
            if (n == 0) next 
            
            # Extract point estimates that don't change across tuning parameters
            pheno_point <- pgs_baseline_pheno[['outcome']]
            pgs_s_point <- pgs_baseline_pheno[[param]]
            pgs_s_prime_point <- pgs_masked_pheno[[param]]
            
            var_baseline <- var(pgs_s_point)
            var_masked <- var(pgs_s_prime_point)
            
            r2_baseline_pheno <- cor(pheno_point, pgs_s_point)^2
            r2_masked_pheno <- cor(pheno_point, pgs_s_prime_point)^2
            p_val <- coef(summary(lm(pheno_point ~ pgs_s_point)))[2,4]
            
            cor_baseline_masked <- cor(pgs_s_point, pgs_s_prime_point)
            cor_baseline_masked_sq <- cor_baseline_masked^2
            
            obs_rel_masked_var <- var_masked / var_baseline
            obs_rel_masked_r2 <- r2_masked_pheno / r2_baseline_pheno
            
            # --- Iterate over Tuning Parameters (Ridge & Ref) ---
            for (ref_val in refs) {
                for (ridge_val in ridges) {
                    
                    # Reconstruct exact column name
                    rec_col_name <- paste0(gsub('_SUM','', param), '_rsq_', prop, '_ridge_', ridge_val, '_ref_', ref_val,'_SUM')
                    
                    if (!(rec_col_name %in% names(pgs_recovered_pheno))) {
                        print(paste("Missing column in recovered scores:", rec_col_name))
                        next
                    }
                    
                    pgs_s_rec_point <- pgs_recovered_pheno[[rec_col_name]]
                    
                    # Extract analytical estimates for this specific combo
                    c_est <- est_dat[est_dat$prop_miss == prop & 
                                     est_dat$beta_set == param & 
                                     est_dat$ridge == ridge_val & 
                                     est_dat$ref == ref_val, ]
                    
                    a_rel_var_rec <- if(nrow(c_est) > 0) c_est$est_rel_var[1] else NA
                    a_cor_sq_rec <- if(nrow(c_est) > 0) c_est$cor2_baseline_imputed[1] else NA
                    
                    # Calculate Recovered Metrics
                    var_recovered <- var(pgs_s_rec_point)
                    r2_recovered_pheno <- cor(pheno_point, pgs_s_rec_point)^2
                    
                    cor_baseline_recovered <- cor(pgs_s_point, pgs_s_rec_point)
                    cor_baseline_recovered_sq <- cor_baseline_recovered^2
                    
                    obs_rel_recovered_var <- var_recovered / var_baseline
                    obs_rel_recovered_r2 <- r2_recovered_pheno / r2_baseline_pheno
                    
                    # Calculate Errors
                    err_rel_var_rec <- a_rel_var_rec - obs_rel_recovered_var
                    err_cor_sq_rec <- a_cor_sq_rec - cor_baseline_recovered_sq
                    
                    # New: Calibration Assumption Error
                    err_assump_rec <- cor_baseline_recovered_sq - obs_rel_recovered_r2
                    
                    # New: Estimated Rel R2 Error
                    err_est_relR2_rec <- a_cor_sq_rec - obs_rel_recovered_r2
                    
                    # --- Store results ---
                    param_results_list[[length(param_results_list) + 1]] <- data.table(
                        param = param,
                        trait = trait,
                        population = pop,
                        prop = prop,
                        ridge = ridge_val,
                        ref = ref_val,
                        n = n,
                        p_baseline_pheno = p_val,
                        
                        # Variance Metrics
                        var_baseline = var_baseline,
                        var_masked = var_masked,
                        var_recovered = var_recovered,
                        
                        # Phenotype R2 Metrics
                        r2_baseline_pheno = r2_baseline_pheno,
                        r2_masked_pheno = r2_masked_pheno,
                        r2_recovered_pheno = r2_recovered_pheno,
                        
                        # Relative Empirical Metrics
                        obs_rel_masked_r2 = obs_rel_masked_r2,
                        obs_rel_recovered_r2 = obs_rel_recovered_r2,
                        obs_rel_masked_var = obs_rel_masked_var,
                        obs_rel_recovered_var = obs_rel_recovered_var,
                        
                        # Cor(s, s') Metrics
                        cor_baseline_masked = cor_baseline_masked,
                        cor_baseline_masked_sq = cor_baseline_masked_sq,
                        cor_baseline_recovered = cor_baseline_recovered,
                        cor_baseline_recovered_sq = cor_baseline_recovered_sq,
                        
                        # Analytical Estimates
                        est_rel_var_rec = a_rel_var_rec,
                        est_cor_sq_rec = a_cor_sq_rec,
                        
                        # Errors
                        err_rel_var_rec = err_rel_var_rec,
                        abs_err_rel_var_rec = abs(err_rel_var_rec),
                        err_cor_sq_rec = err_cor_sq_rec,
                        abs_err_cor_sq_rec = abs(err_cor_sq_rec),
                        
                        # New: Assumption & Estimate Errors
                        err_assump_rec = err_assump_rec,
                        abs_err_assump_rec = abs(err_assump_rec),
                        err_est_relR2_rec = err_est_relR2_rec,
                        abs_err_est_relR2_rec = abs(err_est_relR2_rec)
                    )
                }
            }
        } # End of param loop
        
        # The parallel worker returns the list of data.tables for this trait
        return(param_results_list) 
        
    } # End of parallel trait loop

    # --- 4. COMBINE RESULTS ---
    if (length(trait_results_list) > 0) {
        final_results_list <- append(final_results_list, trait_results_list)
    }
  } # End of prop loop

} # End of sequential pop loop

results_dt <- rbindlist(final_results_list)
results_dt[, pgs_method := ifelse(str_detect(param, 'sbayesrc'), 'SBayesRC', 'pT+clump')]
results_dt$call_rate <- paste0('Call rate = ', 1 - results_dt$prop)

write.csv(results_dt, '~/wgs_sparse_recovered_results_rimp_tuning.csv', row.names=F)
system('dx upload ~/wgs_sparse_recovered_results_rimp_tuning.csv --path mind_the_gap/results/')

```

***

##### Selected Parameters

```{bash}
# refresh mount
sudo umount /mnt; sudo -E /home/dnanexus/dxfuse -readOnly /mnt /home/dnanexus/.dxfuse_manifest.json

tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_ukb_wgs_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_recovered_rimp_ukb_wgs_gwas.tar

```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# --- 1. CONFIGURATION & DATA LOADING ---

# Define analysis parameters
populations <- c('EUR')
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')
base_dir_baseline <- '~/pgs_wgs/wgs_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/wgs_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'
props <- c(0.05, 0.25, 0.5)

n_boot <- 5 # Number of bootstrap replicates (Increase this to 100 or 1000 for final run)
set.seed(123)  # For reproducible results

# Read in estimated relative variance and cor(s,s') early so we can pass it into the bootstrap
# Read in estimated relative variance and cor(s,s')
est_dat <- fread('/mnt/project/mind_the_gap/results/est_rsq.wgs.sparse.recovered.rimp.txt')
est_dat<-est_dat[, c('beta_set','relative_variance_imputed','relative_R2_imputed','prop'), with = F]
names(est_dat)<-c('beta_set','est_rel_var', 'cor2_baseline_imputed', 'prop_miss')
est_dat$beta_set<-paste0(est_dat$beta_set, '_SUM')

# Create the list to hold all results
final_results_list <- list()

# --- 2. RUN SEQUENTIAL OUTER LOOP (for pop) ---
for (pop in populations) {
  print(paste("Processing Population:", pop))
  
  # Read in baseline PGS
  baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
  pgs_baseline <- fread(baseline_file)
  
  # Read in PGS from recovered score file
  recovered_pgs_all <- fread('/home/rstudio-server/pgs_wgs/wgs_sparse/recovered_rimp/ukb.EUR.sparse.sscore')
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Read in masked PGS
    masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.prop_', prop, '.sscore')
    pgs_masked <- fread(masked_file)

    # Subset set recovered PGS to corresponding proportion of masking
    pgs_recovered <- recovered_pgs_all[, c('IID', names(recovered_pgs_all)[grepl(paste0('rsq_', prop, '_SUM'), names(recovered_pgs_all))]), with = F]
    names(pgs_recovered) <- gsub(paste0('_rsq_', prop), '', names(pgs_recovered))
    
    # Identify overlapping individuals
    pgs_baseline <- pgs_baseline[pgs_baseline$IID %in% pgs_masked$IID, ]
    pgs_masked <- pgs_masked[pgs_masked$IID %in% pgs_baseline$IID, ]
    pgs_recovered <- pgs_recovered[pgs_recovered$IID %in% pgs_masked$IID, ]

    # --- 3. RUN PARALLEL INNER LOOP (for trait) ---
    trait_results_list <- foreach(
        trait = traits,
        .combine = 'append', 
        .packages = 'data.table',
        .export = c('pheno_dir', 'n_boot', 'pgs_baseline', 'pgs_masked', 'pgs_recovered', 'pop', 'est_dat') 
    ) %dopar% {
        
        pheno_file <- paste0(pheno_dir, trait, '.', pop, '.txt')
        if (!file.exists(pheno_file)) {
            return(list()) 
        }
        
        pheno <- fread(pheno_file)

        # Subset PGS data
        pgs_baseline_i <- pgs_baseline[, c('IID', names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) &
                                                                      grepl(trait, names(pgs_baseline))]), with = F]
        pgs_masked_i <- pgs_masked[, c('IID', names(pgs_masked)[grepl('SUM', names(pgs_masked)) &
                                                                    grepl(trait, names(pgs_masked))]), with = F]
        pgs_recovered_i <- pgs_recovered[, c('IID', names(pgs_recovered)[grepl('SUM', names(pgs_recovered)) &
                                                                        grepl(trait, names(pgs_recovered))]), with = F]
        
        # Merge PGS data
        pgs_baseline_pheno <- merge(pheno, pgs_baseline_i, by = 'IID', sort = FALSE)
        pgs_masked_pheno <- merge(pheno, pgs_masked_i, by = 'IID', sort = FALSE)
        pgs_recovered_pheno <- merge(pheno, pgs_recovered_i, by = 'IID', sort = FALSE)
        
        # --- CRITICAL ALIGNMENT STEP ---
        # Ensure both data.tables have the same individuals in the same order
        common_iids <- Reduce(intersect, list(pgs_baseline_pheno$IID, 
                                              pgs_masked_pheno$IID, 
                                              pgs_recovered_pheno$IID))

        pgs_baseline_pheno <- pgs_baseline_pheno[IID %in% common_iids][order(IID)]
        pgs_masked_pheno <- pgs_masked_pheno[IID %in% common_iids][order(IID)]
        pgs_recovered_pheno <- pgs_recovered_pheno[IID %in% common_iids][order(IID)]

        # Restrict to common parameters
        common_params <- Reduce(intersect, list(names(pgs_baseline_i)[-1], 
                                                names(pgs_masked_i)[-1], 
                                                names(pgs_recovered_i)[-1]))

        pgs_baseline_pheno <- pgs_baseline_pheno[, c('IID', 'outcome', common_params), with = F]
        pgs_masked_pheno <- pgs_masked_pheno[, c('IID', 'outcome', common_params), with = F]
        pgs_recovered_pheno <- pgs_recovered_pheno[, c('IID', 'outcome', common_params), with = F]

        param_results_list <- list() 

        for (param in common_params) {
            if (!(param %in% names(pgs_masked_pheno))) next
            
            n <- nrow(pgs_baseline_pheno)
            if (n == 0) next 
            
            # Extract analytical estimates for this specific parameter and proportion
            c_est <- est_dat[est_dat$prop_miss == prop & est_dat$beta_set == param, ]
            a_rel_var_rec <- if(nrow(c_est) > 0) c_est$est_rel_var[1] else NA
            a_cor_sq_rec <- if(nrow(c_est) > 0) c_est$cor2_baseline_imputed[1] else NA
            
            # --- BOOTSTRAPPING SECTION ---
            boot_stats <- replicate(n_boot, {
                indices <- sample(1:n, n, replace = TRUE)
                
                boot_base <- pgs_baseline_pheno[indices]
                boot_masked <- pgs_masked_pheno[indices]
                boot_recovered <- pgs_recovered_pheno[indices]
                
                pheno_vals_boot <- boot_base[['outcome']]
                pgs_vals_base <- boot_base[[param]]
                pgs_vals_masked <- boot_masked[[param]]
                pgs_vals_recovered <- boot_recovered[[param]]
                
                # R2 stats
                r2_baseline_boot <- tryCatch(cor(pheno_vals_boot, pgs_vals_base)^2, error = function(e) NA)
                r2_masked_boot <- tryCatch(cor(pheno_vals_boot, pgs_vals_masked)^2, error = function(e) NA)
                r2_recovered_boot <- tryCatch(cor(pheno_vals_boot, pgs_vals_recovered)^2, error = function(e) NA)
                
                # Variances
                var_baseline_boot <- tryCatch(var(pgs_vals_base), error = function(e) NA)
                var_masked_boot <- tryCatch(var(pgs_vals_masked), error = function(e) NA)
                var_recovered_boot <- tryCatch(var(pgs_vals_recovered), error = function(e) NA)
                
                rel_var_masked_boot <- if (is.na(var_baseline_boot) || is.na(var_masked_boot) || var_baseline_boot == 0) NA else var_masked_boot / var_baseline_boot
                rel_var_recovered_boot <- if (is.na(var_baseline_boot) || is.na(var_recovered_boot) || var_baseline_boot == 0) NA else var_recovered_boot / var_baseline_boot
                
                # Correlations
                cor_baseline_masked_boot <- tryCatch(cor(pgs_vals_base, pgs_vals_masked), error = function(e) NA)
                cor_baseline_masked_sq_boot <- cor_baseline_masked_boot^2

                cor_baseline_recovered_boot <- tryCatch(cor(pgs_vals_base, pgs_vals_recovered), error = function(e) NA)
                cor_baseline_recovered_sq_boot <- cor_baseline_recovered_boot^2
                
                # Relative R2
                rel_r2_masked_boot <- if (is.na(r2_baseline_boot) || is.na(r2_masked_boot) || r2_baseline_boot == 0) NA else r2_masked_boot / r2_baseline_boot
                rel_r2_recovered_boot <- if (is.na(r2_baseline_boot) || is.na(r2_recovered_boot) || r2_baseline_boot == 0) NA else r2_recovered_boot / r2_baseline_boot
                
                # Evaluation Metrics for Recovered PGS
                err_rel_var_rec <- a_rel_var_rec - rel_var_recovered_boot
                abs_err_rel_var_rec <- abs(err_rel_var_rec)
                
                err_cor_sq_rec <- a_cor_sq_rec - cor_baseline_recovered_sq_boot
                abs_err_cor_sq_rec <- abs(err_cor_sq_rec)
                
                err_assump_rec <- cor_baseline_recovered_sq_boot - rel_r2_recovered_boot
                abs_err_assump_rec <- abs(err_assump_rec)
                
                err_est_relR2_rec <- a_cor_sq_rec - rel_r2_recovered_boot
                abs_err_est_relR2_rec <- abs(err_est_relR2_rec)

                # Return a named vector
                c(
                    r2_baseline = r2_baseline_boot,
                    r2_masked = r2_masked_boot,
                    r2_recovered = r2_recovered_boot,
                    rel_var_masked = rel_var_masked_boot,
                    rel_var_recovered = rel_var_recovered_boot,
                    cor_baseline_masked = cor_baseline_masked_boot,
                    cor_baseline_masked_sq = cor_baseline_masked_sq_boot,
                    cor_baseline_recovered = cor_baseline_recovered_boot,
                    cor_baseline_recovered_sq = cor_baseline_recovered_sq_boot,
                    rel_r2_masked = rel_r2_masked_boot,
                    rel_r2_recovered = rel_r2_recovered_boot,
                    err_rel_var_rec = err_rel_var_rec,
                    abs_err_rel_var_rec = abs_err_rel_var_rec,
                    err_cor_sq_rec = err_cor_sq_rec,
                    abs_err_cor_sq_rec = abs_err_cor_sq_rec,
                    err_assump_rec = err_assump_rec,
                    abs_err_assump_rec = abs_err_assump_rec,
                    err_est_relR2_rec = err_est_relR2_rec,
                    abs_err_est_relR2_rec = abs_err_est_relR2_rec
                )
            })
            
            # Transpose boot_stats
            boot_stats_t <- t(boot_stats)
            
            # --- Calculate 95% CIs and Means ---
            safe_quantile <- function(col_name) {
                if (!col_name %in% colnames(boot_stats_t) || all(is.na(boot_stats_t[, col_name]))) {
                    return(c(NA, NA))
                }
                quantile(boot_stats_t[, col_name], c(0.025, 0.975), na.rm = TRUE)
            }
            safe_mean <- function(col_name) {
                if (!col_name %in% colnames(boot_stats_t) || all(is.na(boot_stats_t[, col_name]))) {
                    return(NA)
                }
                mean(boot_stats_t[, col_name], na.rm = TRUE)
            }
            
            r2_baseline_ci <- safe_quantile("r2_baseline")
            r2_masked_ci <- safe_quantile("r2_masked")
            r2_recovered_ci <- safe_quantile("r2_recovered")
            rel_var_masked_ci <- safe_quantile("rel_var_masked")
            rel_var_recovered_ci <- safe_quantile("rel_var_recovered")
            cor_baseline_masked_ci <- safe_quantile("cor_baseline_masked")
            cor_baseline_masked_sq_ci <- safe_quantile("cor_baseline_masked_sq")
            cor_baseline_recovered_ci <- safe_quantile("cor_baseline_recovered")
            cor_baseline_recovered_sq_ci <- safe_quantile("cor_baseline_recovered_sq")
            rel_r2_masked_ci <- safe_quantile("rel_r2_masked")
            rel_r2_recovered_ci <- safe_quantile("rel_r2_recovered")
            
            err_rel_var_rec_ci <- safe_quantile("err_rel_var_rec")
            abs_err_rel_var_rec_ci <- safe_quantile("abs_err_rel_var_rec")
            err_cor_sq_rec_ci <- safe_quantile("err_cor_sq_rec")
            abs_err_cor_sq_rec_ci <- safe_quantile("abs_err_cor_sq_rec")
            
            err_assump_rec_ci <- safe_quantile("err_assump_rec")
            abs_err_assump_rec_ci <- safe_quantile("abs_err_assump_rec")
            err_est_relR2_rec_ci <- safe_quantile("err_est_relR2_rec")
            abs_err_est_relR2_rec_ci <- safe_quantile("abs_err_est_relR2_rec")

            # --- Calculate Point Estimates ---
            pheno_point <- pgs_baseline_pheno[['outcome']]
            pgs_s_point <- pgs_baseline_pheno[[param]]
            pgs_s_prime_point <- pgs_masked_pheno[[param]]
            pgs_s_rec_point <- pgs_recovered_pheno[[param]]
            
            r2_baseline_pheno <- cor(pheno_point, pgs_s_point)^2
            r2_masked_pheno <- cor(pheno_point, pgs_s_prime_point)^2
            r2_recovered_pheno <- cor(pheno_point, pgs_s_rec_point)^2
            p_val <- coef(summary(lm(pheno_point ~ pgs_s_point)))[2,4]
            
            cor_baseline_masked <- cor(pgs_s_point, pgs_s_prime_point)
            cor_baseline_masked_sq <- cor_baseline_masked^2
            cor_baseline_recovered <- cor(pgs_s_point, pgs_s_rec_point)
            cor_baseline_recovered_sq <- cor_baseline_recovered^2
            
            var_baseline <- var(pgs_s_point)
            var_masked <- var(pgs_s_prime_point)
            var_recovered <- var(pgs_s_rec_point)
            rel_var_masked_point <- var_masked / var_baseline
            rel_var_recovered_point <- var_recovered / var_baseline

            err_assump_rec <- cor_baseline_recovered_sq - (r2_recovered_pheno / r2_baseline_pheno)
            err_est_relR2_rec <- a_cor_sq_rec - (r2_recovered_pheno / r2_baseline_pheno)

            # --- Store results ---
            param_results_list[[length(param_results_list) + 1]] <- data.table(
                param = param,
                trait = trait,
                population = pop,
                prop = prop,
                n = nrow(pgs_baseline_pheno),
                p_baseline_pheno = p_val,
                
                var_baseline = var_baseline,
                var_masked = var_masked,
                var_recovered = var_recovered,
                
                r2_baseline_pheno = r2_baseline_pheno,
                r2_baseline_low_ci = r2_baseline_ci[1],
                r2_baseline_high_ci = r2_baseline_ci[2],
                
                r2_masked_pheno = r2_masked_pheno,
                r2_masked_low_ci = r2_masked_ci[1],
                r2_masked_high_ci = r2_masked_ci[2],
                
                r2_recovered_pheno = r2_recovered_pheno,
                r2_recovered_low_ci = r2_recovered_ci[1],
                r2_recovered_high_ci = r2_recovered_ci[2],
                
                obs_rel_masked_r2 = r2_masked_pheno / r2_baseline_pheno,
                rel_r2_masked_low_ci = rel_r2_masked_ci[1],
                rel_r2_masked_high_ci = rel_r2_masked_ci[2],
                
                obs_rel_recovered_r2 = r2_recovered_pheno / r2_baseline_pheno,
                rel_r2_recovered_low_ci = rel_r2_recovered_ci[1],
                rel_r2_recovered_high_ci = rel_r2_recovered_ci[2],
                
                obs_rel_masked_var = rel_var_masked_point,
                rel_var_masked_low_ci = rel_var_masked_ci[1],
                rel_var_masked_high_ci = rel_var_masked_ci[2],
                
                obs_rel_recovered_var = rel_var_recovered_point,
                rel_var_recovered_low_ci = rel_var_recovered_ci[1],
                rel_var_recovered_high_ci = rel_var_recovered_ci[2],
                
                cor_baseline_masked = cor_baseline_masked,
                cor_baseline_masked_low_ci = cor_baseline_masked_ci[1],
                cor_baseline_masked_high_ci = cor_baseline_masked_ci[2],
                
                cor_baseline_masked_sq = cor_baseline_masked_sq,
                cor_baseline_masked_sq_low_ci = cor_baseline_masked_sq_ci[1],
                cor_baseline_masked_sq_high_ci = cor_baseline_masked_sq_ci[2],
                
                cor_baseline_recovered = cor_baseline_recovered,
                cor_baseline_recovered_low_ci = cor_baseline_recovered_ci[1],
                cor_baseline_recovered_high_ci = cor_baseline_recovered_ci[2],
                
                cor_baseline_recovered_sq = cor_baseline_recovered_sq,
                cor_baseline_recovered_sq_low_ci = cor_baseline_recovered_sq_ci[1],
                cor_baseline_recovered_sq_high_ci = cor_baseline_recovered_sq_ci[2],
                
                # --- New Validation Columns ---
                est_rel_var_rec = a_rel_var_rec,
                est_cor_sq_rec = a_cor_sq_rec,
                
                err_rel_var_rec = a_rel_var_rec - rel_var_recovered_point,
                err_rel_var_rec_low_ci = err_rel_var_rec_ci[1],
                err_rel_var_rec_high_ci = err_rel_var_rec_ci[2],
                
                mad_rel_var_rec = safe_mean("abs_err_rel_var_rec"),
                mad_rel_var_rec_low_ci = abs_err_rel_var_rec_ci[1],
                mad_rel_var_rec_high_ci = abs_err_rel_var_rec_ci[2],
                
                err_cor_sq_rec = a_cor_sq_rec - cor_baseline_recovered_sq,
                err_cor_sq_rec_low_ci = err_cor_sq_rec_ci[1],
                err_cor_sq_rec_high_ci = err_cor_sq_rec_ci[2],
                
                mad_cor_sq_rec = safe_mean("abs_err_cor_sq_rec"),
                mad_cor_sq_rec_low_ci = abs_err_cor_sq_rec_ci[1],
                mad_cor_sq_rec_high_ci = abs_err_cor_sq_rec_ci[2],
                
                err_assump_rec = err_assump_rec,
                err_assump_rec_low_ci = err_assump_rec_ci[1],
                err_assump_rec_high_ci = err_assump_rec_ci[2],
                
                mad_assump_rec = safe_mean("abs_err_assump_rec"),
                mad_assump_rec_low_ci = abs_err_assump_rec_ci[1],
                mad_assump_rec_high_ci = abs_err_assump_rec_ci[2],

                err_est_relR2_rec = err_est_relR2_rec,
                err_est_relR2_rec_low_ci = err_est_relR2_rec_ci[1],
                err_est_relR2_rec_high_ci = err_est_relR2_rec_ci[2],
                
                mad_est_relR2_rec = safe_mean("abs_err_est_relR2_rec"),
                mad_est_relR2_rec_low_ci = abs_err_est_relR2_rec_ci[1],
                mad_est_relR2_rec_high_ci = abs_err_est_relR2_rec_ci[2]

            )
        } # End of param loop
        
        # The parallel worker returns the list of data.tables for this trait
        return(param_results_list) 
        
    } # End of parallel trait loop

    # --- 4. COMBINE RESULTS (within sequential loop) ---
    if (length(trait_results_list) > 0) {
        final_results_list <- append(final_results_list, trait_results_list)
    }
  } # End of prop loop

} # End of sequential pop loop

results_dt <- rbindlist(final_results_list)
results_dt[, pgs_method := ifelse(str_detect(param, 'sbayesrc'), 'SBayesRC', 'pT+clump')]
results_dt$call_rate <- paste0('Call rate = ', 1 - results_dt$prop)

write.csv(results_dt, '~/wgs_sparse_recovered_results_rimp.csv', row.names=F)
system('dx upload ~/wgs_sparse_recovered_results_rimp.csv --path mind_the_gap/results/')

```

***

## UKB dosage-based GWAS

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_imp_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_ukb_imp_gwas.tar
```

***

### INFO-based masking

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_ukb_imp_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_ukb_imp_gwas.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# Populations to analyze
populations <- c('EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/imp_sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/imp_sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_boot <- 10              # Bootstraps of INDIVIDUALS 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_boot')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            
            # --- INFO MASKING (Single File, Deterministic) ---
            masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.prop_', prop, '.sscore')
            if (!file.exists(masked_file)) next
            
            pgs_masked <- fread(masked_file)
            if (!param_val %in% names(pgs_masked)) next
            
            pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
            full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
            
            n_curr <- nrow(full_data)
            if (n_curr < 10) next 
            
            # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
            y_vec <- full_data[['outcome']]
            s_vec <- full_data[[param_val]]
            sp_vec <- full_data[[paste0(param_val, "_masked")]]
            
            # --- INDIVIDUAL BOOTSTRAP LOOP ---
            boot_stats <- replicate(n_boot, {
                idx <- sample(n_curr, n_curr, replace=TRUE)
                
                # Use vector indexing instead of data.table subsetting
                y <- y_vec[idx]
                s <- s_vec[idx]
                s_prime <- sp_vec[idx]
                
                r2_s <- cor(y, s)^2
                r2_s_prime <- cor(y, s_prime)^2
                
                var_s <- var(s)
                var_s_prime <- var(s_prime)
                rel_var_obs <- var_s_prime / var_s
                
                # Handle extremely low baseline R2 safely
                obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                est_rel_r2 <- cor(s, s_prime)^2 
                
                error_val <- est_rel_r2 - obs_rel_r2
                abs_error_val <- abs(error_val)
                
                c(obs_rel_r2, est_rel_r2, error_val, abs_error_val, r2_s, r2_s_prime, var_s, var_s_prime, rel_var_obs)
            })
            
            # Transpose and remove failures (NAs)
            boot_stats <- t(boot_stats)
            boot_stats <- boot_stats[complete.cases(boot_stats), , drop=FALSE]
            
            if (nrow(boot_stats) == 0) next
            
            # Summarize the distribution
            means <- colMeans(boot_stats)
            cis <- apply(boot_stats, 2, quantile, probs=c(0.025, 0.975))
            
            trait_res_list[[length(trait_res_list) + 1]] <- data.table(
                population = pop,
                trait = trait_val,
                param = param_val,
                pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                prop_missing = prop,
                n_samples = n_curr,
                total_bootstraps = nrow(boot_stats),
                
                obs_rel_r2 = means[1],
                obs_rel_r2_low = cis[1, 1],
                obs_rel_r2_high = cis[2, 1],

                est_rel_r2 = means[2],
                est_rel_r2_low = cis[1, 2],
                est_rel_r2_high = cis[2, 2],
                
                error = means[3],
                error_low = cis[1, 3],
                error_high = cis[2, 3],
                
                mad = means[4],
                mad_low = cis[1, 4],
                mad_high = cis[2, 4],
                
                mean_r2_base = means[5],
                r2_base_low = cis[1, 5],
                r2_base_high = cis[2, 5],
                
                mean_r2_mask = means[6],
                r2_mask_low = cis[1, 6],
                r2_mask_high = cis[2, 6],
                
                mean_var_base = means[7],
                var_base_low = cis[1, 7],
                var_base_high = cis[2, 7],
                
                mean_var_mask = means[8],
                var_mask_low = cis[1, 8],
                var_mask_high = cis[2, 8],
                
                mean_rel_var_obs = means[9],
                rel_var_obs_low = cis[1, 9],
                rel_var_obs_high = cis[2, 9]
            )
        }
        return(rbindlist(trait_res_list))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_sparse_imp_results.csv', row.names=F)
system('dx upload ~/wgs_sparse_imp_results.csv --path mind_the_gap/results/')

```

***

## MVP dosage-based GWAS

***

### Random masking

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_mvp_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_random_masked_mvp_gwas.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# --- 1. CONFIGURATION ---

# Populations to analyze
populations <- c('EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_iterations <- 2          # Number of random masking iterations per trait/prop
n_boot_per_iter <- 2      # Bootstraps of INDIVIDUALS per iteration (10 * 100 = 1000 total)

# Variable to store AFR sample size dynamically
n_downsample_target <- NULL 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_iterations', 'n_boot_per_iter', 'n_downsample_target')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            n_full <- nrow(base_merged) # Capture the full sample size for this trait
            
            # --- ITERATION LOOP (Random Masking) ---
            # We will collect all 1000 bootstrap estimates (10 iters * 100 boots) for this trait/param
            boot_estimates_full <- list()
            boot_estimates_down <- list()
            
            for (iter in 1:n_iterations) {
                masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.random.prop_', prop, '.iteration_', iter, '.sscore')
                if (!file.exists(masked_file)) next
                
                pgs_masked <- fread(masked_file)
                if (!param_val %in% names(pgs_masked)) next
                
                pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
                full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
                
                # Define data subsets to analyze
                analysis_sets <- list(list(type="Full", data=full_data))
                
                # --- INDIVIDUAL BOOTSTRAP LOOP ---
                for (aset in analysis_sets) {
                    df <- aset$data
                    n_curr <- nrow(df)
                    if (n_curr < 10) next 
                    
                    # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
                    # Subsetting a data.table 100,000 times is memory-intensive. 
                    # Indexing numeric vectors directly is vastly faster.
                    y_vec <- df[['outcome']]
                    s_vec <- df[[param_val]]
                    sp_vec <- df[[paste0(param_val, "_masked")]]
                    
                    # Bootstrap individuals n_boot_per_iter times
                    boot_stats <- replicate(n_boot_per_iter, {
                        idx <- sample(n_curr, n_curr, replace=TRUE)
                        
                        # Use vector indexing instead of data.table subsetting
                        y <- y_vec[idx]
                        s <- s_vec[idx]
                        s_prime <- sp_vec[idx]
                        
                        r2_s <- cor(y, s)^2
                        r2_s_prime <- cor(y, s_prime)^2
                        
                        # Variances
                        var_s <- var(s)
                        var_s_prime <- var(s_prime)
                        rel_var_obs <- var_s_prime / var_s
                        
                        # Slopes
                        slope_base <- cov(y, s) / var_s
                        slope_mask <- cov(y, s_prime) / var_s_prime
                        
                        # Handle extremely low baseline R2 safely
                        obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                        est_rel_r2 <- cor(s, s_prime)^2 
                        
                        error_val <- est_rel_r2 - obs_rel_r2
                        abs_error_val <- abs(error_val)
                        
                        # Rescaled Slope (Calibration Factor = sqrt(Relative R2 / Relative Variance))
                        cf <- sqrt(est_rel_r2 / rel_var_obs)
                        slope_rescaled <- if (!is.na(cf) && cf > 1e-8) slope_mask / cf else NA
                        
                        # Slope Ratios
                        slope_ratio_uncalib <- if (!is.na(slope_base) && abs(slope_base) > 1e-8) slope_mask / slope_base else NA
                        slope_ratio_calib <- if (!is.na(slope_base) && abs(slope_base) > 1e-8 && !is.na(slope_rescaled)) slope_rescaled / slope_base else NA
                        
                        c(obs_rel_r2, est_rel_r2, error_val, abs_error_val,
                          var_s, var_s_prime, rel_var_obs,
                          slope_base, slope_mask, slope_rescaled,
                          slope_ratio_uncalib, slope_ratio_calib, r2_s, r2_s_prime)
                    })
                    
                    # Store bootstrap results
                    # boot_stats is a matrix of 12 rows x 100 cols
                    if (aset$type == "Full") {
                        boot_estimates_full[[iter]] <- t(boot_stats)
                    } else {
                        boot_estimates_down[[iter]] <- t(boot_stats)
                    }
                } 
            } # End Iteration Loop
            
            # --- SUMMARIZE ALL 1000 BOOTSTRAPS FOR THIS TRAIT ---
            process_bootstraps <- function(boot_list, type_label, n_samples_val) {
                if (length(boot_list) == 0) return(NULL)
                
                # Combine all 100 matrices into one 1000x12 matrix
                all_boots <- do.call(rbind, boot_list)
                all_boots <- all_boots[complete.cases(all_boots), , drop=FALSE] # Remove NAs from low R2
                
                if (nrow(all_boots) == 0) return(NULL)
                
                means <- colMeans(all_boots)
                cis <- apply(all_boots, 2, quantile, probs=c(0.025, 0.975))
                
                data.table(
                    population = pop,
                    trait = trait_val,
                    param = param_val,
                    pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                    prop_missing = prop,
                    sample_type = type_label,
                    n_samples = n_samples_val,
                    total_bootstraps = nrow(all_boots),
                    
                    obs_rel_r2 = means[1],
                    obs_rel_r2_low = cis[1, 1],
                    obs_rel_r2_high = cis[2, 1],

                    est_rel_r2 = means[2],
                    est_rel_r2_low = cis[1, 2],
                    est_rel_r2_high = cis[2, 2],
                    
                    error = means[3],
                    error_low = cis[1, 3],
                    error_high = cis[2, 3],
                    
                    mad = means[4],
                    mad_low = cis[1, 4],
                    mad_high = cis[2, 4],
                    
                    # Newly added metrics
                    mean_var_base = means[5],
                    var_base_low = cis[1, 5],
                    var_base_high = cis[2, 5],
                    
                    mean_var_mask = means[6],
                    var_mask_low = cis[1, 6],
                    var_mask_high = cis[2, 6],
                    
                    mean_rel_var_obs = means[7],
                    rel_var_obs_low = cis[1, 7],
                    rel_var_obs_high = cis[2, 7],
                    
                    mean_slope_base = means[8],
                    slope_base_low = cis[1, 8],
                    slope_base_high = cis[2, 8],
                    
                    mean_slope_mask = means[9],
                    slope_mask_low = cis[1, 9],
                    slope_mask_high = cis[2, 9],
                    
                    mean_slope_rescaled = means[10],
                    slope_rescaled_low = cis[1, 10],
                    slope_rescaled_high = cis[2, 10],
                    
                    mean_slope_ratio_uncalib = means[11],
                    slope_ratio_uncalib_low = cis[1, 11],
                    slope_ratio_uncalib_high = cis[2, 11],
                    
                    mean_slope_ratio_calib = means[12],
                    slope_ratio_calib_low = cis[1, 12],
                    slope_ratio_calib_high = cis[2, 12],
                    
                    mean_r2_base = means[13],
                    r2_base_low = cis[1, 13],
                    r2_base_high = cis[2, 13],
                    
                    mean_r2_mask = means[14],
                    r2_mask_low = cis[1, 14],
                    r2_mask_high = cis[2, 14]
                )
            }

            res_full <- process_bootstraps(boot_estimates_full, "Full", n_full)
            res_down <- process_bootstraps(boot_estimates_down, "Downsampled", n_down)
            
            trait_res_list[[length(trait_res_list) + 1]] <- rbind(res_full, res_down)
        }
        return(rbindlist(trait_res_list))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_mvp_sparse_results_10it_random.csv', row.names=F)
system('dx upload ~/wgs_mvp_sparse_results_10it_random.csv --path mind_the_gap/results/')

```

***

### INFO-based masking

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_mvp_gwas.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_mvp_gwas.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# Populations to analyze
populations <- c('EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_boot <- 10              # Bootstraps of INDIVIDUALS 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_boot')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            
            # --- INFO MASKING (Single File, Deterministic) ---
            masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse.prop_', prop, '.sscore')
            if (!file.exists(masked_file)) next
            
            pgs_masked <- fread(masked_file)
            if (!param_val %in% names(pgs_masked)) next
            
            pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
            full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
            
            n_curr <- nrow(full_data)
            if (n_curr < 10) next 
            
            # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
            y_vec <- full_data[['outcome']]
            s_vec <- full_data[[param_val]]
            sp_vec <- full_data[[paste0(param_val, "_masked")]]
            
            # --- INDIVIDUAL BOOTSTRAP LOOP ---
            boot_stats <- replicate(n_boot, {
                idx <- sample(n_curr, n_curr, replace=TRUE)
                
                # Use vector indexing instead of data.table subsetting
                y <- y_vec[idx]
                s <- s_vec[idx]
                s_prime <- sp_vec[idx]
                
                r2_s <- cor(y, s)^2
                r2_s_prime <- cor(y, s_prime)^2
                
                var_s <- var(s)
                var_s_prime <- var(s_prime)
                rel_var_obs <- var_s_prime / var_s
                
                # Handle extremely low baseline R2 safely
                obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                est_rel_r2 <- cor(s, s_prime)^2 
                
                error_val <- est_rel_r2 - obs_rel_r2
                abs_error_val <- abs(error_val)
                
                c(obs_rel_r2, est_rel_r2, error_val, abs_error_val, r2_s, r2_s_prime, var_s, var_s_prime, rel_var_obs)
            })
            
            # Transpose and remove failures (NAs)
            boot_stats <- t(boot_stats)
            boot_stats <- boot_stats[complete.cases(boot_stats), , drop=FALSE]
            
            if (nrow(boot_stats) == 0) next
            
            # Summarize the distribution
            means <- colMeans(boot_stats)
            cis <- apply(boot_stats, 2, quantile, probs=c(0.025, 0.975))
            
            trait_res_list[[length(trait_res_list) + 1]] <- data.table(
                population = pop,
                trait = trait_val,
                param = param_val,
                pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                prop_missing = prop,
                n_samples = n_curr,
                total_bootstraps = nrow(boot_stats),
                
                obs_rel_r2 = means[1],
                obs_rel_r2_low = cis[1, 1],
                obs_rel_r2_high = cis[2, 1],

                est_rel_r2 = means[2],
                est_rel_r2_low = cis[1, 2],
                est_rel_r2_high = cis[2, 2],
                
                error = means[3],
                error_low = cis[1, 3],
                error_high = cis[2, 3],
                
                mad = means[4],
                mad_low = cis[1, 4],
                mad_high = cis[2, 4],
                
                mean_r2_base = means[5],
                r2_base_low = cis[1, 5],
                r2_base_high = cis[2, 5],
                
                mean_r2_mask = means[6],
                r2_mask_low = cis[1, 6],
                r2_mask_high = cis[2, 6],
                
                mean_var_base = means[7],
                var_base_low = cis[1, 7],
                var_base_high = cis[2, 7],
                
                mean_var_mask = means[8],
                var_mask_low = cis[1, 8],
                var_mask_high = cis[2, 8],
                
                mean_rel_var_obs = means[9],
                rel_var_obs_low = cis[1, 9],
                rel_var_obs_high = cis[2, 9]
            )
        }
        return(rbindlist(trait_res_list))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_mvp_sparse_results.csv', row.names=F)
system('dx upload ~/wgs_mvp_sparse_results.csv --path mind_the_gap/results/')

```

***

### INFO-based masking (sqrt_info correction)

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_mvp_gwas_sqrtinfo.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_mvp_gwas_sqrtinfo.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# Populations to analyze
populations <- c('EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_boot <- 10              # Bootstraps of INDIVIDUALS 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse_sqrt_info_pre.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_boot')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            
            # --- INFO MASKING (Single File, Deterministic) ---
            masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse_sqrt_info_pre.prop_', prop, '.sscore')
            if (!file.exists(masked_file)) next
            
            pgs_masked <- fread(masked_file)
            if (!param_val %in% names(pgs_masked)) next
            
            pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
            full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
            
            n_curr <- nrow(full_data)
            if (n_curr < 10) next 
            
            # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
            y_vec <- full_data[['outcome']]
            s_vec <- full_data[[param_val]]
            sp_vec <- full_data[[paste0(param_val, "_masked")]]
            
            # --- INDIVIDUAL BOOTSTRAP LOOP ---
            boot_stats <- replicate(n_boot, {
                idx <- sample(n_curr, n_curr, replace=TRUE)
                
                # Use vector indexing instead of data.table subsetting
                y <- y_vec[idx]
                s <- s_vec[idx]
                s_prime <- sp_vec[idx]
                
                r2_s <- cor(y, s)^2
                r2_s_prime <- cor(y, s_prime)^2
                
                var_s <- var(s)
                var_s_prime <- var(s_prime)
                rel_var_obs <- var_s_prime / var_s
                
                # Handle extremely low baseline R2 safely
                obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                est_rel_r2 <- cor(s, s_prime)^2 
                
                error_val <- est_rel_r2 - obs_rel_r2
                abs_error_val <- abs(error_val)
                
                c(obs_rel_r2, est_rel_r2, error_val, abs_error_val, r2_s, r2_s_prime, var_s, var_s_prime, rel_var_obs)
            })
            
            # Transpose and remove failures (NAs)
            boot_stats <- t(boot_stats)
            boot_stats <- boot_stats[complete.cases(boot_stats), , drop=FALSE]
            
            if (nrow(boot_stats) == 0) next
            
            # Summarize the distribution
            means <- colMeans(boot_stats)
            cis <- apply(boot_stats, 2, quantile, probs=c(0.025, 0.975))
            
            trait_res_list[[length(trait_res_list) + 1]] <- data.table(
                population = pop,
                trait = trait_val,
                param = param_val,
                pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                prop_missing = prop,
                n_samples = n_curr,
                total_bootstraps = nrow(boot_stats),
                
                obs_rel_r2 = means[1],
                obs_rel_r2_low = cis[1, 1],
                obs_rel_r2_high = cis[2, 1],

                est_rel_r2 = means[2],
                est_rel_r2_low = cis[1, 2],
                est_rel_r2_high = cis[2, 2],
                
                error = means[3],
                error_low = cis[1, 3],
                error_high = cis[2, 3],
                
                mad = means[4],
                mad_low = cis[1, 4],
                mad_high = cis[2, 4],
                
                mean_r2_base = means[5],
                r2_base_low = cis[1, 5],
                r2_base_high = cis[2, 5],
                
                mean_r2_mask = means[6],
                r2_mask_low = cis[1, 6],
                r2_mask_high = cis[2, 6],
                
                mean_var_base = means[7],
                var_base_low = cis[1, 7],
                var_base_high = cis[2, 7],
                
                mean_var_mask = means[8],
                var_mask_low = cis[1, 8],
                var_mask_high = cis[2, 8],
                
                mean_rel_var_obs = means[9],
                rel_var_obs_low = cis[1, 9],
                rel_var_obs_high = cis[2, 9]
            )
        }
        return(rbindlist(trait_res_list))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_mvp_sparse_sqrt_info_pre_results.csv', row.names=F)
system('dx upload ~/wgs_mvp_sparse_sqrt_info_pre_results.csv --path mind_the_gap/results/')

```

***

### INFO-based masking (INFO > 0.9)

```{bash}
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_baseline_mvp_gwas_rsq_90.tar
tar -xvf /mnt/project/mind_the_gap/pgs_wgs_sparse_masked_mvp_gwas_rsq_90.tar
```

```{r}
library(data.table)
library(scales)
library(ggplot2)
library(cowplot)
library(stringr)
library(foreach)
library(doMC)
registerDoMC(8)

# Populations to analyze
populations <- c('EUR') 
traits <- c('HT', 'HB', 'PLT', 'BMI', 'BWT', 'MCHC', 'SBP', 'NEU', 'HDL', 'TC')

# File paths
base_dir_baseline <- '~/pgs_wgs/sparse/baseline/'
base_dir_masked <- '~/pgs_wgs/sparse/masked/'
pheno_dir <- '/mnt/project/mind_the_gap/phenotype/'

# Simulation parameters
props <- c(0.05, 0.25, 0.5) # Proportions of missingness
n_boot <- 10              # Bootstraps of INDIVIDUALS 

set.seed(123)

# List to collect results
final_results_list <- list()

# --- 2. OUTER LOOP: POPULATIONS ---
for (pop in populations) {
    print(paste("Processing Population:", pop))
  
  for (prop in props) {
    print(paste("Processing prop:", prop))
    
    # Check baseline file existence
    baseline_file <- paste0(base_dir_baseline, 'ukb.', pop, '.sparse_rsq_90.sscore')
    if (!file.exists(baseline_file)) {
        print(paste("Skipping baseline not found:", baseline_file))
        next
    }
    pgs_baseline <- fread(baseline_file)
    pgs_baseline <- pgs_baseline[, !grepl('AVG', names(pgs_baseline)), with = F]
    
    # --- 3. INNER LOOP: TRAITS (Parallel) ---
    trait_results_list <- foreach(
        trait_val = traits,
        .combine = 'rbind',
        .packages = 'data.table',
        .export = c('pheno_dir', 'base_dir_masked', 'pgs_baseline', 'pop', 'prop', 'n_boot')
    ) %dopar% {
        
        # Load Phenotype
        pheno_file <- paste0(pheno_dir, trait_val, '.', pop, '.txt')
        if (!file.exists(pheno_file)) return(NULL)
        pheno <- fread(pheno_file)

        # Extract relevant PGS column for this trait from baseline
        pgs_cols <- names(pgs_baseline)[grepl('SUM', names(pgs_baseline)) & grepl(trait_val, names(pgs_baseline))]
        
        trait_res_list <- list()
        
        for (param_val in pgs_cols) {
            pgs_base_sub <- pgs_baseline[, c('IID', param_val), with=FALSE]
            base_merged <- merge(pheno, pgs_base_sub, by='IID', sort=FALSE)
            
            # --- INFO MASKING (Single File, Deterministic) ---
            masked_file <- paste0(base_dir_masked, 'ukb.', pop, '.sparse_rsq_90.prop_', prop, '.sscore')
            if (!file.exists(masked_file)) next
            
            pgs_masked <- fread(masked_file)
            if (!param_val %in% names(pgs_masked)) next
            
            pgs_mask_sub <- pgs_masked[, c('IID', param_val), with=FALSE]
            full_data <- merge(base_merged, pgs_mask_sub, by='IID', suffixes=c("", "_masked"))
            
            n_curr <- nrow(full_data)
            if (n_curr < 10) next 
            
            # PERFORMANCE OPTIMIZATION: Extract to basic vectors outside the loop
            y_vec <- full_data[['outcome']]
            s_vec <- full_data[[param_val]]
            sp_vec <- full_data[[paste0(param_val, "_masked")]]
            
            # --- INDIVIDUAL BOOTSTRAP LOOP ---
            boot_stats <- replicate(n_boot, {
                idx <- sample(n_curr, n_curr, replace=TRUE)
                
                # Use vector indexing instead of data.table subsetting
                y <- y_vec[idx]
                s <- s_vec[idx]
                s_prime <- sp_vec[idx]
                
                r2_s <- cor(y, s)^2
                r2_s_prime <- cor(y, s_prime)^2
                
                var_s <- var(s)
                var_s_prime <- var(s_prime)
                rel_var_obs <- var_s_prime / var_s
                
                # Handle extremely low baseline R2 safely
                obs_rel_r2 <- if(r2_s > 1e-8) r2_s_prime / r2_s else NA
                est_rel_r2 <- cor(s, s_prime)^2 
                
                error_val <- est_rel_r2 - obs_rel_r2
                abs_error_val <- abs(error_val)
                
                c(obs_rel_r2, est_rel_r2, error_val, abs_error_val, r2_s, r2_s_prime, var_s, var_s_prime, rel_var_obs)
            })
            
            # Transpose and remove failures (NAs)
            boot_stats <- t(boot_stats)
            boot_stats <- boot_stats[complete.cases(boot_stats), , drop=FALSE]
            
            if (nrow(boot_stats) == 0) next
            
            # Summarize the distribution
            means <- colMeans(boot_stats)
            cis <- apply(boot_stats, 2, quantile, probs=c(0.025, 0.975))
            
            trait_res_list[[length(trait_res_list) + 1]] <- data.table(
                population = pop,
                trait = trait_val,
                param = param_val,
                pgs_method = ifelse(grepl('sbayesrc', param_val, ignore.case=T), 'SBayesRC', 'pT+clump'),
                prop_missing = prop,
                n_samples = n_curr,
                total_bootstraps = nrow(boot_stats),
                
                obs_rel_r2 = means[1],
                obs_rel_r2_low = cis[1, 1],
                obs_rel_r2_high = cis[2, 1],

                est_rel_r2 = means[2],
                est_rel_r2_low = cis[1, 2],
                est_rel_r2_high = cis[2, 2],
                
                error = means[3],
                error_low = cis[1, 3],
                error_high = cis[2, 3],
                
                mad = means[4],
                mad_low = cis[1, 4],
                mad_high = cis[2, 4],
                
                mean_r2_base = means[5],
                r2_base_low = cis[1, 5],
                r2_base_high = cis[2, 5],
                
                mean_r2_mask = means[6],
                r2_mask_low = cis[1, 6],
                r2_mask_high = cis[2, 6],
                
                mean_var_base = means[7],
                var_base_low = cis[1, 7],
                var_base_high = cis[2, 7],
                
                mean_var_mask = means[8],
                var_mask_low = cis[1, 8],
                var_mask_high = cis[2, 8],
                
                mean_rel_var_obs = means[9],
                rel_var_obs_low = cis[1, 9],
                rel_var_obs_high = cis[2, 9]
            )
        }
        return(rbindlist(trait_res_list))
    }
    
    if (!is.null(trait_results_list)) {
        final_results_list[[length(final_results_list) + 1]] <- trait_results_list
    }
  }
}

# Combine all trait-level summaries
results_dt <- rbindlist(final_results_list)

write.csv(results_dt, '~/wgs_mvp_sparse_rsq_90_results.csv', row.names=F)
system('dx upload ~/wgs_mvp_sparse_rsq_90_results.csv --path mind_the_gap/results/')

```

***
