---
title: Genetic analysis of Mirocals trial
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 2
    css: styles/styles.css
    includes:
      in_header: header.html
      after_body: footer.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{css, echo=F}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

***

# Proposal

The Mirocals trial tested the efficacy of low dose IL-2 for treatment of ALS. The trial showed IL-2 + riluzole treatment provided a significant improvement in survival compared to those only receiving placebo + riluzole. Low dose IL-2 has been shown to increase the number of regulatory T-cells (Tregs).

This trial has collected whole genome sequence data and methylation data (across several time points). This presents an opportunity to identify genetic predictors/biomarkers of response to IL-2 treatment. The sample size of the trial is limited (N drug = 110, N placebo = 110), which means we need to test specific hypotheses and leverage larger external datasets.

***

## Outcome

There are two key outcome variables from the trial, including survival and regulatory T-cell numbers. It seems survival is the more meaningful outcome for patients, but perhaps there are benefits to using T-cell numbers that I am unaware of. If we are using survival, then we should using a cox regression, but if use T cell numbers linear regression.

***

## Polygenic scores (PGS)

List of polygenic scores that might predict response to IL-2:

* ALS risk, survival, age at onset
* Risk of other neurodegenerative diseases
* Interleukin-2 receptor subunit alpha (IL2RA) protein levels
* Regulatory T-cell molecule levels
* Autoimmune disease risk
* Other inflammatory markers (CRP, IL-6, etc.)

I would suggest using SBayesRC as the method to generate the PGS. If GWAS from multiple populations are available we can using SBayesRC-multi.

We could also further stratify the polygenic scores by functional annotations, such as pathway membership. This could be particularly useful, as it may be that only certain pathways are relevant to IL-2 efficacy. For example, ALS risk mediated by inflammatory pathways may be particularly relevant for IL-2 efficacy. 

***

## Methylation scores

We can use previously published EWAS summary statistics to do the same thing.

***

# Calculate PGS

```{bash}
mkdir -p /users/k1806347/oliverpainfel/Data/miraclestrial/original

# Create index for each batch
module add bcftools/1.12-gcc-13.2.0-python-3.11.6

####
# Format for GenoPred
####
# Split into per chromosome files

##
# Batch 1
## 
for chr in $(seq 1 22);do
    sbatch --mem 10G -n 1 -p interruptible_cpu -t 1:00:00 --wrap="/users/k1806347/oliverpainfel/Software/plink2 \
        --vcf /scratch/prj/mirocalstrial/recovered/DNAscan/variantcalling/merged_strelka_stringent.vcf.gz \
        --chr ${chr} \
        --out /users/k1806347/oliverpainfel/Data/miraclestrial/processed/miracles_batch1.chr${chr} \
        --allow-extra-chr \
        --export vcf bgz"
done

# The file is truncated mid chr16, and missing chr 17-22

# Recreate merged dataset 
module add bcftools/1.12-gcc-13.2.0-python-3.11.6

bcftools merge \
  -l sample_vcf_list.txt \
  --missing-to-ref \
  -o /users/k1806347/oliverpainfel/Data/miraclestrial/original/miracles_batch1.vcf.gz

# Replace corrupt version
cp /users/k1806347/oliverpainfel/Data/miraclestrial/original/miracles_batch1.vcf.gz /scratch/prj/mirocalstrial/recovered/DNAscan/variantcalling/merged_strelka_stringent_fixed.vcf.gz

# Split by chromosome
for chr in $(seq 1 22);do
    sbatch --mem 10G -n 1 -p interruptible_cpu -t 1:00:00 --wrap="/users/k1806347/oliverpainfel/Software/plink2 \
        --vcf /users/k1806347/oliverpainfel/Data/miraclestrial/original/miracles_batch1.vcf.gz \
        --chr ${chr} \
        --out /users/k1806347/oliverpainfel/Data/miraclestrial/processed/miracles_batch1.chr${chr} \
        --allow-extra-chr \
        --export vcf bgz"
done

##
# Batch 2
## 
for chr in $(seq 1 22);do
    sbatch --mem 10G -n 1 -p interruptible_cpu -t 1:00:00 --wrap="/users/k1806347/oliverpainfel/Software/plink2 \
        --vcf /scratch/prj/mirocalstrial/recovered/DNAscan_batch2/variantcalling/merged_strelka_stringent.vcf.gz \
        --chr ${chr} \
        --out /users/k1806347/oliverpainfel/Data/miraclestrial/processed/miracles_batch2.chr${chr} \
        --allow-extra-chr \
        --export vcf bgz"
done

# Batch 2 seems fine.

```

```{r}
library(data.table)

dir.create('/users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/output', recursive = T)
dir.create('/users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config', recursive = T)

# Create target_list
target_list <- NULL
target_list <- rbind(target_list, data.table(
  name = 'miracles_batch1',
  path = '/users/k1806347/oliverpainfel/Data/miraclestrial/processed/miracles_batch1',
  type = 'vcf',
  indiv_report = F))

target_list <- rbind(target_list, data.table(
  name = 'miracles_batch2',
  path = '/users/k1806347/oliverpainfel/Data/miraclestrial/processed/miracles_batch2',
  type = 'vcf',
  indiv_report = F))

write.table(
  target_list,
  '/users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config/target_list.txt',
  col.names = T,
  row.names = F,
  quote = F,
  sep = ' '
)

# Create gwas_list
gwas_list <- data.table(
  name='yengo_eur',
  path = '/users/k1806347/oliverpainfel/Data/GWAS_sumstats/opensnp_test/yengo_2022_height_eur.txt',
  population = 'EUR',
  n = NA,
  sampling = NA,
  prevalence = NA,
  mean = NA,
  sd = NA,
  label = "\"Yengo 2022 Height EUR\"")

write.table(
  gwas_list,
  '/users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config/gwas_list.txt',
  col.names = T,
  row.names = F,
  quote = F,
  sep = ' '
)

# Create config file
conf <- c(
  'outdir: /users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/output',
  'config_file: /users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config/config.yaml',
  'target_list: /users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config/target_list.txt',
  'gwas_list: /users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config/gwas_list.txt',
  "pgs_methods: ['dbslmm']"
)

write.table(
  conf,
  '/users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config/config.yaml',
  col.names = F,
  row.names = F,
  quote = F
)

```

```{bash}
cd /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline
conda activate genopred

snakemake \
  --profile slurm \
  --use-conda \
  --configfile=/users/k1806347/oliverpainfel/Data/miraclestrial/GenoPred/config/config.yaml \
  output_all -n 
```

The vast majority of the clinical trial participants are of European ancestry (Non-EUR N = 15, EUR N = 267). We should restrict the primary analysis to EUR, and then rerun key analyses in full sample. 

***

Analyses:

1. Replicate primary analysis of clinical trial (is this published/can I see details on this)
2. Then test for interaction between PGS and IL-22 on primary endpoint
  - We will have very little power to detect main/interaction effects at statistical significance
  - Could test PGS for risk as primary analysis, as previous studies have done, and then do a hypothesis-free analysis across many PGS, controlling for multiple testing?
    - ALS risk GWAS Could build multi-trait and multi-population model. Build elastic net containing many scores using ProjectMine and independent GWAS
    

What about Jiajing's PGS? seemed to predict risk very highly.

I can generate the PGS and hand them over, or do the analysis? Depends on how Ahmad wants to structure the papers.

***

# Clinical data

I have been sent some preliminary clinical data for the trial. Lets have a look at it:

```{r}
library(data.table)
library(readxl)

data <- read_excel("~/oliverpainfel/scratch_tmp/Data/miraclestrial/original/Survival-MIR.xlsx")
data <- data.table(data)

# Rename variables
original_names<-names(data)
names(data) <- c('reference','label','id','c9','unc13a','diagnosis','country','site','reference_2','sex','age_at_diagnosis','age_at_onset','site_of_onset','arm','sex_2','familial','alfrs','svc','age_at_death','age_at_death_rounded_down','randomisation','reference_3','gender','age_c','site_of_onset_2','arm_2')

# Overview of data
str(data)
summary(data)

# Check references are identical
all(data$reference == data$reference_2)
all(data$reference == data$reference_3) # We can delete reference_3 as identical to reference
data$reference_3 <- NULL
data[data$reference != data$reference_2,] # There is an reference value matches different row in reference_2.

# Check sex are identical
all(data$sex == data$sex_2)
all(data$sex == data$gender)
all(data$sex_2 == data$gender) # We can delete gender as identical to sex_2
data$gender <- NULL
data[data$sex != data$sex_2,] # The same rows with discordant reference values have discordant sex values.

# Check site of onset
all(data$site_of_onset == data$site_of_onset_2)
data[data$site_of_onset != data$site_of_onset_2,] # One individual has different site of onset (also mismatched for reference)

# Check arm
all(data$arm == data$arm_2, na.rm = T) # They are identical
data$arm_2 <- NULL
data$arm

# Check age variables
# age_c seems to be age_at_diagnosis, just rounded to 1 decimal point, so delete age_c
data$age_c <- NULL

# Create variable indicating death before end of trial
data$age_at_death_rounded_down <- NULL
data$died_during_trial <- ifelse(grepl('Alive at cut off date', data$age_at_death), F, T)
data$died_during_trial[grepl('NA', data$age_at_death)] <- NA
data$age_at_death <- as.numeric(data$age_at_death)

# Check site of onset
all(data$site_of_onset == data$site_of_onset_2)

#####
# Run cox regression
#####
library(survival)

# Subset those in the trial
trial_data <- data[!is.na(data$arm),]
trial_data <- trial_data[trial_data$randomisation == 'RN',]
nrow(trial_data) # 218 - This is two less than the publication

# We need to create a time to event variable
# We can assume age at diagnosis is the date of enrolment
# The trial ran for 640 days, so we can insert time_to_event for those that survived
trial_data$time_to_event <- trial_data$age_at_death - trial_data$age_at_diagnosis
trial_data$time_to_event[trial_data$died_during_trial == F] <- trial_data$age_at_death[trial_data$died_during_trial == F] - trial_data$age_at_diagnosis[trial_data$died_during_trial == F] + (640/365)
trial_data$status <- as.numeric(trial_data$died_during_trial)
trial_data$arm_numeric <- ifelse(trial_data$arm == 'IL2', 1, 0)

cox_model <- coxph(Surv(time_to_event, status) ~ arm_numeric, data = trial_data)
summary(cox_model)

fit <- survfit(Surv(time_to_event, status) ~ arm_numeric, data = trial_data)
plot(fit, col = c("blue", "red"), xlab = "Time (years)", ylab = "Survival probability")
legend("bottomleft", legend = c("Control", "Treatment"), col = c("blue", "red"), lty = 1)

# Shows non-significant increase in survival in those given IL-2
# The real trial used prognostic covariates which we do not have access to currently.
```




