---
title: Model Builder Top1 Script Demo
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 2
    css: styles/styles.css
    includes:
      in_header: header.html
      after_body: footer.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{css, echo=F}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

***

This document provides a demonstration of the `model_builder_top1.R` script. This script derives and evaluates predictive models using a structured approach:

1. **Top1 Model**: Identifies the best predictor within a specified `top1` group.
2. **Multi Model**: Constructs a generalized linear model (GLM) incorporating all best predictors from different `top1` groups within the same `multi` group.
3. **Evaluation**: Uses cross-validation to train and evaluate both `top1` and `multi` models.
4. **Comparison**: Applies Williams' test to compare the predictive utility of the models (using `psych::paired.r` function).

***

## Input Files

### Outcome File (`--outcome`)

- A file containing the outcome data to be predicted.
- Column names are required.
- Acceptable delimiters: space, tab, or comma.
- Required format:

  1. **FID**: Family ID
  2. **IID**: Individual ID
  3. **Outcome Variable** (any column name is acceptable)

### Predictor List File (`--predictors`)

- A file listing the predictor data files and their respective grouping information.
- Must contain three columns with headers:

  - **predictor**: Path to the predictor data file.
  - **multi**: Grouping variable specifying which `top1` models should be combined into a `multi` model.
  - **top1**: Group variable indicating which predictor files should be considered together when creating a `top1` model.

Each **predictor file** must follow this format:

- Column names are required.
- Acceptable delimiters: space, tab, or comma.
- Required format:

  1. **FID**: Family ID
  2. **IID**: Individual ID
  3. Additional columns: Predictor variables (any column names are acceptable)

***

## Generating Demo Data

Below is an example script to simulate a phenotype and polygenic scores (PGS) across different p-value thresholds, derived from two separate GWAS datasets.

```{r}
# Create a directory for the demo
dir.create('~/test_model_builder')
set.seed(1)
n <- 1000
ids <- 1:n
data <- data.frame(FID = ids, IID = ids, outcome = rnorm(n))

# Simulate PGS for two GWAS datasets
for(i in 1:2){
  pgs_tmp <- data$outcome + rnorm(n)
  for(j in c(0.05, 0.01, 0.5)){
    data[[paste0('pgs',i,'_',j)]] <- pgs_tmp + rnorm(n, 0, j*5)
  }
}

# Save outcome data
write.table(data[, c('FID','IID','outcome')], '~/test_model_builder/outcome.txt', col.names=TRUE, row.names=FALSE, quote=FALSE)

# Save PGS data from each GWAS
write.table(data[, grepl('FID|IID|pgs1', names(data))], '~/test_model_builder/pgs1.txt', col.names=TRUE, row.names=FALSE, quote=FALSE)
write.table(data[, grepl('FID|IID|pgs2', names(data))], '~/test_model_builder/pgs2.txt', col.names=TRUE, row.names=FALSE, quote=FALSE)
```

***

## Creating the Predictor List File

The following script generates a predictor list file, linking predictor files to their respective `multi` and `top1` groups.

```{r}
pred <- data.frame(
  predictor = c('~/test_model_builder/pgs1.txt','~/test_model_builder/pgs2.txt'),
  multi = 'combo',
  top1 = c('pgs1','pgs2')
)

write.table(pred, '~/test_model_builder/predictor_list.txt', col.names=TRUE, row.names=FALSE, quote=FALSE)
```

***

## Running the Model Builder Script

To execute the `model_builder_top1.R` script, activate the appropriate Conda environment and run the script with the required input files. Note, a .yaml file to create the model_builder environment can be found [here](https://github.com/opain/GenoPred/blob/gwas_grouping/pipeline/envs/model_builder.yaml).

```{bash}
conda activate model_builder

Rscript ../Scripts/model_builder/model_builder_top1.R \
  --outcome ~/test_model_builder/outcome.txt \
  --predictors ~/test_model_builder/predictor_list.txt \
  --out ~/test_model_builder/res
```

***

## Output Files

Upon execution, the script will generate the following outputs:

- `<out>.log` - Log file recording the execution process.
- `<out>.pred_eval.txt` - Evaluation results indicating the predictive utility of each model.
- `<out>.pred_comp.txt` - Comparative results of predictive utility across models.
- `<out>.group_list.txt` - Summary of the number of predictors used in each `multi` and `top1` model.


