---
title: Preparing LDpred2 LD matrices
output: 
  html_document:
    theme: cosmo
    css: styles/styles.css
    includes:
      in_header: header.html
      after_body: footer.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{css, echo=F}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

***

# 1KG+HGDP

Here we will calculate LD matrices for LDpred2 using HapMap3 SNPs in the 1KG+HGDP super populations. This is to allow fair comparison across methods for a cross-population polygenic scoring study. 

The code will follow that used to generate the LD matrices for EUR, in this paper ([link](https://doi.org/10.1016/j.xhgg.2022.100136)).

***

## Full reference

```{bash}
# Subset 1KG+HGDP into super populations and merge across chromosomes
for pop in $(echo EUR EAS AFR); do
  mkdir -p /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/hm3/${pop}

  # Create mergelist
  rm -f /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/hm3/${pop}/mergelist.txt
  for chr in $(seq 1 22); do
    echo /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/resources/data/ref/ref.chr${chr} >> /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/hm3/${pop}/mergelist.txt
  done
  
  /users/k1806347/oliverpainfel/Software/plink2_linux_avx2_20241011/plink2 \
    --pmerge-list /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/hm3/${pop}/mergelist.txt \
    --make-bed \
    --keep /users/k1806347/oliverpainfel/Software/MyGit/GenoPred/pipeline/resources/data/ref/keep_files/${pop}.keep \
    --out /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/hm3/${pop}/ref.${pop}.GW
done
```

```{r}
# export OPENBLAS_NUM_THREADS=1
# conda activate .snakemake/conda/329e0288cb99508f5e6c50a0996b234c_
library(bigsnpr)
library(bigreadr)

# Create object for genetic data in ldpred2 format, and impute missing values
for(pop in c('EUR', 'EAS', 'AFR')){
  # Read in reference data
  snp_readBed(
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/ref.',
      pop,
      '.GW.bed'
    )
  )
  
  # Attach the ref object in R session
  ref <-
    snp_attach(
      paste0(
        '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
        pop,
        '/ref.',
        pop,
        '.GW.rds'
      )
    )
  G <- ref$genotypes

  #### Impute missing values (bigsnpr can't handle missing data in most functions)
  G_imp <- snp_fastImputeSimple(G, method = "mean2", ncores = nb_cores())

  # Save imputed reference
  ref$genotypes<-G_imp
  saveRDS(ref, paste0(
        '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
        pop,
        '/ref.',
        pop,
        '.GW.rds'
      ))
}

################
# Now calculate LD matrices following code from https://doi.org/10.1016/j.xhgg.2022.100136

library(dplyr)
library(bigreadr)
library(bigsnpr)
library(future.batchtools)
library(furrr)

for(pop in c('EUR', 'EAS', 'AFR')){
  obj.hgdp_1kg <- snp_attach(paste0(
          '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
          pop,
          '/ref.',
          pop,
          '.GW.rds'
        ))
  G <- obj.hgdp_1kg$genotypes
  CHR <- obj.hgdp_1kg$map$chromosome
  POS <- obj.hgdp_1kg$map$physical.pos
  POS2 <- snp_asGeneticPos(CHR, POS, dir ='/users/k1806347/oliverpainfel/Data/Genetic_Map/CEU', ncores = nb_cores()) # Note this is not appropriate but I think Florian did the same as used distance from UKB
  MAF <- snp_MAF(G, ncores = nb_cores())
  keep <- (MAF > 0.02)

  plan(batchtools_slurm(resources = list(
    time = "12:00:00",
    ntasks = 5,
    mem = "20g",
    partition = "neurohack_cpu"
  )))
  
  all_final_grp <- furrr::future_map_dfr(1:22, function(chr) {
    ind.chr2 <- which(CHR == chr & keep)

    corr0 <- runonce::save_run({
      snp_cor(
        G,
        ind.col = ind.chr2,
        infos.pos = POS2[ind.chr2],
        size = 3 / 1000,
        ncores = nb_cores()
      )
    }, file = paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/LD_chr',
      chr,
      ".rds"
    ))
    
    # find nearly independent LD blocks
    # Playing around with the snp_ldsplit function, the max_r2 needs to be increased to 0.2 for AFR populations.
    # This might be because we are using an inappropriate recombination map (EUR)?
    # Use max_r2 = 0.15 for EUR and EAS as in Florian's code, but increase to 0.2 for AFR. Note. this is still lower than the default of 0.3.
    max_r2 <- ifelse(pop %in% c('EUR','EAS'), 0.15, 0.25)
    m <- length(ind.chr2)
    (SEQ <- round(seq_log(m / 30, m / 5, length.out = 20)))
    splits <- snp_ldsplit(corr0, thr_r2 = 0.05, min_size = 50, max_size = SEQ, max_r2 = max_r2)
    splits$cost2 <- sapply(splits$all_size, function(sizes) sum(sizes^2))
  
    best_split <- splits %>%
      arrange(cost2 * sqrt(5 + cost)) %>%
      print() %>%
      slice(1) %>%
      print()
  
    (all_size <- best_split$all_size[[1]])
    best_grp <- rep(seq_along(all_size), all_size)
  
    runonce::save_run({
      corr0T <- as(corr0, "dgTMatrix")
      corr0T@x <-
        ifelse(best_grp[corr0T@i + 1L] == best_grp[corr0T@j + 1L], corr0T@x, 0)
      as(Matrix::drop0(corr0T), "symmetricMatrix")
    }, file = paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/LD_with_blocks_chr',
      chr,
      ".rds"
    ))
  
    # return
    tibble(best_split, ind = list(ind.chr2))
  })
  
  saveRDS(
    all_final_grp,
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/all_final_grp.rds'
    )
  )
  
  ###
  # Create map.rds
  ###
  map <- obj.hgdp_1kg$map[keep,]

  # Compute allele frequency
  map$af<-big_colstats(G, ind.col = which(keep), ncores = nb_cores())$sum / (2 * nrow(G))
  
  # Compute LD scores
  map$ld <- do.call('c', lapply(1:22, function(chr) {
    cat(chr, ".. ", sep = "")
    corr_chr <- readRDS(
      paste0(
        '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
        pop,
        '/LD_chr',
        chr,
        ".rds"
      )
    )
    Matrix::colSums(corr_chr ^ 2)
  }))
  
  map <- map[, c('chromosome','physical.pos','allele2','allele1','marker.ID','af','ld')]
  names(map) <- c('chr','pos','a0','a1','rsid','af','ld')
  
  saveRDS(
    map,
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/map.rds'
    )
  )
}

```

***

## Split reference


To use GWAS subsampling techniques
Only the correlations part needs to be derived using subset data
Follow X-wing strategy first - Split into three parts

- Subset the plink1 files into three parts
- Calculate LD matrices in subset 2
- Calculate freq files for subset 3

```{r}
# conda activate .snakemake/conda/329e0288cb99508f5e6c50a0996b234c_
# Load dependencies
library(GenoUtils)
library(data.table)
source('../functions/misc.R')
source_all('../functions')
  
library(dplyr)
library(bigreadr)
library(bigsnpr)
library(future.batchtools)
library(furrr)

# Create temp directory
tmp_dir<-tempdir()

opt<-list()
opt$ref_plink_chr<-'resources/data/ref/ref.chr'
opt$plink<-'plink'
opt$plink2<-'plink2'

for(pop in c('EUR','EAS','AFR')){
  opt$output <- paste0('/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/subset/', pop, '/', pop)
  opt$output_dir <- paste0(dirname(opt$output),'/')
  system(paste0('mkdir -p ',opt$output_dir))

  opt$ref_keep<-paste0('resources/data/ref/keep_files/', pop, '.keep')
  
  keep <- fread(opt$ref_keep)
  
  # Split reference into three groups
  lines <- readLines(opt$ref_keep)
  set.seed(1)
  shuffled_lines <- sample(lines)

  # Calculate the split points
  n <- length(shuffled_lines)
  split1 <- floor(n / 3)
  split2 <- 2 * split1
  
  # Split the lines into three groups
  keep1 <- shuffled_lines[1:split1]
  keep2 <- shuffled_lines[(split1 + 1):split2]
  keep3 <- shuffled_lines[(split2 + 1):n]
  
  # Write each group to a separate file
  write.table(data.frame(0, keep1), paste0(tmp_dir, "/keep1"), col.names = F, row.names = F, quote = F)
  write.table(data.frame(0, keep2), paste0(tmp_dir, "/keep2"), col.names = F, row.names = F, quote = F)
  write.table(data.frame(0, keep3), paste0(tmp_dir, "/keep3"), col.names = F, row.names = F, quote = F)

  for(subset_i in 1:3){
    # Save plink format files
    system(paste0(opt$plink, ' --bfile ', tmp_dir, '/ref_merge --keep ', tmp_dir, '/keep', subset_i, ' --make-bed --out ', opt$output, '.subset_', subset_i))
  }

  # Calculate freq files for subset 3, as required by LEOPARD
  for(subset_i in 3:3){
    system(paste0(opt$plink, ' --bfile ', opt$output, '.subset_', subset_i, ' --freq --out ', opt$output, '.subset_', subset_i))
  }

  # Read in 1KG+HGDP rds file
  # Subset to subset 2 individuals
  obj.hgdp_1kg <- snp_attach(paste0(
          '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
          pop,
          '/ref.',
          pop,
          '.GW.rds'
        ))
  G <- obj.hgdp_1kg$genotypes
  CHR <- obj.hgdp_1kg$map$chromosome
  POS <- obj.hgdp_1kg$map$physical.pos
  POS2 <- snp_asGeneticPos(CHR, POS, dir ='/users/k1806347/oliverpainfel/Data/Genetic_Map/CEU', ncores = nb_cores()) # Note this is not appropriate but I think Florian did the same as used distance from UKB
  MAF <- snp_MAF(G, ncores = nb_cores())
  keep <- (MAF > 0.02)

  plan(batchtools_slurm(resources = list(
    time = "12:00:00",
    ntasks = 5,
    mem = "20g",
    partition = "neurohack_cpu"
  )))
  
  all_final_grp <- furrr::future_map_dfr(1:22, function(chr) {
    ind.chr2 <- which(CHR == chr & keep)

    corr0 <- runonce::save_run({
      snp_cor(
        G,
        ind.col = ind.chr2,
        infos.pos = POS2[ind.chr2],
        size = 3 / 1000,
        ncores = nb_cores()
      )
    }, file = paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/subset',
      pop,
      '/LD_chr',
      chr,
      ".rds"
    ))
    
    # find nearly independent LD blocks
    # Playing around with the snp_ldsplit function, the max_r2 needs to be increased to 0.2 for AFR populations.
    # This might be because we are using an inappropriate recombination map (EUR)?
    # Use max_r2 = 0.15 for EUR and EAS as in Florian's code, but increase to 0.2 for AFR. Note. this is still lower than the default of 0.3.
    max_r2 <- ifelse(pop %in% c('EUR','EAS'), 0.15, 0.25)
    m <- length(ind.chr2)
    (SEQ <- round(seq_log(m / 30, m / 5, length.out = 20)))
    splits <- snp_ldsplit(corr0, thr_r2 = 0.05, min_size = 50, max_size = SEQ, max_r2 = max_r2)
    splits$cost2 <- sapply(splits$all_size, function(sizes) sum(sizes^2))
  
    best_split <- splits %>%
      arrange(cost2 * sqrt(5 + cost)) %>%
      print() %>%
      slice(1) %>%
      print()
  
    (all_size <- best_split$all_size[[1]])
    best_grp <- rep(seq_along(all_size), all_size)
  
    runonce::save_run({
      corr0T <- as(corr0, "dgTMatrix")
      corr0T@x <-
        ifelse(best_grp[corr0T@i + 1L] == best_grp[corr0T@j + 1L], corr0T@x, 0)
      as(Matrix::drop0(corr0T), "symmetricMatrix")
    }, file = paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/LD_with_blocks_chr',
      chr,
      ".rds"
    ))
  
    # return
    tibble(best_split, ind = list(ind.chr2))
  })
  
  saveRDS(
    all_final_grp,
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/all_final_grp.rds'
    )
  )
  
  ###
  # Create map.rds
  ###
  map <- obj.hgdp_1kg$map[keep,]

  # Compute allele frequency
  map$af<-big_colstats(G, ind.col = which(keep), ncores = nb_cores())$sum / (2 * nrow(G))
  
  # Compute LD scores
  map$ld <- do.call('c', lapply(1:22, function(chr) {
    cat(chr, ".. ", sep = "")
    corr_chr <- readRDS(
      paste0(
        '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
        pop,
        '/LD_chr',
        chr,
        ".rds"
      )
    )
    Matrix::colSums(corr_chr ^ 2)
  }))
  
  map <- map[, c('chromosome','physical.pos','allele2','allele1','marker.ID','af','ld')]
  names(map) <- c('chr','pos','a0','a1','rsid','af','ld')
  
  saveRDS(
    map,
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/map.rds'
    )
  )
  
  system(paste0('rm -r ', tmp_dir, '/*'))
}

```


