---
title: Preparing LDpred2 LD matrices
output: 
  html_document:
    theme: cosmo
    css: styles/styles.css
    includes:
      in_header: header.html
      after_body: footer.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{css, echo=F}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

***

# 1KG

Here we will calculate LD matrices for LDpred2 using HapMap3 SNPs in the 1KG super populations. This is to allow fair comparison across methods for a cross-population polygenic scoring study. 

The code will follow that used to generate the LD matrices for EUR, in this paper ([link](https://doi.org/10.1016/j.xhgg.2022.100136)).

```{bash}
# Subset 1KG into super populations and merge across chromosomes
for pop in $(echo EUR EAS AFR); do
  mkdir -p /users/k1806347/oliverpainfel/Data/1KG/ldpred2/${pop}

  # Create mergelist
  rm -f /users/k1806347/oliverpainfel/Data/1KG/ldpred2/${pop}/mergelist.txt
  for chr in $(seq 1 22); do
    echo ~/oliverpainfel/Data/1KG/GenoPred/v2.0.0/ref.chr${chr} >> /users/k1806347/oliverpainfel/Data/1KG/ldpred2/${pop}/mergelist.txt
  done
  
  ~/oliverpainfel/Software/plink1.9/plink \
    --merge-list /users/k1806347/oliverpainfel/Data/1KG/ldpred2/${pop}/mergelist.txt \
    --make-bed \
    --keep /users/k1806347/oliverpainfel/Data/1KG/GenoPred/v2.0.0/keep_files/${pop}.keep \
    --out /users/k1806347/oliverpainfel/Data/1KG/ldpred2/${pop}/ref.${pop}.GW
done
```

```{r}
library(bigsnpr)
library(bigreadr)

# Create object for genetic data in ldpred2 format, and impute missing values
for(pop in c('EUR', 'EAS', 'AFR')){
  # Read in reference data
  snp_readBed(
    paste0(
      '/users/k1806347/oliverpainfel/Data/1KG/ldpred2/',
      pop,
      '/ref.',
      pop,
      '.GW.bed'
    )
  )
  
  # Attach the ref object in R session
  ref <-
    snp_attach(
      paste0(
        '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
        pop,
        '/ref.',
        pop,
        '.GW.rds'
      )
    )
  G <- ref$genotypes

  #### Impute missing values (bigsnpr can't handle missing data in most functions)
  G_imp <- snp_fastImputeSimple(G, method = "mean2", ncores = nb_cores())

  # Save imputed reference
  ref$genotypes<-G_imp
  saveRDS(ref, paste0(
        '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
        pop,
        '/ref.',
        pop,
        '.GW.rds'
      ))
}

################
# Now calculate LD matrices following code from https://doi.org/10.1016/j.xhgg.2022.100136

library(dplyr)
library(bigreadr)
library(bigsnpr)
library(future.batchtools)
library(furrr)

for(pop in c('EUR', 'EAS', 'AFR')){
  obj.1000G <- snp_attach(paste0(
          '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
          pop,
          '/ref.',
          pop,
          '.GW.rds'
        ))
  G <- obj.1000G$genotypes
  CHR <- obj.1000G$map$chromosome
  POS <- obj.1000G$map$physical.pos
  POS2 <- snp_asGeneticPos(CHR, POS, dir ='/users/k1806347/oliverpainfel/Data/Genetic_Map/CEU', ncores = nb_cores()) # Note this is not appropriate but I think Florian did the same as used distance from UKB
  MAF <- snp_MAF(G, ncores = nb_cores())
  keep <- (MAF > 0.02)

  plan(batchtools_slurm(resources = list(
    time = "12:00:00",
    ntasks = 5,
    mem = "20g",
    partition = "neurohack_cpu"
  )))
  
  all_final_grp <- furrr::future_map_dfr(1:22, function(chr) {
    ind.chr2 <- which(CHR == chr & keep)

    corr0 <- runonce::save_run({
      snp_cor(
        G,
        ind.col = ind.chr2,
        infos.pos = POS2[ind.chr2],
        size = 3 / 1000,
        ncores = nb_cores()
      )
    }, file = paste0(
      '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
      pop,
      '/LD_chr',
      chr,
      ".rds"
    ))
    
    # find nearly independent LD blocks
    # Playing around with the snp_ldsplit function, the max_r2 needs to be increased to 0.2 for AFR populations.
    # This might be because we are using an inappropriate recombination map (EUR)?
    # Use max_r2 = 0.15 for EUR and EAS as in Florian's code, but increase to 0.2 for AFR. Note. this is still lower than the default of 0.3.
    max_r2 <- ifelse(pop %in% c('EUR','EAS'), 0.15, 0.25)
    m <- length(ind.chr2)
    (SEQ <- round(seq_log(m / 30, m / 5, length.out = 20)))
    splits <- snp_ldsplit(corr0, thr_r2 = 0.05, min_size = 50, max_size = SEQ, max_r2 = max_r2)
    splits$cost2 <- sapply(splits$all_size, function(sizes) sum(sizes^2))
  
    best_split <- splits %>%
      arrange(cost2 * sqrt(5 + cost)) %>%
      print() %>%
      slice(1) %>%
      print()
  
    (all_size <- best_split$all_size[[1]])
    best_grp <- rep(seq_along(all_size), all_size)
  
    runonce::save_run({
      corr0T <- as(corr0, "dgTMatrix")
      corr0T@x <-
        ifelse(best_grp[corr0T@i + 1L] == best_grp[corr0T@j + 1L], corr0T@x, 0)
      as(Matrix::drop0(corr0T), "symmetricMatrix")
    }, file = paste0(
      '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
      pop,
      '/LD_with_blocks_chr',
      chr,
      ".rds"
    ))
  
    # return
    tibble(best_split, ind = list(ind.chr2))
  })
  
  saveRDS(
    all_final_grp,
    paste0(
      '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
      pop,
      '/all_final_grp.rds'
    )
  )
  
  ###
  # Create map.rds
  ###
  map <- obj.1000G$map[keep,]

  # Compute allele frequency
  map$af<-big_colstats(G, ind.col = which(keep), ncores = nb_cores())$sum / (2 * nrow(G))
  
  # Compute LD scores
  map$ld <- do.call('c', lapply(1:22, function(chr) {
    cat(chr, ".. ", sep = "")
    corr_chr <- readRDS(
      paste0(
        '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
        pop,
        '/LD_chr',
        chr,
        ".rds"
      )
    )
    Matrix::colSums(corr_chr ^ 2)
  }))
  
  map <- map[, c('chromosome','physical.pos','allele2','allele1','marker.ID','af','ld')]
  names(map) <- c('chr','pos','a0','a1','rsid','af','ld')
  
  saveRDS(
    map,
    paste0(
      '/scratch/prj/oliverpainfel/Data/1KG/ldpred2/',
      pop,
      '/map.rds'
    )
  )
}


```

***

# 1KG+HGDP

Here we will calculate LD matrices for LDpred2 using HapMap3 SNPs in the 1KG+HGDP super populations. This is to allow fair comparison across methods for a cross-population polygenic scoring study. 

The code will follow that used to generate the LD matrices for EUR, in this paper ([link](https://doi.org/10.1016/j.xhgg.2022.100136)).

```{bash}
# Subset 1KG+HGDP into super populations and merge across chromosomes
for pop in $(echo EUR EAS AFR); do
  mkdir -p /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/${pop}
  mkdir -p /users/k1806347/oliverpainfel/Data/hgdp_1kg/prscs/ldblk_hgdp_1kg_$(echo ${pop} | tr '[:upper:]' '[:lower:]')

  # Create mergelist
  rm -f /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/${pop}/mergelist.txt
  for chr in $(seq 1 22); do
    echo /users/k1806347/oliverpainfel/Data/hgdp_1kg/genopred/ref/ref.chr${chr} >> /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/${pop}/mergelist.txt
  done
  
  /users/k1806347/oliverpainfel/Software/plink2_linux_avx2_20241011/plink2 \
    --pmerge-list /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/${pop}/mergelist.txt \
    --make-bed \
    --keep /users/k1806347/oliverpainfel/Data/hgdp_1kg/genopred/ref/keep_files/${pop}.keep \
    --out /users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/${pop}/ref.${pop}.GW
done
```

```{r}
library(bigsnpr)
library(bigreadr)

# Create object for genetic data in ldpred2 format, and impute missing values
for(pop in c('EUR', 'EAS', 'AFR')){
  # Read in reference data
  snp_readBed(
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/ref.',
      pop,
      '.GW.bed'
    )
  )
  
  # Attach the ref object in R session
  ref <-
    snp_attach(
      paste0(
        '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
        pop,
        '/ref.',
        pop,
        '.GW.rds'
      )
    )
  G <- ref$genotypes

  #### Impute missing values (bigsnpr can't handle missing data in most functions)
  G_imp <- snp_fastImputeSimple(G, method = "mean2", ncores = nb_cores())

  # Save imputed reference
  ref$genotypes<-G_imp
  saveRDS(ref, paste0(
        '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
        pop,
        '/ref.',
        pop,
        '.GW.rds'
      ))
}

################
# Now calculate LD matrices following code from https://doi.org/10.1016/j.xhgg.2022.100136

library(dplyr)
library(bigreadr)
library(bigsnpr)
library(future.batchtools)
library(furrr)

for(pop in c('EUR', 'EAS', 'AFR')){
  obj.hgdp_1kg <- snp_attach(paste0(
          '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
          pop,
          '/ref.',
          pop,
          '.GW.rds'
        ))
  G <- obj.hgdp_1kg$genotypes
  CHR <- obj.hgdp_1kg$map$chromosome
  POS <- obj.hgdp_1kg$map$physical.pos
  POS2 <- snp_asGeneticPos(CHR, POS, dir ='/users/k1806347/oliverpainfel/Data/Genetic_Map/CEU', ncores = nb_cores()) # Note this is not appropriate but I think Florian did the same as used distance from UKB
  MAF <- snp_MAF(G, ncores = nb_cores())
  keep <- (MAF > 0.02)

  plan(batchtools_slurm(resources = list(
    time = "12:00:00",
    ntasks = 5,
    mem = "20g",
    partition = "neurohack_cpu"
  )))
  
  all_final_grp <- furrr::future_map_dfr(1:22, function(chr) {
    ind.chr2 <- which(CHR == chr & keep)

    corr0 <- runonce::save_run({
      snp_cor(
        G,
        ind.col = ind.chr2,
        infos.pos = POS2[ind.chr2],
        size = 3 / 1000,
        ncores = nb_cores()
      )
    }, file = paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/LD_chr',
      chr,
      ".rds"
    ))
    
    # find nearly independent LD blocks
    # Playing around with the snp_ldsplit function, the max_r2 needs to be increased to 0.2 for AFR populations.
    # This might be because we are using an inappropriate recombination map (EUR)?
    # Use max_r2 = 0.15 for EUR and EAS as in Florian's code, but increase to 0.2 for AFR. Note. this is still lower than the default of 0.3.
    max_r2 <- ifelse(pop %in% c('EUR','EAS'), 0.15, 0.25)
    m <- length(ind.chr2)
    (SEQ <- round(seq_log(m / 30, m / 5, length.out = 20)))
    splits <- snp_ldsplit(corr0, thr_r2 = 0.05, min_size = 50, max_size = SEQ, max_r2 = max_r2)
    splits$cost2 <- sapply(splits$all_size, function(sizes) sum(sizes^2))
  
    best_split <- splits %>%
      arrange(cost2 * sqrt(5 + cost)) %>%
      print() %>%
      slice(1) %>%
      print()
  
    (all_size <- best_split$all_size[[1]])
    best_grp <- rep(seq_along(all_size), all_size)
  
    runonce::save_run({
      corr0T <- as(corr0, "dgTMatrix")
      corr0T@x <-
        ifelse(best_grp[corr0T@i + 1L] == best_grp[corr0T@j + 1L], corr0T@x, 0)
      as(Matrix::drop0(corr0T), "symmetricMatrix")
    }, file = paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/LD_with_blocks_chr',
      chr,
      ".rds"
    ))
  
    # LD for PRS-CS (based on the same data and LD blocks)
    hdf5_file <- paste0("/users/k1806347/oliverpainfel/Data/hgdp_1kg/prscs/ldblk_hgdp_1kg_", tolower(pop)  , "/ldblk_hgdp_1kg_chr", chr, ".hdf5")
    runonce::skip_run_if({
      ind_block <- split(ind.chr2, best_grp)
      bigparallelr::set_blas_ncores(nb_cores())
      for (ic in seq_along(ind_block)) {
        ind <- ind_block[[ic]]
        ld <- big_cor(G, ind.col = ind)[]
        rhdf5::h5write(list(ldblk = ld, snplist = paste0("snp", ind)),
                       file = hdf5_file, name = paste0("blk_", ic), level = 9)
      }
    }, files = hdf5_file)
  
    # return
    tibble(best_split, ind = list(ind.chr2))
  })
  
  saveRDS(
    all_final_grp,
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/all_final_grp.rds'
    )
  )
  
  ###
  # Create map.rds
  ###
  map <- obj.hgdp_1kg$map[keep,]

  # Compute allele frequency
  map$af<-big_colstats(G, ind.col = which(keep), ncores = nb_cores())$sum / (2 * nrow(G))
  
  # Compute LD scores
  map$ld <- do.call('c', lapply(1:22, function(chr) {
    cat(chr, ".. ", sep = "")
    corr_chr <- readRDS(
      paste0(
        '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
        pop,
        '/LD_chr',
        chr,
        ".rds"
      )
    )
    Matrix::colSums(corr_chr ^ 2)
  }))
  
  map <- map[, c('chromosome','physical.pos','allele2','allele1','marker.ID','af','ld')]
  names(map) <- c('chr','pos','a0','a1','rsid','af','ld')
  
  saveRDS(
    map,
    paste0(
      '/users/k1806347/oliverpainfel/Data/hgdp_1kg/ldpred2/',
      pop,
      '/map.rds'
    )
  )
  
  # GOT to here
  
  # need also some .bim and some snpinfo for PRS-CS
  ind_keep <- unlist(all_final_grp$ind)
  snpinfo <- transmute(obj.hgdp_1kg$map, chromosome, marker.ID = paste0("snp", row_number()),
                       genetic.dist = 0, physical.pos, allele1 = "A", allele2 = "C")[ind_keep, ]
  bigsnpr:::write.table2(snpinfo, "/users/k1806347/oliverpainfel/Data/hgdp_1kg/prscs/ldblk_hgdp_1kg_", tolower(pop)  , "/for_prscs.bim")
  
  snpinfo2 <- transmute(snpinfo, CHR = chromosome, SNP = marker.ID,
                        BP = physical.pos, A1 = allele1, A2 = allele2)
  snpinfo2$MAF <- snp_MAF(G, ind.row = ind_EAS, ind.col = ind_keep, ncores = nb_cores())
  bigreadr::fwrite2(snpinfo2, "/users/k1806347/oliverpainfel/Data/hgdp_1kg/prscs/ldblk_hgdp_1kg_", tolower(pop)  , "/snpinfo_hgdp_1kg_hm3", sep = "\t")
}


```





