---
title: "PTS portability"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 2
    css: styles/styles.css
    includes:
      in_header: header.html
      after_body: footer.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(knitr)
```

***

# Phase 1: Data Preparation & Harmonization

TWAS models have already been created based on EUR eQTLGen sumstats and AFR SABR sumstats.
UKB sample has already been split into EUR training and EUR/AFR test subsets, and GWAS for 10 traits have been conducted in the training subset.

***

# Phase 2: TWAS & Gene Expression Prediction

***

## 2.1 Run TWAS (Discovery)

Run a TWAS using the eQTLGen sumstats and EUR UKB GWAS sumstats.

```{bash}

conda activate ldsc
mkdir -p /users/k1806347/oliverpainfel/Data/GWAS_sumstats/mind_the_gap/ldsc_munged
for pheno in $(cut -f 1 -d' ' /users/k1806347/oliverpainfel/Analyses/mind_the_gap/GenoPred/config/gwas_list_mvp.txt | tail -n +2); do
sbatch -p neurohack_cpu,interruptible_cpu -t 1:00:00 --wrap="
~/oliverpainfel/Software/ldsc/munge_sumstats.py \
  --sumstats /users/k1806347/oliverpainfel/Analyses/mind_the_gap/GenoPred/output_sbayesrc_sparse/reference/gwas_sumstat/UKB_WGS_${pheno}/UKB_WGS_${pheno}-cleaned.gz \
  --out /users/k1806347/oliverpainfel/Data/GWAS_sumstats/mind_the_gap/ldsc_munged/${pheno} \
  --chunksize 50000 \
  --merge-alleles /users/k1806347/oliverpainfel/Data/ldsc/w_hm3.snplist
"
done

conda activate eqtl_to_twas
for pheno in $(cut -f 1 -d' ' /users/k1806347/oliverpainfel/Analyses/mind_the_gap/GenoPred/config/gwas_list_mvp.txt | tail -n +2); do
mkdir -p /users/k1806347/oliverpainfel/Analyses/pts_portability/twas/${pheno}

for chr in $(seq 1 22); do
sbatch -p neurohack_cpu,interruptible_cpu -t 1:00:00 --wrap="
Rscript /users/k1806347/oliverpainfel/Software/MyGit/eQTL_to_TWAS/FUSION.assoc_test.edit.R \
    --sumstats /users/k1806347/oliverpainfel/Data/GWAS_sumstats/mind_the_gap/ldsc_munged/${pheno}.sumstats.gz \
    --weights /users/k1806347/oliverpainfel/Data/eQTLGen/eQTL_to_TWAS/eQTLGen.eQTL.pos \
    --weights_dir /users/k1806347/oliverpainfel/Data/eQTLGen/eQTL_to_TWAS/ \
    --ref_ld_chr /users/k1806347/oliverpainfel/Data/1KG/FUSION/1000G.EUR. \
    --out /users/k1806347/oliverpainfel/Analyses/pts_portability/twas/${pheno}/${pheno}.eqtlgen.chr${chr} \
    --chr ${chr}
"
done
done

```

```{r}
# Combine per chromosome TWAS results
phenos <- fread('/users/k1806347/oliverpainfel/Analyses/mind_the_gap/GenoPred/config/gwas_list_mvp.txt')$name

for(pheno in phenos){
  res <- NULL
  for(i in 1:22){
    res <- rbind(
      res,
      fread(
        paste0(
          '/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/',pheno,'/', pheno,'.eqtlgen.chr', i
        )
      )
    )
  }
  fwrite(res, 
         paste0(
          '/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/',pheno,'/', pheno,'.eqtlgen.gw.txt.gz'
         ),
         sep = ' ', quote = F, na = 'NA')
}
```

```{bash}
for pheno in $(cut -f 1 -d' ' /users/k1806347/oliverpainfel/Analyses/mind_the_gap/GenoPred/config/gwas_list_mvp.txt | tail -n +2); do
  dx upload /users/k1806347/oliverpainfel/Analyses/pts_portability/twas/${pheno}/${pheno}.eqtlgen.gw.txt.gz --path pts_portability/
done
```

***

## 2.2 Predict Expression (Reference)

```{bash}
conda activate eqtl_to_twas

Rscript ~/MyGit/Predicting-TWAS-features/FeaturePred.V2.0.R \
	--PLINK_prefix_chr /users/k1806347/oliverpainfel/Data/1KG/FUSION/1000G.EUR. \
  --weights /users/k1806347/oliverpainfel/Data/eQTLGen/eQTL_to_TWAS/eQTLGen.eQTL.pos \
  --weights_dir /users/k1806347/oliverpainfel/Data/eQTLGen/eQTL_to_TWAS/ \
	--ref_ld_chr /users/k1806347/oliverpainfel/Data/1KG/FUSION/1000G.EUR. \
	--plink /users/k1806347/oliverpainfel/Software/plink1.9/plink \
	--ref_maf /users/k1806347/oliverpainfel/Data/1KG/FUSION/1000G.EUR. \
	--pigz /users/k1806347/oliverpainfel/Software/pigz \
	--n_cores 50 \
	--model prscs \
	--memory 80000 \
	--output /users/k1806347/oliverpainfel/Data/1KG/pred_exp/EUR.GW
	
```

***

# Phase 3: Construct PTS (OmicScores)

```{r}
# Combine per chromosome TWAS results
phenos <- fread('/users/k1806347/oliverpainfel/Analyses/mind_the_gap/GenoPred/config/gwas_list_mvp.txt')$name

for(pheno in phenos){
  
  # Restrict to prscs model
  twas <-
    fread(
      paste0(
        '/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/',
        pheno,
        '/',
        pheno,
        '.eqtlgen.gw.txt.gz'
      )
    )
  twas<-twas[twas$MODEL == 'prscs',]
  fwrite(twas, 
         paste0(
           '/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/',
           pheno,
           '/',
           pheno,
           '.eqtlgen.prscs.gw.txt.gz'
         ),
         sep = ' ', quote = F, na = 'NA')
}
```

```{bash}
# Create the score file using IFRisk
mkdir -p /users/k1806347/oliverpainfel/Analyses/pts_portability/pts/eqtlgen/eur/

for pheno in $(cut -f 1 -d' ' /users/k1806347/oliverpainfel/Analyses/mind_the_gap/GenoPred/config/gwas_list_mvp.txt | tail -n +2); do
  sbatch -p neurohack_cpu,interruptible_cpu -t 1:00:00 --wrap="
    Rscript ~/MyGit/Inferred-functional-risk-scoring/IFRisk.V1.0.R \
      --twas_results /users/k1806347/oliverpainfel/Analyses/pts_portability/twas/${pheno}/${pheno}.eqtlgen.prscs.gw.txt.gz \
      --target_gene_exp /users/k1806347/oliverpainfel/Data/1KG/pred_exp/EUR.GW/FeaturePredictions_eQTLGen.eQTL.txt.gz \
      --clump_thresh 0.1 \
      --output /users/k1806347/oliverpainfel/Analyses/pts_portability/pts/eqtlgen/eur/${pheno}.eqtlgen.eur
      
    dx upload /users/k1806347/oliverpainfel/Analyses/pts_portability/pts/eqtlgen/eur/${pheno}.eqtlgen.eur.score --path pts_portability/
  "
done

```

***

## Experiment with restrict results to genes with significant prediction

```{r}
# Restrict to genes that show significant positive prediction in GTEx
twas<-fread('/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/HT/HT.eqtlgen.gw.txt.gz')
twas<-twas[twas$MODEL == 'prscs',]
sig_genes<-fread('/users/k1806347/oliverpainfel/Data/eQTLGen/eQTL_to_TWAS/eQTLGen.eQTL.sig.pos')
twas<-twas[twas$ID %in% sig_genes$ID,]
fwrite(twas, 
       paste0(
        '/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/',pheno,'/', pheno,'.eqtlgen.prscs.sig.gw.txt.gz'
       ),
       sep = ' ', quote = F, na = 'NA')
```

```{bash}
Rscript ~/MyGit/Inferred-functional-risk-scoring/IFRisk.V1.0.R \
  --twas_results /users/k1806347/oliverpainfel/Analyses/pts_portability/twas/HT/HT.eqtlgen.prscs.sig.gw.txt.gz \
  --target_gene_exp /users/k1806347/oliverpainfel/Data/1KG/pred_exp/EUR.GW/FeaturePredictions_eQTLGen.eQTL.txt.gz \
  --clump_thresh 0.1 \
  --output /users/k1806347/oliverpainfel/Analyses/pts_portability/pts/eqtlgen/eur/HT.eqtlgen.sig.eur
  
dx upload /users/k1806347/oliverpainfel/Analyses/pts_portability/pts/eqtlgen/eur/HT.eqtlgen.sig.eur.score --path pts_portability/

```

Note. using height, this filter reduced the variance explained of PTS.

***

## Experiment with restricting results to genes present in both eQTLGen and SABR

```{r}
# Restrict to genes that show significant positive prediction in GTEx
twas<-fread('/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/HT/HT.eqtlgen.gw.txt.gz')
twas<-twas[twas$MODEL == 'prscs',]
sabr_genes<-fread('/users/k1806347/oliverpainfel/Data/SABR/processed/eqtl_to_twas/SABR.eQTL.pos')
twas<-twas[twas$ID %in% sabr_genes$ID,]
fwrite(twas, 
       paste0(
        '/users/k1806347/oliverpainfel/Analyses/pts_portability/twas/',pheno,'/', pheno,'.eqtlgen.prscs.in_sabr.gw.txt.gz'
       ),
       sep = ' ', quote = F, na = 'NA')
```

```{bash}
Rscript ~/MyGit/Inferred-functional-risk-scoring/IFRisk.V1.0.R \
  --twas_results /users/k1806347/oliverpainfel/Analyses/pts_portability/twas/HT/HT.eqtlgen.prscs.in_sabr.gw.txt.gz \
  --target_gene_exp /users/k1806347/oliverpainfel/Data/1KG/pred_exp/EUR.GW/FeaturePredictions_eQTLGen.eQTL.txt.gz \
  --clump_thresh 0.1 \
  --output /users/k1806347/oliverpainfel/Analyses/pts_portability/pts/eqtlgen/eur/HT.eqtlgen.in_sabr.eur
  
dx upload /users/k1806347/oliverpainfel/Analyses/pts_portability/pts/eqtlgen/eur/HT.eqtlgen.in_sabr.eur.score --path pts_portability/

```
